\documentclass[11pt]{article}

\input{macros}

\usepackage[margin=1in]{geometry}

\title{VegasCore: A Technical Report on the Current Mechanization}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
VegasCore is a Lean~4 development for expressing finite strategic games as
programs.  A Vegas program combines ordinary deterministic computation,
finite-support probabilistic sampling, strategic commitment, and explicit
reveal of private values.  The central design idea is that information
structure is represented by the program context: each binding is either
public, or hidden and owned by one player.  Player strategies are then typed
over that player's view of the context, rather than over the full program
state, in the user-facing legal pure and behavioral strategy spaces.

The current repository provides a checked-program boundary, legal strategy
spaces, denotational and trace semantics, strategic-form game wrappers,
protocol-level solution concepts, a canonical sequential denotation through
factored-observation stochastic games, a finite Kuhn-style
mixed-to-behavioral realization theorem with a native guard-legal Vegas
behavioral witness, preserving outcome distributions, and a MAID backend used
for an additional graphical-game compilation route.  This
report is intended as a map of what is implemented and why the main design
choices were made.  It deliberately avoids cataloguing every internal lemma:
the useful level of detail is the shape of the objects, the invariants they
carry, and the theorems a Vegas user can rely on.
\end{abstract}

\tableofcontents

%% =====================================================================
\section{Orientation}\label{sec:orientation}
%% =====================================================================

VegasCore sits between two worlds.  On one side is a programming-language
view of games: a game is a program with variables, contexts, sampling,
branching-by-strategic-choice, and final payoff computation.  On the other
side is game theory: strategies, outcome kernels, expected utility,
equilibrium concepts, and sequential-game theorems such as Kuhn's theorem.
The repository's job is to make those two views agree in Lean.

\paragraph{What a Vegas user gets.}
The main user-facing object is a checked program:
\[
  g : \WFP.
\]
From such a program the repository exposes:
\begin{itemize}[nosep]
\item guard-legal behavioral and pure strategy spaces;
\item strategic-form kernel games for behavioral and pure strategies;
\item Vegas-facing definitions of Nash equilibrium, best response, dominance,
      correlated equilibrium, welfare, approximate Nash, potential games,
      zero-sum games, and related properties;
\item protocol-level versions of the same solution concepts, stated directly
      on the Vegas outcome semantics, with correspondence theorems to the
      strategic-form wrappers;
\item a finite mixed-to-behavioral realization theorem preserving the
      distribution over final payoff outcomes, with a native Vegas
      behavioral witness;
\item a MAID backend that compiles the program to a graphical representation
      and proves outcome preservation for that backend.
\end{itemize}

\begin{figure}[ht]
\centering
\fbox{\parbox{0.92\textwidth}{
\begin{center}
\small
\begin{tabular}{@{}rcl@{}}
Vegas program & \(\longrightarrow\) & \(\WFP\) \\[4pt]
\(\WFP\) & \(\xrightarrow{\;\toKG,\;\toStrategicKG\;}\) &
  strategic form: Nash, dominance, welfare \\[4pt]
\(\WFP\) & \(\xrightarrow{\;\mathrm{toFOSG}\;}\) &
  sequential form: bounded horizon, Kuhn \\[4pt]
\(\WFP\) & \(\xrightarrow{\;\mathrm{MAID}\;}\) &
  graphical backend: outcome preservation
\end{tabular}
\end{center}
}}
\caption{The main denotational routes.  \WFP{} is the boundary where raw
syntax becomes a checked game.  Kernel games provide strategic-form concepts;
FOSG provides the canonical sequential denotation; MAID remains a graphical
backend.}
\label{fig:pipeline}
\end{figure}

\paragraph{What is intentionally internal.}
The development contains a substantial amount of proof infrastructure:
cursor-indexed program states, suffix witnesses, dependent transports between
program fragments, executable MAID compilation states, FDist/PMF bridges, and
finite-history constructions for the GameTheory library.  Those are not
separate user concepts.  They exist to justify the public surfaces listed
above.  In this report they appear only where they explain a design decision
or a theorem boundary.

\paragraph{Current scale.}
At the time of writing, the Vegas tree contains roughly 24,000 lines of Lean
across 42 files.  The GameTheory submodule contains roughly 44,500 lines of
Lean across 169 files.  The split matters: VegasCore defines the programming
language and its denotations; GameTheory supplies generic game-theoretic
objects and theorems that Vegas instantiates.


%% =====================================================================
\section{The Programming Model}\label{sec:model}
%% =====================================================================

\subsection{Contexts, Visibility, and Views}

The language is parameterized by a player type $\Players$ and an expression
language interface $L$.  A Vegas context is a list of variable bindings:
\[
  \Gamma ::= [] \mid (x,\beta)::\Gamma.
\]
The binding type $\beta$ records both the value type and its visibility:
\[
  \beta ::= \pub\;\tau \mid \hid(i,\tau),
\]
where $\tau$ is a type of the expression language and $i$ is a player.

There are three projections that occur throughout the development:
\begin{itemize}[nosep]
\item $\Gamma\!\!\downarrow_i$ is player $i$'s view: public bindings plus
      hidden bindings owned by $i$;
\item $\Gamma^{\pub}$ is the public subcontext;
\item $\|\Gamma\|$ erases visibility and keeps only ordinary expression
      types.
\end{itemize}

The point of these projections is not just notation.  They determine which
environment is available to which part of the semantics.  Public computation
uses the public erasure; guard evaluation uses the full erased environment;
player strategies use the player's view.  This separation is the core
information-flow discipline of Vegas.

\subsection{Program Constructs}

The core language has five constructs:
\[
\begin{array}{rcl}
p &::=& \Ret\;\mathit{payoffs}\\
  &\mid& \Let\;x = e\;\mathbf{in}\;p\\
  &\mid& \Sample\;x \sim D\;\mathbf{in}\;p\\
  &\mid& \Commit\;x\;\mathbf{by}\;i\;\mathbf{where}\;R\;\mathbf{in}\;p\\
  &\mid& \Reveal\;y\;\mathbf{from}\;x\;\mathbf{in}\;p.
\end{array}
\]

\paragraph{Return.}
$\Ret$ terminates the game with one integer-valued payoff expression per
player.  Payoff expressions are typed over the public context only.  A payoff
cannot directly inspect a still-hidden secret.

\paragraph{Let.}
$\Let$ binds a deterministic public expression.  It is administratively useful
even though it has no strategic content.  One could imagine desugaring it to
a sample from a point-mass distribution, but keeping it separate avoids
requiring the expression language to expose point-mass distribution
expressions or substitution operations.  It also preserves sharing: a named
public computation can be referenced many times downstream without duplicating
the expression into every use site.

\paragraph{Sample.}
$\Sample$ is public chance.  It draws from a finite-support weighted
distribution expression evaluated in the current public environment and binds
the result as a public variable.

\paragraph{Commit.}
$\Commit$ is strategic choice.  Player $i$ chooses a value of type $\tau$
subject to guard $R$, and the chosen value is bound as hidden state owned by
$i$.  The guard is typed over the full erased context extended by the proposed
value.  A separate view-scoping judgment then requires the guard's
dependencies to lie in the proposed value plus player $i$'s view.

That design is deliberate.  Typing guards directly over
$\Gamma\!\!\downarrow_i$ would make the core syntax depend more deeply on the
committing player and would force more dependent transports through the
semantics.  The current design keeps raw syntax simple and puts the
information-flow obligation in a decidable well-formedness predicate.

\paragraph{Reveal.}
$\Reveal$ makes a hidden value public by binding a fresh public copy.  Reveal
is an explicit program construct because the timing of information disclosure
is strategic.  Committing both players' moves before either reveal models a
simultaneous move; revealing one player's move before another player commits
models sequential information.

\subsection{Examples}

\paragraph{Matching pennies.}
The simultaneous version commits both choices first and reveals them only
afterward:
\[
\begin{array}{l}
\Commit\;a\;\mathbf{by}\;\pl{0}\;\mathbf{where}\;\mathit{true}\\
\Commit\;b\;\mathbf{by}\;\pl{1}\;\mathbf{where}\;\mathit{true}\\
\Reveal\;a'\;\mathbf{from}\;a\\
\Reveal\;b'\;\mathbf{from}\;b\\
\Ret\;[\pl{0}\mapsto \mathbf{if}\;a'=b'\;\mathbf{then}\;1\;\mathbf{else}\;-1,\;
      \pl{1}\mapsto \mathbf{if}\;a'=b'\;\mathbf{then}\;-1\;\mathbf{else}\;1].
\end{array}
\]
At player~1's commit site, player~0's hidden value is not in player~1's
view.  Therefore player~1's strategy cannot depend on it.  The simultaneity is
not a convention; it is enforced by the type of strategies.

\paragraph{Sequential reveal.}
If the program reveals $a$ before player~1 commits, then $a'$ is public at
player~1's commit site.  Player~1's strategy type now permits dependence on
$a'$.  The difference between simultaneous and sequential play is represented
by program order and visibility, not by changing the game-theory layer.

\paragraph{Guarded choice.}
A guard can make the set of legal actions depend on visible state.  For
example, after revealing $a'$, player~1 can have a commit guard depending on
$a'$.  If for some visible state no action satisfies the guard, the program is
not legal as a game.  This is why legality is a separate checked obligation:
view scoping says the guard only reads allowed information; legality says the
guard does not deadlock the protocol.


%% =====================================================================
\section{Checked Programs}\label{sec:wfprogram}
%% =====================================================================

The raw core syntax is intentionally permissive.  It is an inductive family
indexed by context, so ordinary expression scoping is intrinsic, but several
semantic obligations are extrinsic.  A raw program can use duplicate variable
names, leave secrets unrevealed, use a guard that depends on another player's
hidden state, or contain a sample distribution whose mass is not one.  Those
terms are useful for metatheory by structural recursion, but they should not
be exported as game-theoretic objects.

\subsection{The Bundle Boundary}

The public game API consumes:
\[
  g : \WFP.
\]
The bundle contains:
\[
  (\Gamma,\;p,\;\rho,\;\mathsf{wf},\;\mathsf{normalized},\;\mathsf{legal}).
\]
Here $\Gamma$ is the initial context, $p$ is the program, $\rho$ is the
initial environment, and the witnesses assert:
\begin{itemize}[nosep]
\item $\WF(p)$: fresh bindings, reveal-completeness, and view scoping;
\item $\NormDists(p)$: every sample distribution has total mass one;
\item $\Legal(p)$: every commit site admits at least one guard-satisfying
      action in every relevant environment.
\end{itemize}

This bundle is not just packaging.  It is the boundary between raw syntax and
coherent games.  Earlier APIs allowed a raw program plus a normalization proof
to be turned into a kernel game.  That was too weak: a view-violating or
duplicate-name program could be mathematically assigned an outcome kernel even
though it did not describe a coherent protocol.  Requiring \WFP{} at all
user-facing game-theoretic entry points closes that gap.

\subsection{The Three Structural Obligations}

\paragraph{Fresh bindings.}
Freshness is an SSA discipline.  Every \(\Let\), \(\Sample\), \(\Commit\),
and \(\Reveal\) introduces a variable name that is fresh for the current
context.  This is not only surface-language hygiene.  Several backends rely
on the fact that a variable identifier names one event in the program.  In
the action graph it becomes a node identifier; in MAID compilation it becomes
part of the dependency map; in FOSG proofs it helps relate the current
environment to the visible observation.

\paragraph{Reveal-completeness.}
Every committed hidden value must eventually be revealed exactly once before
the program returns.  This is stronger than merely requiring payoffs to be
publicly scoped.  It gives a simple closed-world reading to a terminated
Vegas execution: all strategically chosen values have become part of the
public transcript by the time final payoffs are evaluated.  It also avoids a
class of awkward backend states in which a secret exists operationally but has
no public interpretation at termination.

\paragraph{View scoping.}
View scoping is the information-flow condition for commit guards.  A guard is
syntactically typed over the full erased environment, but the dependency
tracker must prove that it actually depends only on the proposed action and
the committing player's view.  This division is important.  The evaluator can
remain uniform and pass the full erased environment to guards, while the
well-formedness proof certifies that hidden information owned by other players
does not affect the result.

\subsection{Normalization and Legality}

The other two \WFP{} fields are not structural in the same sense, but they are
needed for game-theoretic interpretation.

\paragraph{Normalized distributions.}
Every \(\Sample\) site must evaluate to total mass one in every public
environment.  This theorem is combined with normalized strategy kernels to
show that the final outcome distribution has total mass one.  Only then can
the FDist-valued semantics be promoted to a PMF-valued outcome kernel.

\paragraph{Program legality.}
Every commit site must have at least one guard-satisfying action in every
erased environment.  This is a no-deadlock condition.  It does not choose an
action and it does not require all actions to be legal; it only says the
protocol never presents a player with an empty action set.

\subsection{Why Not Bundle Everything Internally?}

Internal semantic functions still operate on raw programs plus the specific
facts they need.  This is the right shape for proofs by induction.  When a
semantic proof descends into a continuation, it naturally has a subprogram in
an extended context.  Reconstructing a fresh \WFP{} bundle at every recursive
call would add dependent bookkeeping without changing the theorem being
proved.  The current architecture is therefore:
\begin{itemize}[nosep]
\item raw program semantics for metatheory;
\item checked program bundles for exported game-theoretic objects.
\end{itemize}

\subsection{Program Legality versus Strategy Legality}

$\Legal(p)$ is a feasibility property of the program: every commit site has
some legal action.  It does not say that every strategy chooses legal actions.
That second property is enforced by the strategy types used by the public
game APIs.

For a checked program $g$, the exported behavioral and pure strategy spaces
are subtypes:
\[
  \mathcal{B}^{\Legal}_i(g),
  \qquad
  \mathcal{P}^{\Legal}_i(g).
\]
A behavioral strategy is legal when, at every owned commit site and every
reachable environment, its distribution is supported on values satisfying the
commit guard.  A pure strategy is legal when its chosen value satisfies the
guard in the same sense.  Thus a Nash equilibrium of a Vegas program is an
equilibrium among protocol-legal strategies, not among arbitrary functions
that could recommend rejected moves.

This distinction is easy to miss, but it is one of the central API decisions
in the current repository.  If the strategy type were not refined, a
``deviation'' in a Nash statement could choose an action that the protocol
would reject.  The resulting theorem would be formally true about the wrapper
game but not about the intended protocol.  By making strategy legality part
of the public strategy type, the usual game-theoretic quantifiers range over
exactly the admissible strategic behavior.


%% =====================================================================
\section{Outcome Semantics}\label{sec:outcomes}
%% =====================================================================

\subsection{Finite Weighted Distributions}

The main semantics uses $\FDist{\alpha}$, a finitely-supported distribution
with nonnegative rational weights.  It is a sub-probability object: the type
does not require total mass one.  Normalization is a theorem, not a field in
every intermediate term.

This design pays off in the denotational semantics.  A program's outcome
distribution is built by repeated monadic bind.  If the semantics used
$\PMF{}$ directly, every intermediate bind would require a normalization
proof.  With $\FDist{}$, the algebraic semantics stays simple and the final
normalization theorem promotes the result to $\PMF{}$ when a game-theoretic
kernel needs an actual probability distribution.

The tradeoff is that not every PMF-valued object can be represented as an
$\FDist{}$.  A $\PMF{}$ may have real-valued weights or arise from the
GameTheory library.  Consequently the repository has both FDist-based
behavioral semantics and PMF-valued strategy/output bridges.  This duplication
is not accidental; it reflects the boundary between a rational, executable
language semantics and the real-valued probability objects used by general
game theory.

\subsection{Denotational Semantics}

Given an operational profile $\sigma$ and environment $\rho$, the
denotational semantics returns:
\[
  \densub{p}{\sigma}(\rho) : \FDist{\Outcome}.
\]
The clauses follow the program structure:
\begin{itemize}[nosep]
\item $\Ret$ evaluates public payoff expressions and returns a point mass;
\item $\Let$ evaluates a public expression and continues in the extended
      environment;
\item $\Sample$ binds over the sampled distribution;
\item $\Commit$ binds over the profile's kernel for the active player and
      commit site;
\item $\Reveal$ copies a hidden value into a new public binding.
\end{itemize}

The semantics is intentionally evaluator-facing: commit kernels are functions
of the full erased environment.  View-based game strategies are converted to
operational kernels by projecting that environment to the player's view before
calling the strategy.  This keeps the low-level evaluator simple while still
making information restrictions structural at the strategy layer.

\subsection{Normalization Theorem}

The main mass theorem has the following shape:
\[
  \NormDists(p) \;\wedge\; \NormOn(\sigma,p)
  \quad\Longrightarrow\quad
  \mathrm{totalWeight}(\densub{p}{\sigma}(\rho)) = 1.
\]
\(\NormDists\) covers the program's chance moves, while \(\NormOn\) covers
the operational strategy kernels at commit sites.  This separation is useful:
normalization of the program is a property of \(p\), while normalization of
strategic behavior is a property of the profile currently being evaluated.
The public checked-program APIs hide most of this split because their legal
strategy spaces already carry normalized kernels.

\subsection{Trace Semantics and Reachability}

The trace semantics records one choice at each probabilistic or strategic
branch.  Deterministic steps carry no choice data.  A trace has:
\begin{itemize}[nosep]
\item an outcome, obtained by replaying the recorded choices;
\item a weight, obtained by multiplying the selected sample and commit
      weights;
\item a legality predicate, requiring sampled values to lie in distribution
      support and committed values to satisfy guards.
\end{itemize}

The central trace theorem is adequacy: the denotational weight of an outcome
equals the sum of weights of traces producing that outcome.  A related support
theorem identifies the support of the outcome distribution with the
inductively reachable outcomes.  These results are important because they
connect the algebraic monadic semantics to the operational reading of a game
execution.

In formula form, adequacy says:
\[
  \densub{p}{\sigma}(\rho)(o)
  =
  \sum_{\substack{t\in\Tr{p}\\ \trout(t,\rho)=o}}
    \trweight{\sigma}(t,\rho).
\]
Support correctness says that \(o\) is assigned positive weight exactly when
there is a reachable trace producing \(o\).  Fair-play lemmas then connect
positive-weight traces to guard legality: if a profile is supported on
guard-satisfying actions, every positive-weight execution is a legal
protocol execution.

\subsection{Commutativity and Action Graphs}

Vegas programs are written sequentially, but some adjacent constructs are
independent.  The repository proves commutativity theorems showing that
independent commits and independent reveals can be swapped without changing
the outcome distribution.  These theorems rely on two ingredients:
\begin{itemize}[nosep]
\item expression dependency tracking supplied by the expression language
      interface;
\item commutativity of independent finite-distribution binds.
\end{itemize}

The action-graph semantics makes this order-independence explicit.  A program
can be flattened into a dependency DAG of let, sample, commit, and reveal
events.  Any ready event can fire once its dependencies are available, and
complete executions that assign the same values to nodes produce the same
final environment.  This layer is useful conceptually because it explains why
the textual order of independent program fragments is not always
game-theoretically meaningful.

The action graph is not currently the primary exported game object.  Its role
is explanatory and structural: it records the dependency skeleton that is also
visible in the MAID backend, and it justifies thinking about Vegas programs as
partially ordered protocols rather than merely as syntax trees.


%% =====================================================================
\section{Strategic Views}\label{sec:strategic}
%% =====================================================================

\subsection{Behavioral Strategic Form}

For a checked program $g$, VegasCore constructs a kernel game:
\[
  \toKG(g) : \KG.
\]
The strategy type for player $i$ is the legal behavioral strategy subtype
$\mathcal{B}^{\Legal}_i(g)$.  The outcome type is the Vegas payoff outcome
type, and the outcome kernel is obtained by running the Vegas behavioral
outcome semantics from $g$'s initial environment and promoting the normalized
FDist result to a PMF.

The expected utility theorem says that the expected utility in this kernel
game is exactly the expected payoff computed from the Vegas outcome
distribution.  This theorem is the justification for transporting
game-theoretic definitions through the bridge.

\paragraph{The strategy shape.}
A behavioral strategy for player \(i\) is not a function from full histories
or full environments.  At a commit site in context \(\Gamma\), it is a kernel
from the view environment
\[
  \Env(\|\Gamma\!\!\downarrow_i\|)
\]
to a distribution over values of the committed type.  The projection from the
full erased environment to this view is the only place where hidden state is
discarded.  Once the strategy has this type, accidental dependence on another
player's secret is impossible.

\paragraph{Why KernelGame is still useful.}
The kernel game is not a competing denotation to the sequential game.  It is
the strategic-form view of the same protocol: choose one strategy per player,
run the program, obtain a distribution over outcomes.  This is exactly the
shape needed by Nash, dominance, welfare, correlated-equilibrium, and
potential-game definitions.  Sequential structure matters for theorems such
as Kuhn; strategic form matters for solution concepts stated as deviations
from complete strategies.

\subsection{Pure Strategic Form}

The repository also constructs a pure-strategy kernel game:
\[
  \toStrategicKG(g) : \KG.
\]
Its strategy type is the legal pure strategy subtype
$\mathcal{P}^{\Legal}_i(g)$.  Pure strategies can be embedded into behavioral
strategies by turning each chosen action into a point mass.  The bridge proves
that the behavioral and pure strategic views agree on the outcome kernel after
this embedding.

This pure strategic form is the right source for mixed-over-pure statements:
a mixed strategy in the Kuhn theorem is a PMF over legal pure strategies.

\paragraph{Mixed strategies.}
The repository follows the standard distinction:
\begin{itemize}[nosep]
\item a pure strategy chooses one action at each of a player's decision sites;
\item a behavioral strategy gives an action distribution at each decision
      site or information state;
\item a mixed strategy is a distribution over pure strategies.
\end{itemize}
Kuhn-style theorems compare the last two.  In VegasCore the mixed side is
represented as an independent product of PMFs over legal pure strategies, one
PMF per player.  The behavioral witness produced by the Kuhn theorem is
PMF-valued because the realization probabilities are real-valued
probabilities produced by the generic GameTheory theorem, not necessarily
representable as rational FDist weights.

\subsection{Equilibrium and Game-Property Wrappers}

The files \texttt{Vegas.Equilibrium}, \texttt{Vegas.PureStrategic}, and
\texttt{Vegas.GameProperties} expose GameTheory concepts directly on
\WFP{}:
\begin{itemize}[nosep]
\item Nash, strict Nash, best response, dominance, and Pareto efficiency;
\item correlated and coarse correlated equilibrium;
\item pure Nash and pure dominance variants;
\item approximate Nash, rationalizability, potential games, zero-sum and
      constant-sum games, security levels, and welfare quantities.
\end{itemize}

These definitions are intentionally thin.  They should not introduce a new
game theory; they give Vegas names to the corresponding GameTheory notions
after applying the appropriate kernel-game bridge.  The corollary files then
transport general GameTheory results back to Vegas.  For example, the fact
that a profile of dominant strategies is Nash is not reproved for Vegas from
scratch; it is inherited through the kernel-game representation.

\subsection{Protocol-Native Concepts}

The strategic wrappers are convenient, but they phrase theorems through an
intermediate kernel game.  \texttt{Vegas.ProtocolSemantics} states the same
core solution concepts directly over the Vegas outcome semantics:
\[
  \protocolEU(g,\sigma,i)
    = \sum_o \Pr_{\sigma,g}[o]\cdot o(i),
\]
and then defines protocol Nash, best response, dominance, and strict Nash in
the expected way.

The correspondence theorems show that these protocol-native definitions agree
with the strategic wrappers.  This is useful for readers: the kernel-game
layer is a representation, not a change in meaning.  A Vegas Nash equilibrium
is exactly the familiar no-profitable-legal-deviation condition stated over
the program's own outcome distribution.


%% =====================================================================
\section{Sequential Denotation and Kuhn}\label{sec:fosg}
%% =====================================================================

\subsection{Why a Sequential Denotation Exists}

The kernel-game views collapse a program to strategic form.  That is the
right representation for many solution concepts, but it hides the sequential
structure of execution.  To use sequential-game theorems from GameTheory, the
repository also gives checked Vegas programs a canonical sequential
denotation.

The chosen target is a factored-observation stochastic game (FOSG).  This is a
good match for Vegas because Vegas observations are factored by construction:
the public state is visible to everyone, and each hidden value is visible only
to its owner until reveal.  A player's observation at a point in the run is
the sequential record of what that player can see.

\subsection{What the FOSG Layer Proves for Vegas}

The public entry point is \texttt{Vegas.FOSG}.  Internally it constructs a
finite cursor-indexed state machine for a checked program.  The cursor records
where the sequential execution is in the program; the state carries the typed
environment; available actions are exactly the program actions at the current
commit site; terminal states correspond to program returns.

The important Vegas-facing facts are:
\begin{itemize}[nosep]
\item the sequential denotation is legal-observable: available moves depend
      only on the acting player's observation;
\item it has bounded horizon, bounded by the remaining program syntax;
\item its run distribution, projected to Vegas payoff outcomes, preserves the
      Vegas outcome distribution.
\end{itemize}

The proof infrastructure behind these facts is substantial, because Lean must
relate cursor states, program suffixes, typed environments, and finite FOSG
histories.  But those are implementation devices.  The semantic statement is
simple: running the sequential denotation of a Vegas program gives the same
distribution over final payoff outcomes as running the Vegas program itself.

\subsection{The Internal Shape of the Sequential Machine}

The FOSG implementation uses two related representations of runtime state.
The first is suffix-oriented: a state contains the current program suffix and
the current typed environment.  This is convenient for proofs by cases on the
next program construct.  The second is cursor-oriented: a finite cursor points
into the original program, and the runtime state carries enough data to
recover the corresponding suffix state.  This is convenient for constructing
a finite FOSG world type.

Both representations are necessary for clean code.  If everything were done
with suffixes, finiteness would be hard to expose to GameTheory.  If
everything were done with cursors, every local proof about the next program
construct would be polluted by cursor decoding.  The bridge between the two
is internal; the user-facing theorem only sees the checked program and the
sequential strategy profile.

\paragraph{Action alphabet.}
There is a similar split for actions.  The evaluator can talk about a broad
action alphabet, essentially a value of the type being committed.  The FOSG
needs a finite action type at each state, so it uses a program-action type
indexed by commit sites in the program.  The bridge theorem says that these
two views agree at the current commit.  This avoids assuming that the entire
expression language has globally finite type codes; finiteness is required
only through the explicit finite-valuation witness used by the theorem.

\paragraph{Observations.}
The private observation recorded for a player contains the current sequential
position and the player's visible environment.  The position matters: the same
visible values at two different commit sites need not have the same available
actions.  The visible environment matters for exactly the reason Vegas exists:
strategies should depend on what the player can observe, not on the full
state.  Legal observability then states that if two histories give a player
the same observation, the set of available legal moves is the same.

\subsection{How the Kuhn Proof Is Routed}

The proved Vegas theorem is not a bespoke probabilistic proof about the
syntax of Vegas.  The route is:
\begin{enumerate}[nosep]
\item compile the checked program to its FOSG sequential denotation;
\item prove the FOSG hypotheses needed by GameTheory: legal observability and
      bounded horizon;
\item map independent mixed Vegas pure strategies to the corresponding
      mixed pure strategies in the FOSG;
\item apply the generic GameTheory FOSG mixed-to-behavioral theorem;
\item project the FOSG run distribution back to Vegas payoff outcomes;
\item use outcome preservation to identify the result with the original Vegas
      mixed-pure outcome distribution.
\end{enumerate}

This division is important for maintainability.  The hard probabilistic
Kuhn argument belongs in GameTheory.  Vegas contributes exactly the facts that
its programming language must prove: how to run the program as a sequential
game, what observations and legal actions are, why the run is finite, and why
the projected outcome distribution is the Vegas one.

\subsection{The Vegas Kuhn Theorem}

The main public sequential-game result is stated back in the native Vegas
strategy space:
\[
\begin{array}{l}
\text{for finite } P \;([{\rm Fintype}\;P])
  \text{ and finite value types }(\mathrm{FiniteValuation}\;L),\\
\text{for every } g:\WFP \text{ and every independent PMF mixture }
  \mu_i \text{ over legal pure strategies,}\\
\text{there exists a total guard-legal Vegas behavioral profile } \beta\\
\text{such that for every } o\in\Outcome,\quad
  \Pr_{\beta}^{\mathrm{beh}}[o]
  =
  \Pr_{\pi\sim\prod_i\mu_i}^{\mathrm{pure}}[o].
\end{array}
\]
In the code this is the Vegas-facing theorem
\texttt{kuhn\_mixedPure\_realizedByBehavioralPMF\_finite}.

The theorem preserves the distribution over payoff outcomes.  Expected
utility preservation is a corollary, not the primitive statement.  This is the
right level: outcome-distribution equality is stronger and composes better
with later utility functions or welfare functionals.

\subsection{How the Sequential Witness Becomes a Vegas Witness}

The generic theorem imported from GameTheory produces a behavioral profile
for the FOSG denotation, indexed by each player's sequential information
state.  Vegas then collapses that witness back to the syntax-recursive
behavioral strategy space:
\[
  \beta : \leanref{LegalProgramBehavioralProfilePMF}\;g.
\]
The transport is not merely a type cast.  The two strategy spaces index
decisions differently:
\begin{itemize}[nosep]
\item the syntax-recursive Vegas behavioral strategy is indexed by commit
      sites and current player-view environments;
\item the sequential behavioral strategy is indexed by player information
      states in the sequential denotation.
\end{itemize}
For well-formed SSA programs these carry the same game-relevant information
along reachable executions.  The proof extracts the active commit site and
current player-view environment from the last sequential observation, proves
that available moves agree at that site, and then runs a step-by-step
outcome-kernel induction.  Off-path Vegas views are completed from an
arbitrary legal pure profile selected from the input mixed profile's support;
this completion is guard-legal and does not affect the outcome distribution.

The repository still exposes the sequential witness type
\[
  \leanref{SequentialBehavioralProfilePMF}\;g
\]
for users who want to inspect or reuse the FOSG denotation directly.  The
unqualified Vegas Kuhn theorem, however, returns the native
\leanref{LegalProgramBehavioralProfilePMF} witness.

Thus the present state is:
\begin{itemize}[nosep]
\item proved: mixed legal pure strategies are realized by a native total
      Vegas PMF behavioral profile with the same outcome distribution
      (\leanref{LegalProgramBehavioralProfilePMF});
\item also exposed: the corresponding sequential-denotation witness, useful
      when working directly with the FOSG semantics.
\end{itemize}

The sequential theorem is still a genuine Vegas theorem: it uses the canonical
sequential denotation and preserves the Vegas outcome distribution.  The
syntax-recursive witness is the main user-facing theorem because it is phrased
entirely in the ordinary Vegas behavioral profile type.


%% =====================================================================
\section{MAID Backend}\label{sec:maid}
%% =====================================================================

\subsection{What the MAID Backend Is For}

The repository also contains a MAID backend.  MAIDs are graphical models with
chance, decision, and utility nodes.  The backend compiles Vegas programs to a
restricted MAID structure and proves that MAID evaluation preserves the Vegas
outcome distribution under the corresponding policy.

Historically, this backend was the main route to sequential-game theorems.
The current public Kuhn theorem instead goes through the FOSG denotation,
because FOSG is the cleaner semantic match for guarded sequential execution.
The MAID backend remains useful:
\begin{itemize}[nosep]
\item it gives an independent graphical-game representation of Vegas
      programs;
\item it validates the relationship between the language's visibility
      discipline and graphical observation parents;
\item it provides a computable compilation route under finite valuation
      assumptions;
\item it contains additional metatheory about perfect recall for the compiled
      graphical structure.
\end{itemize}

\subsection{Compilation Shape}

The MAID compilation maps program constructs as follows:
\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}lll@{}}
\toprule
Vegas construct & MAID node & Role \\
\midrule
\(\Sample\;x\sim D\) & chance node & public random value \\
\(\Commit\;x\;\mathbf{by}\;i\) & decision node & player \(i\)'s action \\
\(\Ret\) & utility nodes & one payoff node per player \\
\(\Let\) & no node & deterministic expression inlined through dependencies \\
\(\Reveal\) & no node & changes visibility/observation structure \\
\bottomrule
\end{tabular}
\end{center}

The significant invariant is that a compiled decision node's observation
parents match the values visible to the player at the corresponding commit
site.  View scoping is what makes this sound: a commit guard cannot depend on
another player's hidden state.

\paragraph{Perfect recall in the compiled MAID.}
The backend proves that the compiled MAID has perfect recall.  Intuitively,
Vegas views are monotone: as the program executes, public values remain
public and a player's own hidden values remain in that player's view.  A
player does not forget previously visible information.  The MAID compilation
records the whole current view as the observation parent set of each decision
node, so later decisions by the same player observe at least what earlier
decisions observed.

\paragraph{Outcome bridge.}
The correctness proof relates two evaluations.  On one side is MAID frontier
evaluation under a policy.  On the other side is the Vegas outcome semantics
under the reflected behavioral profile.  The proof reconstructs typed Vegas
environments from raw MAID node assignments and shows that expression
evaluation agrees on the variables each expression depends on.  This is one
of the places where the expression language's dependency-soundness laws are
load-bearing.

\subsection{Guard Caveat}

MAIDs in the current GameTheory library do not natively carry the same
guard-legal action subtype that Vegas exposes to users.  A compiled MAID
decision node has an action type determined by the value type, whereas a
Vegas commit site may restrict available actions by a state-dependent guard.
This means the MAID backend is outcome-preserving under the compiled policy
correspondence, but it is not the cleanest public home for the final guarded
Vegas Kuhn theorem.

This is why the report treats FOSG as the canonical sequential denotation and
MAID as a useful backend rather than as the main public game-theoretic
semantics.

\subsection{Relation to the MAID Kuhn Theorem}

The MAID backend contains a raw-program Kuhn theorem through the compiled
MAID.  It is useful evidence that the graphical compilation has the expected
perfect-recall structure, but it is not the final public theorem for checked
Vegas programs.  The current public theorem quantifies over legal pure
strategies of a \WFP{} and produces a sequential behavioral witness through
the FOSG denotation.  This distinction prevents a misleading theorem shape:
the MAID route can talk about the full value type at decision nodes, while
the Vegas API must respect state-dependent commit guards.


%% =====================================================================
\section{Expression Languages and Finiteness}\label{sec:expr}
%% =====================================================================

\subsection{The Expression Interface}

VegasCore is generic over an expression language interface \(L\).  The
interface supplies:
\begin{itemize}[nosep]
\item a universe of expression types, including Booleans and integers;
\item values for each type;
\item expressions and distribution expressions indexed by contexts;
\item evaluators for expressions and distribution expressions;
\item dependency tracking for both expressions and distributions;
\item soundness laws saying that evaluation depends only on tracked
      variables;
\item structural operations needed to extend and drop context entries in
      commutativity proofs.
\end{itemize}

This is why most Vegas metatheory is parametric in \(L\).  The repository
includes a concrete simple expression language with Booleans, integers,
arithmetic, equality, conditionals, and simple finite distributions for
examples and testing.

\subsection{Finite Valuation}

Finiteness is not part of the core expression-language interface.  The
separate finite-valuation assumption says that every value type \(L.\Val(\tau)\)
is finite.  It is needed for:
\begin{itemize}[nosep]
\item enumerating pure strategies;
\item compiling finite action spaces for backend game structures;
\item applying finite FOSG and MAID theorems;
\item stating the current finite Kuhn theorem.
\end{itemize}

The denotational semantics, trace semantics, behavioral strategic bridge, and
many equilibrium wrappers do not need finite valuation.  This separation keeps
the core language more general while making the finite-game results available
when the expression language supports enumeration.


%% =====================================================================
\section{Repository Map}\label{sec:repo}
%% =====================================================================

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}ll@{}}
\toprule
Area & Main files \\
\midrule
Core syntax and expression layer &
  \texttt{Core}, \texttt{ExprSimple}, \texttt{VegasSimple} \\
Well-formedness and checked programs &
  \texttt{WF}, \texttt{Scope}, \texttt{WFProgram} \\
Operational and denotational semantics &
  \texttt{Operational}, \texttt{BigStep}, \texttt{Strategic},
  \texttt{StrategicPMF} \\
Trace and action graph semantics &
  \texttt{TraceSemantics}, \texttt{TraceCorollaries},
  \texttt{ActionGraph} \\
Pure and behavioral strategic views &
  \texttt{PureStrategic}, \texttt{ViewKernel}, \texttt{LegalNonempty} \\
Game-theoretic wrappers &
  \texttt{Equilibrium}, \texttt{GameProperties}, \texttt{Corollaries/*} \\
Protocol and sequential theorems &
  \texttt{ProtocolSemantics}, \texttt{Kuhn}, \texttt{FOSG} \\
FOSG implementation internals &
  \texttt{FOSG/Basic}, \texttt{FOSG/Observation},
  \texttt{FOSG/Observed/*} \\
MAID backend &
  \texttt{MAID/*} \\
Examples &
  \texttt{Examples} \\
\bottomrule
\end{tabular}
\end{center}

The public import file \texttt{Vegas.lean} gathers the supported Vegas
surface.  The \texttt{Explorations/} directory is not part of the main
surface; it preserves earlier design experiments.

\subsection{Suggested Reading Order}

For understanding the current repository, the most efficient path is:
\begin{enumerate}[nosep]
\item \texttt{WFProgram}: the boundary between syntax and games.
\item \texttt{Strategic} and \texttt{PureStrategic}: legal strategy spaces and
      kernel-game construction.
\item \texttt{Equilibrium} and \texttt{GameProperties}: the public
      strategic-form vocabulary.
\item \texttt{ProtocolSemantics}: why the strategic wrappers agree with
      protocol-native statements.
\item \texttt{Kuhn} and \texttt{FOSG}: the public sequential theorem and its
      denotational entry point.
\item \texttt{FOSG/Observed/Kernel}: the deepest part of the FOSG proof route,
      only after the public statement is clear.
\item \texttt{MAID}: the graphical backend and its separate theorem route.
\end{enumerate}

Reading the FOSG internals first is usually counterproductive.  Much of that
code is there to satisfy Lean's dependent typechecker and GameTheory's finite
interfaces.  The public semantic shape is clearer from \texttt{Kuhn},
\texttt{FOSG}, and \texttt{ProtocolSemantics}.


%% =====================================================================
\section{Design Lessons}\label{sec:lessons}
%% =====================================================================

\paragraph{The checked-program boundary is essential.}
It is tempting in a mechanized development to expose the raw inductive syntax
everywhere and ask each theorem to carry the needed side conditions.  That
works internally, but it gives users too many ways to construct
game-theoretic objects from incoherent syntax.  The \WFP{} bundle makes the
safe boundary explicit.

\paragraph{Legality must appear in the strategy type.}
Program legality only prevents deadlock.  It does not prevent an arbitrary
strategy from recommending illegal actions.  Moving guard admissibility into
the public strategy subtypes makes the equilibrium API match the intended
protocol semantics.

\paragraph{FOSG is the sequential denotation; KernelGame is the strategic
view.}
These are not competing backends.  A sequential game and its strategic form
answer different questions.  FOSG keeps the sequential information structure
visible, which is why it is the natural route to Kuhn.  KernelGame packages
the induced strategic interaction, which is why it is the natural route to
Nash, dominance, welfare, and other strategic-form concepts.

\paragraph{Outcome preservation is the right invariant.}
Expected utility equalities are useful, but they are consequences of a
stronger statement: equality of outcome distributions.  The current
FOSG-based Kuhn theorem is stated at the distribution level, which makes it
robust under changes to utility interpretation and welfare aggregation.

\paragraph{MAID remains valuable but is no longer the main theorem path.}
The MAID backend explains a lot about the graphical structure of Vegas games,
and it supplies a useful compilation/evaluation route.  But state-dependent
guards make FOSG the cleaner target for the public guarded sequential theorem.

\paragraph{Dependent transport is the main engineering cost.}
Most difficult proofs are not mathematically surprising.  They are about
showing that two dependently typed views of the same program state agree:
cursor states versus suffix states, erased environments versus typed
environments, reachable histories versus current observations, and FDist
versus PMF presentations.  The stable architecture is the one that keeps
those transports internal and exposes simple semantic statements at the
boundary.


%% =====================================================================
\section{Current Status and Open Ends}\label{sec:status}
%% =====================================================================

\subsection{Stable Public Results}

The repository currently has:
\begin{itemize}[nosep]
\item checked-program game APIs based on \WFP{};
\item legal behavioral and pure strategy spaces;
\item behavioral and pure kernel-game views preserving the Vegas outcome
      semantics;
\item Vegas-facing equilibrium and game-property wrappers;
\item protocol-native definitions corresponding to the kernel-game wrappers;
\item FOSG sequential denotation preserving outcome distributions;
\item finite mixed-to-behavioral realization preserving outcome
      distributions, with a native Vegas behavioral witness;
\item MAID compilation and associated outcome-preservation/perfect-recall
      metatheory.
\end{itemize}

\subsection{Known Limits}

\begin{itemize}[nosep]
\item The finite Kuhn theorem requires finite players and finite value types.
\item The MAID backend does not by itself expose the same guard-legal action
      subtype as the public Vegas strategy API.
\item The concrete executable path is better through finite backend
      compilation than through reducing the abstract FDist denotational
      semantics in the Lean kernel.
\end{itemize}

\subsection{Natural Future Work}

\begin{itemize}[nosep]
\item Prove a clean correspondence between syntax-recursive Vegas behavioral
      profiles and sequential behavioral profiles, if the syntax-recursive
      strategy space is needed as the witness type for Kuhn-style theorems.
\item Strengthen or adapt generic GameTheory backends so state-dependent
      action availability is represented uniformly across sequential and
      graphical games.
\item Add more expression language instances and surface-language tooling.
\item Extend the game-theoretic theorem surface beyond Kuhn and strategic-form
      wrappers, using FOSG as the canonical sequential denotation.
\end{itemize}


%% =====================================================================
\section{Related Work}\label{sec:related}
%% =====================================================================

\paragraph{Game description languages.}
GDL, Gambit, and OpenSpiel provide ways to describe and solve games.  Their
focus is execution and solving, whereas VegasCore focuses on a mechanized
semantics with a type-directed account of information.

\paragraph{Mechanized game theory.}
The GameTheory submodule supplies the generic library of kernel games, MAIDs,
FOSGs, extensive-form games, and theorem infrastructure.  VegasCore's
contribution is to connect a programming language for games to those
objects with verified semantic bridges.

\paragraph{Information-flow type systems.}
Visibility annotations resemble security labels, but the purpose differs.
Security type systems prevent unintended leakage.  Vegas uses visibility to
state what information a player has when choosing an action, so the same style
of structure enforces game-theoretic information constraints.

\paragraph{Commit-reveal protocols.}
Blockchain commit-reveal protocols motivate the language design, but Vegas is
not a cryptographic model.  Commit and reveal are semantic constructs for
information timing.  Cryptographic enforcement is outside the calculus.


%% =====================================================================
\section{Conclusion}\label{sec:conclusion}
%% =====================================================================

VegasCore now has a coherent end-to-end story.  A checked Vegas program is a
game.  Its strategic form is available through guard-legal kernel-game
wrappers.  Its protocol-level solution concepts agree with those wrappers.
Its sequential denotation is a FOSG that preserves the Vegas outcome
distribution.  For finite games, independent mixtures over legal pure
strategies can be realized by sequential behavioral strategies with the same
outcome distribution.  The MAID backend remains as a graphical compilation
route and an additional body of metatheory, but the canonical sequential
denotation for current public theorem statements is FOSG.

The important design boundary is that public theorem statements should be
simple and Vegas-facing, while proof machinery stays behind the denotational
interfaces.  That boundary is what makes the current repository usable despite
the large amount of mechanized infrastructure underneath.

\begin{thebibliography}{9}

\bibitem{koller2003multi}
D.~Koller and B.~Milch.
\newblock Multi-agent influence diagrams for representing and solving games.
\newblock {\em Games and Economic Behavior}, 45(1):181--221, 2003.

\bibitem{kovarik2019rethinking}
V.~Kova\v{r}\'ik, M.~Schmid, N.~Burch, M.~Bowling, and V.~Lis\'y.
\newblock Rethinking formal models of partially observable multiagent decision
  making.
\newblock {\em Artificial Intelligence}, 277:103171, 2019.

\end{thebibliography}

\end{document}
