\documentclass[11pt]{article}

\input{macros}

\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,calc,shapes.geometric,decorations.pathreplacing}

% Be more tolerant about line breaking, since identifier-heavy paragraphs
% can produce unbreakable tokens longer than a column.
\sloppy
\emergencystretch=3em
\hbadness=10000
\hfuzz=4pt

% Local additions for the supplement.
\newcommand{\TT}[1]{\texttt{\detokenize{#1}}}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{TODO:} #1}}
\newcommand{\DefRef}[1]{\hyperref[def:#1]{\TT{#1}}}

\title{VegasCore: A Definitional Supplement\\
\large{A Cross-Discipline Reading Guide for the Lean Mechanization}}
\author{}
\date{}

\begin{document}
\maketitle

\begin{abstract}
This document is a companion to the technical report \emph{VegasCore:
Event-Graph-Native Protocol Semantics}. The report describes the architecture in
the language of types and theorems; this supplement collects the
\emph{definitions with mathematical content} that a careful reader needs in
order to navigate the Lean codebase. Each entry pairs the Lean form with an
ordinary mathematical equivalent and a short prose explanation of what the
construct is for.

The intended audience is twofold. We assume programming-languages background
(typed contexts, monads, dependent types, well-formedness judgements) and
provide informal definitions of the game-theoretic vocabulary; conversely we
assume a reader familiar with Nash equilibria, behavioural strategies, and
finite extensive-form games but new to the Lean formalisation, and explain the
PL machinery in terms they can map to standard game forms.

The supplement is organised bottom-up along the same path as the main report:
the ambient probabilistic and language layers, then visibility-aware syntax,
then well-formedness, then the event graph and its machine semantics,
then FOSG views, kernel games, and the public solution-concept API. It
finishes with a tour of the worked examples.
\end{abstract}

\tableofcontents

% =====================================================================
\section{How To Read This Document}

\subsection{Conventions}
For each definition we present three items: (i) the Lean signature, often
shortened to the names and types that carry mathematical content, with
implementation noise stripped; (ii) a mathematical equivalent in standard
notation; (iii) a short prose paragraph explaining the construct and
\emph{why it exists}. The Lean name is given in monospace
(e.g.~\TT{Vegas.WFProgram}\label{def:Vegas.WFProgram}). Cross-references match the names used in the
main report.

We use the macros from the main paper: $\WFP$, $\Outcome$, $\FWeight{\cdot}$,
$\PMF{\cdot}$, $\bind$, $\Players$, $\view{\cdot}$, $\pubctx{\cdot}$,
$\erasectx{\cdot}$, $\erasepub{\cdot}$, $\Reach$, $\CanReach$. Where the
supplement introduces additional notation we mark it locally.

\subsection{Bird's-eye view}
\label{sec:birdseye}
A Vegas program is a syntactic description of a finite game. Its meaning is
obtained through the layers below, each of which contributes its own
vocabulary. We list them in the order definitions will appear.

\begin{center}
\begin{tikzpicture}[
  every node/.style = {draw, rounded corners, align=center, font=\small,
    minimum width=2.7cm, minimum height=1cm},
  arr/.style = {-{Stealth[length=2mm,width=2mm]}, thick},
]
\node (lang)  {\textbf{Language}\\\TT{IExpr}, \TT{FWeight}};
\node[right=4mm of lang] (vis) {\textbf{Visibility}\\\TT{VCtx}, \TT{viewVCtx}};
\node[right=4mm of vis] (syn) {\textbf{Source}\\\TT{VegasLang},\\\TT{VegasCore},\\\TT{WFProgram}};
\node[below=10mm of syn] (graph) {\textbf{Event graph}\\\TT{EventGraph}\\\TT{Configuration}};
\node[left=4mm of graph] (mach) {\textbf{Machine}\\\TT{Machine},\\events};
\node[left=4mm of mach] (fosg) {\textbf{FOSG view}\\\TT{Machine.FOSGView}};
\node[below=10mm of fosg] (kg) {\textbf{Kernel game}\\\TT{KernelGame}\\(public API)};
\draw[arr] (lang) -- (vis);
\draw[arr] (vis) -- (syn);
\draw[arr] (syn) -- (graph);
\draw[arr] (graph) -- (mach);
\draw[arr] (mach) -- (fosg);
\draw[arr] (fosg) -- (kg);
\end{tikzpicture}
\end{center}

\noindent The arrows are read as ``defines on top of''. The substantive
mathematical work happens at the event-graph level, where a syntax tree
is turned into a DAG with \emph{extensional} state — the configuration
records only which nodes have produced which results, not the schedule
that produced them — and at the FOSG checkpoint level, where the
game-theoretic round structure is recovered.

\subsection{Code organisation}
\label{sec:code-organisation}\label{def:Lang.ToCore}\label{def:Core.ToEventGraph}\label{def:EventGraph.ToMachine}\label{def:Fields}\label{def:Obligations}\label{def:Construction}

The Lean modules follow the semantic ownership of the construction, not a
single compiler package.  The internal Vegas pipeline uses source-owned
\TT{ToX}\label{def:ToX} names:
\[
  \TT{Lang.ToCore}
  \;\longrightarrow\;
  \TT{Core.ToEventGraph}
  \;\longrightarrow\;
  \TT{EventGraph.ToMachine}.
\]
These are the definitional arrows of Vegas itself.  In contrast,
\TT{GameBridge/} contains optional bridges from the Vegas semantics to
external game-form infrastructure, such as FOSG and EFG.  Operational
implementation refinement lives in \TT{Backend/}; it is not a game-form
bridge.  Native Vegas strategic semantics live in \TT{Strategic/}; removing
the FOSG bridge would not remove the fact that checked Vegas programs have
strategy carriers, profiles, outcome kernels, and equilibrium predicates.

\begin{center}\small
\begin{tabular}{@{}p{0.25\textwidth}p{0.66\textwidth}@{}}
\toprule
Folder & Role \\
\midrule
\TT{Base/}
  & Common finite-weight, expression, context, visibility, environment, and
    payoff infrastructure shared by core syntax and event graphs. \\
\TT{Lang/}
  & Concrete/surface language layers, including \TT{VegasLang} and its
    lowering to \TT{VegasCore}. \\
\TT{Core/}
  & \TT{VegasCore}, static well-formedness, checked programs, view kernels,
    and the elaboration \TT{Core.ToEventGraph}.  The elaboration is split into
    \TT{ToEventGraph/Nodes}, \TT{Fields}, \TT{Obligations}, and
    \TT{Construction}. \\
\TT{EventGraph/}
  & The central event-graph object, configurations, frontier facts, and the
    internal interpretation \TT{EventGraph.ToMachine}. \\
\TT{Machine/}
  & Generic operational machine carrier and primitive trace semantics. \\
\TT{Strategic/}
  & Vegas-native strategic semantics: strategy carriers, kernel games,
    expected utility, equilibrium vocabulary, dominance, and transport. \\
\TT{GameBridge/}
  & Bridges from Vegas semantics to FOSG/EFG machinery and the FOSG-based Kuhn
    realisation bridge. \\
\TT{Backend/}
  & Runtime/backend refinement obligations and backend blocked-trace
    kernel-game transport. \\
\bottomrule
\end{tabular}
\end{center}

\subsection{A glossary for the cross-disciplinary reader}

\paragraph{For PL readers.}
A \emph{(finite) game} consists of a player set $\Players$, per-player action
sets, and a probabilistic mechanism that turns chosen profiles into payoff
vectors. A \emph{strategy} is what a player intends to do; a \emph{profile} is
one strategy per player. \emph{Expected utility} $\EU$ is the expected payoff
under a profile. \emph{Nash equilibrium} is a profile from which no player
gains by unilateral deviation. \emph{Mixed} strategies randomise over pure
ones; \emph{behavioural} strategies randomise locally at each information set.
\emph{Kuhn's theorem} says that under perfect recall mixed and behavioural
strategies are equivalent. A \emph{factored-observation stochastic game} (FOSG)
is the modern formalism for finite extensive-form games with imperfect
information; players observe local signals at each round.

\paragraph{For game-theory readers.}
A \emph{typed context} $\VCtx$ is just a list of $(\text{name}, \text{type})$
pairs, like a typing environment in a static type checker. A
\emph{de~Bruijn-like} membership proof \TT{HasVar}\label{def:HasVar}~$\Gamma$~$x$~$\tau$ is the
evidence that variable $x$ has type $\tau$ in $\Gamma$. A
\emph{well-formedness} predicate is a side condition collected by the type
checker (here: bindings have unique names, every commitment is revealed,
distributions sum to one, every commit can be played by at least one
action). \emph{Erasure} forgets the visibility annotation on a context to
hand the result to a plain expression evaluator. The \emph{denotation
brackets} $\den{p}$ are not needed for individual definitions; they are
bookkeeping for ``the meaning of $p$''.

\paragraph{A two-column glossary.}

\begin{center}\small
\begin{tabular}{@{}p{0.32\textwidth}p{0.6\textwidth}@{}}
\toprule
PL term & Game-theoretic counterpart \\
\midrule
typed binding $(x, \pub\,\tau)$
  & a public observation channel of type $\tau$ \\
typed binding $(x, \hid\,p\,\tau)$
  & a sealed move/private signal of player $p$ \\
\Sample~$x \sim D$
  & a chance node drawn from $D$, observable by all \\
\Commit{} (player $p$) $x \mid R$
  & a player-$p$ move at an information set; $R$ is the legality predicate \\
\Reveal~$y := x$
  & a public announcement of a previously sealed value \\
\Ret~$\bar{p}$
  & terminal node carrying the payoff vector $\bar p$ \\
\midrule
\TT{EventGraph}
  & the game tree, but DAG-shaped and order-quotiented \\
\TT{Configuration}
  & a partially-played history (closed under causal deps) \\
frontier of a configuration
  & set of currently-active moves \\
\TT{Machine}
  & operational carrier of the game \\
\TT{Machine.FOSGView}
  & the same game read at round-level, with information sets \\
\TT{KernelGame}
  & strategic-form game (strategies, outcome kernel, utility) \\
\TT{pureKernelGame}, \TT{pmfBehavioralKernelGame}
  & finite normal-form / behavioural-form view of a Vegas program \\
\bottomrule
\end{tabular}
\end{center}

% =====================================================================
\section{The Probability and Language Layer}

\subsection{Finite-support weights: \TT{FWeight}}
\label{def:FWeight}

\paragraph{Lean.}
\begin{quote}
\TT{abbrev FWeight (}$\alpha$\TT{) [DecidableEq }$\alpha$\TT{] := }$\alpha \to_0 \mathbb{Q}_{\ge 0}$
\end{quote}
together with the operations \TT{pure}\label{def:pure}, \TT{zero}\label{def:zero}, \TT{bind}\label{def:bind}, \TT{map}\label{def:map},
\TT{ofList}\label{def:ofList}, \TT{totalWeight}\label{def:totalWeight}, and the bridge \TT{toPMF}\label{def:toPMF} into Mathlib's
measure-theoretic $\PMF{\cdot}$.

\paragraph{Math.}
$\FWeight{\alpha}$ is the set of finitely-supported, non-negative-rational
weight functions on $\alpha$:
\[
  \FWeight{\alpha} \;=\; \bigl\{ \mu : \alpha \to \NNQ \;\bigm|\; |\mathrm{supp}\,\mu| < \infty \bigr\}.
\]
The total mass $w(\mu) = \sum_{a} \mu(a)$ ranges over $[0, \infty)$;
normalised weights, written \TT{FProb}\label{def:FProb} in Lean, are those with $w(\mu)=1$.
The standard monad operations are
\[
  \mathrm{pure}(x)(a) = [a = x],\quad
  \mu \bind f = \sum_{a} \mu(a)\cdot f(a),\quad
  \mathrm{map}\,g\,\mu = \sum_{a} \mu(a)\cdot \mathrm{pure}(g(a)).
\]
(The trio $\mathrm{pure}$, $\bind$, $\mathrm{map}$ is the standard PL
vocabulary for a structured computational effect: $\mathrm{pure}$ injects a
value as a degenerate distribution, $\bind$ sequences a stochastic step
with a continuation indexed by the result, and $\mathrm{map}$ is the
pushforward through a deterministic function.)

\paragraph{Why this and not $\PMF{\cdot}$.}
Three reasons: (i) raw weights allow intermediate \texttt{bind}-trees to be
built without per-step normalisation proofs; (ii) rational weights make
finite weights decidably equal and concrete examples mechanically
inspectable; (iii) decidable support is necessary to construct
\texttt{Finsupp}-backed weights computably. The conversion \TT{toPMF}
requires a proof of total weight $1$, is one-way, and is the single exit
point into Mathlib's measure theory.

\paragraph{Key laws.}
The \emph{monad laws} (left/right identity for $\mathrm{pure}$, and
associativity of $\bind$) hold; in particular finite weights \emph{commute}
($\mu_1 \bind \lambda x.\, \mu_2 \bind \lambda y.\, f(x,y) = \mu_2 \bind \lambda y.\, \mu_1 \bind \lambda x.\, f(x,y)$),
which is what licenses commit-commit commutation in the program semantics.
Normalisation propagates: if $w(\mu)=1$ and $w(f(a))=1$ for every $a$ in
$\mathrm{supp}(\mu)$, then $w(\mu \bind f) = 1$.

\subsection{The expression-language interface: \TT{IExpr}}
\label{def:IExpr}

\paragraph{Lean.}
\TT{IExpr} is a record with the following fields, grouped by role:
\begin{itemize}\itemsep1pt
\item \emph{Types and values}: $\Ty : \mathbf{Type}$;
  $\Val : \Ty \to \mathbf{Type}$; \TT{decEqTy : DecidableEq Ty};
  \TT{decEqVal :} $\forall \tau,$ \TT{DecidableEq (Val}~$\tau$\TT{)};
  distinguished $\bool, \intt \in \Ty$ with projections to Lean's
  \TT{Bool}\label{def:Bool} and \TT{Int}\label{def:Int}.
\item \emph{Expression syntax}: \TT{Expr}\label{def:Expr}~$\Gamma$~$\tau$; an evaluator
  \TT{eval}\label{def:eval}; a static dependency over-approximation \TT{exprDeps}\label{def:exprDeps}; and
  \TT{expr_deps_context}\label{def:expr_deps_context}, which proves dependencies mention only variables
  in the expression context.
\item \emph{Distribution syntax}: \TT{DistExpr}\label{def:DistExpr}~$\Gamma$~$\tau$;
  \TT{evalDist}\label{def:evalDist} into $\FWeight{\Val(\tau)}$; \TT{distDeps}\label{def:distDeps}.
\item \emph{Structural lemmas}: \TT{extendAfterHead}\label{def:extendAfterHead} (insert an unrelated
  binding after the head) and \TT{dropAfterHead}\label{def:dropAfterHead} (remove a head-after binding
  the expression does not depend on).
\item \emph{Soundness laws}: the dependency sets faithfully predict the
  read sets:
\[
  \mathrm{AgreesOn}(\rho_1, \rho_2, \mathtt{exprDeps}\,e) \;\implies\;
  \mathtt{eval}\,e\,\rho_1 = \mathtt{eval}\,e\,\rho_2,
\]
and the analogous law for distributions.
\end{itemize}

\paragraph{Math.}
This is the abstract interface of a typed expression language with finite
weights. The \emph{soundness} laws are the only nontrivial axioms; they
justify treating syntactic dependency as a sound visibility analysis.

\paragraph{Trusted interface obligation.}
\TT{IExpr} is an abstraction boundary. The Vegas proofs do not inspect the
internal syntax of an expression language; they rely on each \TT{IExpr}
instance to prove that \TT{exprDeps} and \TT{distDeps} are sound
over-approximations of semantic reads. An instance may over-report
dependencies, but it cannot under-report a semantically relevant read and
still satisfy \TT{expr_deps_sound}\label{def:expr_deps_sound} or \TT{dist_deps_sound}\label{def:dist_deps_sound}.

\paragraph{Concrete instance.}
\TT{simpleExpr}\label{def:simpleExpr} (in \TT{Vegas/Lang/ExprSimple.lean}) instantiates \TT{IExpr} with
$\Ty = \{\intt, \bool, \mathsf{range}\,\ell\,h,
\mathsf{option}\,b\}$ and the obvious values, where
$\Val(\mathsf{option}\,b)=\mathrm{Option}(\Val(b))$. Its expression syntax
has constants, addition, equality, boolean conjunction and negation,
\texttt{ite}, and the option constructors/eliminators
\TT{none}\label{def:none}, \TT{some}\label{def:some}, \TT{isSome}\label{def:isSome}, \TT{isNone}\label{def:isNone}, and \TT{getD}\label{def:getD}, plus
distributions built either as a weighted list of $(v, q)$ pairs or by guarded
choice. The \TT{VegasLang} programs in this development use \TT{simpleExpr}.
The \TT{VegasLang} nullable-yield layer is intentionally concrete over this
instance; another expression language would need its own option type,
non-nullable commit-payload predicate, and default-value operation.

Two small type classes are part of the concrete instance rather than the
generic \TT{IExpr} interface. \TT{CommitPayloadTy}\label{def:CommitPayloadTy} marks the non-nullable
types that may be written as explicit surface commit payloads; in particular
\TT{not_commitPayloadTy_option}\label{def:not_commitPayloadTy_option} rules out explicit user commits of
\TT{BaseTy.option b}. \TT{DefaultVal}\label{def:DefaultVal} supplies the fallback needed to embed a
guard over $b$ into a guard over \TT{option b}; the fallback is used only under
\TT{getD} in the non-null branch.

\subsection{Plain typed contexts: \TT{Ctx}, \TT{HasVar}, \TT{Env}}
\label{def:Ctx}

\paragraph{Lean.}
\begin{quote}
\TT{abbrev Ctx (Ty : Type) := List (VarId} $\times$ \TT{Ty)} \\
\TT{inductive HasVar : Ctx Ty} $\to$ \TT{VarId}\label{def:VarId} $\to$ \TT{Ty} $\to$ \TT{Type}\label{def:Type}\\
\TT{def Env (Val : Ty} $\to$ \TT{Type) (}$\Gamma$\TT{ : Ctx Ty) :=} \TT{(}$x$\TT{,}$\tau$\TT{)} $\to$ \TT{HasVar }$\Gamma$\TT{ }$x$\TT{ }$\tau$ $\to$ \TT{Val}\label{def:Val}~$\tau$
\end{quote}

\paragraph{Math.}
A \emph{context} $\Gamma$ is a list of typed variable bindings; an
\emph{environment} $\rho \in \Env(\Gamma)$ assigns a value to each binding.
\TT{HasVar} is a typed de~Bruijn membership proof; under \TT{Nodup}\label{def:Nodup} (no
repeated names) it is a \emph{subsingleton} (any two such proofs are
definitionally equal), so environment lookup is \emph{proof-irrelevant}
(the value depends only on the variable and its type, not on which
membership witness was used).

\paragraph{Why \TT{HasVar} is in \TT{Type} and not \TT{Prop}.}
The \TT{Env}\label{def:Env} datatype pattern-matches on \TT{HasVar} to retrieve values;
this requires \TT{HasVar} to live in the data universe. The single lemma
\TT{Env.get_eq_of_nodup}\label{def:Env.get_eq_of_nodup} discharges the resulting proof-position equality
once and for all.

% =====================================================================
\section{Visibility-aware Syntax}

The protocol layer cares not just about variables but about \emph{who can see
which variable}. The standard PL apparatus is extended with a visibility tag
on each binding.

\subsection{Visibility tags: \TT{BindTy}}
\label{def:BindTy}

\paragraph{Lean.}
\TT{inductive Visibility (Player : Type)} with constructors
\TT{pub}\label{def:pub} and \TT{hidden (owner : Player)}. Then
\TT{structure BindTy (Player : Type) (L : IExpr)} has fields
\TT{base : L.Ty} and \TT{visibility : Visibility Player}. The namespace
\TT{BindTy} provides smart constructors \TT{.pub (}$\tau$\TT{)} and
\TT{.hidden (owner : Player) (}$\tau$\TT{)}.

\paragraph{Math.}
A binding type is either \emph{public} ($\pub\,\tau$, visible to all) or
\emph{hidden} ($\hid\,p\,\tau$, visible only to its owner $p$). The information
model is intentionally \emph{owner-or-all}: there are no group-visibility
constructors. Internally, the base type is an immutable field and visibility is
independent metadata: reveal-like operations change visibility by introducing a
fresh public alias, not by changing the variable's base type.

\subsection{Visibility contexts and views: \TT{VCtx}, \TT{viewVCtx}, \TT{pubVCtx}}
\label{def:VCtx}\label{def:pubVCtx}\label{def:viewVCtx}

\paragraph{Lean.}
\begin{quote}
\TT{abbrev VCtx (P : Type) (L : IExpr) := List (VarId} $\times$ \TT{BindTy P L)} \\
\TT{def viewVCtx (p : P) : VCtx P L} $\to$ \TT{VCtx P L} \quad\textemdash\quad keeps only bindings $p$ can see.\\
\TT{def pubVCtx : VCtx P L} $\to$ \TT{VCtx P L} \quad\textemdash\quad keeps only public bindings.
\end{quote}

\paragraph{Math.}
$\view{p}$ is the sublist of bindings of $\Gamma$ observable by player $p$;
$\pubctx{\Gamma}$ is the sublist of public bindings:
\[
  \view{p} \;=\; \bigl[(x, \tau) \in \Gamma \;\big|\; p \text{ can see } \tau\bigr],
  \qquad
  \pubctx{\Gamma} \;=\; \bigl[(x, \pub\,\tau) \in \Gamma\bigr].
\]
Public bindings are visible to every player, so $\pubctx{\Gamma} \subseteq
\view{p}$ as multisets.

\subsection{Erasure to plain contexts: \TT{eraseVCtx}, \TT{erasePubVCtx}}
\label{def:erase}\label{def:erasePubVCtx}\label{def:eraseVCtx}

\paragraph{Lean.}
\TT{def eraseVCtx} forgets the visibility tag, returning a plain
\TT{Ctx L.Ty}. \TT{def erasePubVCtx} keeps only public variables and
forgets the tag.

\paragraph{Math.}
There are two erasures because there are two consumers:
\[
  \erasectx{\Gamma} = [(x, \tau.\mathrm{base}) \mid (x,\tau) \in \Gamma], \quad
  \erasepub{\Gamma} = [(x, \tau) \mid (x, \pub\,\tau) \in \Gamma].
\]
The key commutation lemma (\TT{eraseVCtx_pubVCtx}\label{def:eraseVCtx_pubVCtx}) says that the two
projections — restrict to the public fragment, and forget visibility
tags — commute:
\[
  \erasectx{\pubctx{\Gamma}} \;=\; \erasepub{\Gamma}.
\]
Either route produces the public erasure, for which we keep the
shorthand $\erasepub{\Gamma}$. Operationally this is what lets
$\erasepub{\Gamma}$-shaped goals be discharged by reasoning about
$\erasectx{\Gamma}$.

\paragraph{Two erased contexts, by design.}
The full erased context $\erasectx{\Gamma}$ is what \emph{commit guards} and
operational machinery type over: a guard is permitted to mention any
variable, but a separate predicate restricts which it may actually depend
on. The public-only erased context $\erasepub{\Gamma}$ is what
\emph{deterministic let-bindings, samples, and payoff expressions} type over,
since these by construction may not read hidden state. This split is the
single design commitment around which the rest of the development is
organised.

\subsection{Source syntax: \TT{VegasCore}}
\label{def:VegasCore}

\paragraph{Lean.}
\TT{inductive VegasCore (P : Type) (L : IExpr) : VCtx P L} $\to$ \TT{Type}
with five constructors. Schematically:
\[
\arraycolsep=1pt
\begin{array}{r@{\;}c@{\;}l}
\Ret~\bar p
  & : & \VCtx \to \mathrm{VegasCore}\,\VCtx \\
\Let~x = e;~k
  & : & \mathsf{Expr}(\erasepub{\Gamma}, b) \times \mathrm{VegasCore}((x,\pub\,b)::\Gamma) \to \mathrm{VegasCore}\,\Gamma \\
\Sample~x \sim D;~k
  & : & \mathsf{DistExpr}(\erasepub{\Gamma}, b) \times \mathrm{VegasCore}((x,\pub\,b)::\Gamma) \to \mathrm{VegasCore}\,\Gamma \\
\Commit_{p}~x \mid R;~k
  & : & \mathsf{Expr}((x,b)::\erasectx{\Gamma}, \bool) \times \mathrm{VegasCore}((x,\hid\,p\,b)::\Gamma) \to \mathrm{VegasCore}\,\Gamma \\
\Reveal~y := x;~k
  & : & \mathrm{VHasVar}\,\Gamma\,x\,(\hid\,p\,b) \times \mathrm{VegasCore}((y,\pub\,b)::\Gamma) \to \mathrm{VegasCore}\,\Gamma \\
\end{array}
\]

\paragraph{Math.}
A program is a typed sequence of \emph{events} terminated by a \Ret{}. Four
constructors model observable activity in a multi-party computation:
\Sample{} (a public coin), \Commit{} (a sealed choice by a player),
\Reveal{} (a sealed value made public), and \Ret{} (the protocol terminates
and pays out). One constructor, \Let, is administrative.

The argument types follow the visibility discipline of the previous
subsection: \Let, \Sample, and the payoffs at \Ret{} type their bodies over
$\erasepub{\Gamma}$ so they cannot mention hidden state, while \Commit's
guard $R$ types over $(x,b)::\erasectx{\Gamma}$ — syntactically it can read
any variable, so a separate predicate \TT{GuardUsesOnly}\label{def:GuardUsesOnly} is needed to enforce
that it only reads what the committing player can see (\S\ref{def:scope}).

\paragraph{Why \Let{} is kept rather than encoded.}
$\Let~x = e;~k$ is semantically equivalent to $\Sample~x \sim \delta_e;~k$
where $\delta_e$ is the Dirac mass at $e$. The codebase keeps \Let{} as a
distinct constructor for four reasons: (i) it would otherwise force \TT{IExpr}
to expose a \TT{Expr} $\to$ \TT{DistExpr} lift; (ii) it keeps program
representation linear in surface size; (iii) it preserves the binding
$(x, \pub\,b)$ in every downstream view, so strategy types do not change; and
(iv) the surface syntax makes randomness visually distinct from bookkeeping.

\subsection{Surface syntax: \TT{VegasLang}}
\label{def:VegasLang}\label{def:simultaneous}

\paragraph{Lean.}
\TT{inductive VegasLang (P : Type) [DecidableEq P] :
VCtx P simpleExpr} $\to$ \TT{Type}
mirrors the five \TT{VegasCore} constructors over \TT{simpleExpr} and adds
simultaneous yield phases plus nullable \TT{yield}\label{def:yield} sugar:
\[
\begin{array}{@{}l@{}}
  \TT{simultaneous} : \\[-1pt]
    \qquad(\phi : \TT{VegasYieldPhase P } \Gamma\,[]\,\Delta)
      \to \phi.\TT{DistinctActors} \to \\[1pt]
    \qquad\TT{VegasLang P } \Delta
      \to \TT{VegasLang P } \Gamma, \\[2pt]
  \Gamma_{\mathsf{yield}} =
    \begin{array}[t]{@{}l@{}}
    ((y,\pub\,(\mathsf{option}\,b))::\\
    \quad(x,\hid\,p\,(\mathsf{option}\,b))::\Gamma),
    \end{array}
    \\[2pt]
  \TT{yield} :
    [\TT{CommitPayloadTy }b] \to [\TT{DefaultVal }b] \to
    \Expr((x,b)::\erasectx{\Gamma},\bool) \to \\[1pt]
    \qquad\TT{VegasLang P }\Gamma_{\mathsf{yield}}
    \to \TT{VegasLang P }\Gamma .
\end{array}
\]
\TT{VegasYieldPhase P}~$\Gamma\,\pi\,\Delta$ is a typed list of same-phase
\TT{yield}s. Its guards are all typed over the original pre-phase context
$\Gamma$, not over the accumulated hidden prefix $\pi$, so a guard cannot
mention another same-phase yielded variable. The phase also carries the
\TT{DistinctActors}\label{def:DistinctActors} side condition.
The relevant lowering functions are:
\begin{quote}
\TT{VegasYieldPhase.lowerWith : VegasYieldPhase P }$\Gamma\,\pi\,\Delta$
  \TT{ -> VegasCore P simpleExpr }$\Delta$
  \TT{ -> VegasCore P simpleExpr }$(\pi\TT{ ++ }\Gamma)$\\
\TT{VegasLang.lower : VegasLang P }$\Gamma$
  \TT{ -> VegasCore P simpleExpr }$\Gamma$
\end{quote}

\paragraph{Math.}
\TT{VegasLang} is an elaborated typed surface language, not a second
denotation. Its meaning is definitionally the \TT{VegasCore} program obtained
by \TT{VegasLang.lower}\label{def:VegasLang.lower}. A thinner parser-level source language need not
construct this intermediate syntax: it can generate a fresh hidden commitment
name for \TT{yield}, translate ordinary reads to the immediately revealed
public name, translate owner-private reads to the hidden commitment name, and
emit \TT{VegasCore} directly.
The simultaneous constructor records that a source author intends several
public strategic moves, by distinct actors, to be presented together. Lowering
first emits all hidden nullable commits in source order and only then emits the
public reveals. After lowering, the program event-graph elaborator computes
the commit/reveal information barrier explicitly: the same-phase commits share
an event-graph frontier, while each reveal waits for all earlier commits in
addition to reading the committed secret it exposes.

\paragraph{Nullable yields.}
The ordinary surface \Commit{} constructor is restricted by
\TT{CommitPayloadTy}, so programmers cannot explicitly commit an optional
value. The user-facing public-move sugar is \TT{yield}. Lowering changes the
internal core payload type from $b$ to \TT{BaseTy.option b}, and replaces the
guard $R$ by
\TT{Expr.nullableCommitGuard R}. This lifted guard accepts \TT{none}
unconditionally and agrees with $R$ on \TT{some v}; hence
\TT{nullableCommitGuard_satisfiable}\label{def:nullableCommitGuard_satisfiable} proves the core \Legal{} obligation by
choosing \TT{none}.
The user-supplied guard is not required to be feasible: it constrains the
\TT{some} branch, while \TT{none} is always a legal lowered commitment.

The strategic cost is intentionally pushed into the continuation. After a
surface \TT{yield}, the revealed public value has type \TT{option b}. Payoff
and continuation expressions must inspect it with \TT{isNone}, \TT{isSome},
or \TT{getD}. Thus feasibility is obtained by construction, while the type
system forces the protocol author to decide how a null/abort branch affects
later payoffs.

\paragraph{Current scope.}
The surface language keeps \Let{} because \TT{VegasCore} keeps \Let{}.
It also deliberately omits quit handlers. Handler support is a later
extension because a strategic quit changes the action space and control flow;
it is not merely a value-level option.

\subsection{Visibility predicates: \TT{publicVars}, \TT{visibleVars},
  \TT{ObsEq}\label{def:ObsEq}, \TT{GuardUsesOnly}, \TT{ViewScoped}\label{def:ViewScoped}}
\label{def:scope}\label{def:publicVars}\label{def:visibleVars}

\paragraph{Lean.}
\begin{align*}
\mathtt{publicVars}\,\Gamma &: \mathrm{Finset}\,\mathrm{VarId} \\
\mathtt{visibleVars}\,p\,\Gamma &: \mathrm{Finset}\,\mathrm{VarId} \\
\mathtt{ObsEq}\,p\,\rho_1\,\rho_2 &\;:\;
  \mathrm{AgreesOn}\,\rho_1\,\rho_2\,(\mathtt{visibleVars}\,p\,\Gamma) \\
\mathtt{GuardUsesOnly}\,R &\;:\; \mathtt{exprDeps}\,R \subseteq
  \mathrm{insert}\,x\,(\mathtt{visibleVars}\,p\,\Gamma) \\
\mathtt{ViewScoped}\,p &\;:\;
  \text{recursive judgement on programs}
\end{align*}

\paragraph{Math.}
$\mathrm{publicVars}(\Gamma)$ and $\mathrm{visibleVars}(p, \Gamma)$ are the
sets of variable identifiers that are public, respectively visible to player
$p$. Two erased environments are \emph{observationally equivalent} for $p$,
written $\rho_1 \sim_p \rho_2$, if they agree on all variables in
$\mathrm{visibleVars}(p, \Gamma)$. The \TT{GuardUsesOnly} predicate says a
commit guard $R$ for player $p$ proposing variable $x$ depends only on
$\{x\} \cup \mathrm{visibleVars}(p, \Gamma)$. Finally
$\mathtt{ViewScoped}\,p$ is the recursive judgement that holds at \Commit{}
nodes and is forwarded transparently across other constructors.

\paragraph{Soundness.}
The pivotal lemma \TT{evalGuard_eq_of_obsEq}\label{def:evalGuard_eq_of_obsEq} says: \emph{provided the
proposed action variable $x$ is fresh in $\Gamma$}, if the guard satisfies
\TT{GuardUsesOnly} and $\rho_1 \sim_p \rho_2$, then evaluating the guard on
a proposed action against $\rho_1$ or $\rho_2$ yields the same result. The
freshness side condition is needed because $x$ shadows the head of the
extended context $(x, b) :: \erasectx{\Gamma}$ and we must know $x$ is not
already in $\Gamma$. This is the formal statement that \emph{a player's
commitment depends only on what they can see}.

\paragraph{Worked example.}
For a four-binding context
\[
  \Gamma \;=\;
    \bigl[\,(x_1,\,\pub\,\bool),\;
            (x_2,\,\hid\,p_2\,\intt),\;
            (x_3,\,\hid\,p_1\,\bool),\;
            (x_4,\,\pub\,\intt)\,\bigr],
\]
the per-player views are
\begin{align*}
  \view{p_1} &=
    \bigl[\,(x_1,\,\pub\,\bool),\;(x_3,\,\hid\,p_1\,\bool),\;(x_4,\,\pub\,\intt)\,\bigr],
    \\
  \view{p_2} &=
    \bigl[\,(x_1,\,\pub\,\bool),\;(x_2,\,\hid\,p_2\,\intt),\;(x_4,\,\pub\,\intt)\,\bigr],
\end{align*}
and $\pubctx{\Gamma} = \bigl[(x_1,\pub\,\bool),\,(x_4,\pub\,\intt)\bigr]$.
Neither player sees both hidden bindings; the two views overlap exactly on
$\pubctx{\Gamma}$.

% =====================================================================
\section{Well-formedness Obligations}

A raw \TT{VegasCore} program is just a tree. To turn it into a game, four
predicates must hold; they are bundled in \WFP.

\subsection{Freshness: \TT{Fresh}, \TT{FreshBindings}, \TT{WFCtx}}
\label{def:fresh}\label{def:Fresh}\label{def:WFCtx}

\paragraph{Math.}
$\mathrm{Fresh}(x, \Gamma)$ says $x$ does not appear in $\Gamma$. A
\emph{well-formed context} $\mathrm{WFCtx}\,\Gamma$ is one with no repeated names
($\mathtt{Nodup}\,(\Gamma.\mathtt{map}\,\mathtt{fst})$). The recursive
predicate $\mathrm{FreshBindings}$ asserts that every binder introduced by
\Let, \Sample, \Commit, or \Reveal{} is fresh in its enclosing context,
i.e.\ the program is in \emph{static-single-assignment} form (every
variable name is bound by exactly one program point and is never
reassigned).

\paragraph{Why.}
Unique names eliminate accidental shadowing in the visibility analysis and
make environment lookups proof-irrelevant. They are decidable.

\subsection{Reveal completeness: \TT{RevealComplete}}
\label{def:reveal}\label{def:RevealComplete}

\paragraph{Math.}
$\mathrm{RevealComplete}(\mathrm{pending}, p)$ tracks the set of currently
sealed commitments. At a \Commit{} node a fresh name is added; at a \Reveal{}
node it must be present and is removed. At \Ret{} the pending set must be
empty. Equivalently: every committed value is revealed exactly once before
the game ends. Decidable.

\paragraph{Why.}
Reveal completeness ensures every committed secret is publicly disclosed
before the game ends. It is \emph{not} a termination property — termination
follows from the finite syntax tree, captured by \TT{syntaxSteps}\label{def:syntaxSteps}, and the
horizon-bounded FOSG presentation. Payoff expressions type-check over the
public context regardless of revelation; what \texttt{RevealComplete} buys
is that all committed information becomes available to all players for
inspection (and, implicitly, that no commitment is irrelevant to the
public outcome).

\subsection{Legality and normalisation: \TT{Legal}, \TT{NormalizedDists}}
\label{def:legal}\label{def:Legal}\label{def:NormalizedDists}

\paragraph{Lean.}
\begin{quote}
\TT{def Legal} (program-level): every \Commit{} site has at least one
proposed action satisfying the guard.\\
\TT{def NormalizedDists}: every \Sample{} distribution sums to weight 1.
\end{quote}

\paragraph{Math.}
\Legal{} says, at each \Commit{} site for player $p$ with guard $R$ over
type $b$ and ambient erased environment $\rho$,
\[
  \exists\, a \in \Val(b).\, \mathrm{evalGuard}(R, a, \rho) = \mathsf{true}.
\]
This is a \emph{strategic non-deadlock}: the game never reaches a position
where the active player has no legal action. \NormDists{} is the
probabilistic well-formedness condition: each sample's evalDist is a
normalised finite weight, hence a true probability distribution after the
\TT{toPMF} bridge.

\paragraph{Decidability caveat.}
Both predicates are decidable when value domains are finite (the relevant
\TT{FiniteType}\label{def:FiniteType} instances are picked up by \TT{FiniteDomains}\label{def:FiniteDomains}). Without a
finite-type assumption, the existential in \Legal{} ranges over an
arbitrarily large value space.

\paragraph{Visibility note.}
\Legal{} permits the witness $a$ to depend on the entire erased
environment, while game-theoretic strategies must depend only on the
player's view. The gap is closed at the strategic API: equilibrium predicates
quantify over \emph{reachable legal} strategies of the FOSG view, in which
illegal actions are not in the carrier (\S\ref{sec:strategic}).

\subsection{The bundled checked program: \TT{WFProgram}}
\label{def:WFProgram}

\paragraph{Lean.}
\TT{structure WFProgram (P : Type) [DecidableEq P] (L : IExpr) where}
\begin{quote}
$\Gamma$\TT{ : VCtx P L} \\
\TT{prog : VegasCore P L }$\Gamma$ \\
\TT{env : VEnv L }$\Gamma$ \\
\TT{wctx : WFCtx }$\Gamma$ \\
\TT{wf : WF prog} \quad\textemdash\quad fresh, reveal-complete, view-scoped\\
\TT{normalized : NormalizedDists prog} \\
\TT{legal : Legal prog}
\end{quote}

\paragraph{Math.}
A \WFP{} is the API boundary at which raw syntax becomes a \emph{game}: a
program in some context $\Gamma$, an initial environment $\rho_0$, and four
static obligations — \TT{wctx}\label{def:wctx}, \TT{wf}, \TT{normalized}\label{def:normalized}, \TT{legal} —
needed to read the program as a finite stochastic extensive-form game.
\TT{wf} itself is a triple of \TT{FreshBindings}\label{def:FreshBindings}, \TT{RevealComplete~[]},
and \TT{ViewScoped}, so the bundle in fact records six independent facts.
Every public game-theoretic API in the codebase consumes \WFP, never raw
\TT{VegasCore}.

\paragraph{Program event-graph obligation bundle.}
The program event-graph elaborator has an additional local packaging layer:
\TT{ProgramNode.ProgramObligations p} bundles exactly the static facts needed
to interpret a source occurrence,
\TT{hctx}\label{def:hctx}, \TT{fresh}, \TT{hscoped}\label{def:hscoped}, \TT{legal}, and \TT{normalized}.  Its
tail projections \TT{letTail}\label{def:letTail}, \TT{sampleTail}\label{def:sampleTail}, \TT{commitTail}\label{def:commitTail}, and
\TT{revealTail}\label{def:revealTail} encode the recursive proof payloads for suffix programs.
\TT{WFProgram.obligations}\label{def:WFProgram.obligations} builds this bundle from a checked program $g$ and
is what \TT{programEventGraph}\label{def:programEventGraph} passes to \TT{ProgramNode.sem}\label{def:ProgramNode.sem},
\TT{prereqs}\label{def:prereqs}, \TT{sliceLegal}\label{def:sliceLegal}, \TT{actionLegal}\label{def:actionLegal}, and
\TT{internalKernel}\label{def:internalKernel}. This is proof/API plumbing: it does not add a new
semantic side condition, but it keeps the generated program event graph layer from
threading five separate proof arguments through every recursive call.

\paragraph{Finite domains.}
The class \TT{FiniteDomains}~$g$ provides instance evidence that all
operational types in $g$ are finite (\TT{FiniteVCtx}\label{def:FiniteVCtx} for the initial context
and \TT{FiniteProgram}\label{def:FiniteProgram} for the program). It is required for the \emph{public
strategic-form} games but is genuinely separate from \WFP{} because payoff
expressions return $\intt$, which is not finite.

% =====================================================================
\section{The Event Graph}

The first nontrivial semantic step elaborates a checked source program into a
DAG with extensional state. This subsection introduces the carrier; its
soundness theorems (frontier diamond and stability) are stated in the main
report.

\subsection{Stored values and write slices}

\paragraph{Lean.}
\begin{quote}
\TT{inductive WriteMode := clear | hidden} \\
\TT{inductive StoredValue (}$\alpha$\TT{) := clear (}$v$\TT{) | hidden (}$v$\TT{)} \\
\TT{abbrev WriteSlice G := (f : G.Field)} $\to$ \TT{Option (StoredValue (L.Val (G.fieldTy f)))}
\end{quote}

\paragraph{Math.}
A \emph{stored value} is a value tagged with whether it is currently public
(\textsf{clear}) or sealed (\textsf{hidden}). A \emph{write slice} is a
partial function from fields to stored values: it records what fields a
single event node writes when it executes. The empty slice is the all-\textsf{none}
function; \texttt{single} writes one field.

\subsection{Event-graph-local expressions, distributions, and guards}
\label{def:event-graph-local}

\paragraph{Lean.}
\TT{ReadEnv}\label{def:ReadEnv} packages the values of a declared finite read set;
\TT{EventExpr}\label{def:EventExpr} (deterministic), \TT{EventDist}\label{def:EventDist} (PMF-valued), and
\TT{EventGuard}\label{def:EventGuard} (boolean, with a \emph{visible}-versus-opaque read split and a
satisfiability obligation) take such read environments and return a value, a
distribution, or a Boolean.

\paragraph{Math.}
An event-graph-local construct evaluates from \emph{only the fields it declares as
reads}. This is an event-graph-level analogue of the source-level \TT{exprDeps}:
the event graph carries explicit local dependency information rather than a
syntactic over-approximation.

\paragraph{The visibility split inside guards.}
A \TT{EventGuard} keeps a subset \TT{visibleReads}\label{def:visibleReads} $\subseteq$ \TT{reads}\label{def:reads}
of \emph{visible} reads, plus a soundness obligation
\TT{eval_eq_of_visible_eq}\label{def:eval_eq_of_visible_eq}: any two read environments that agree on the
visible part must give the same Boolean. Conceptually, the guard's value
depends on the public-or-owner-visible reads only; the additional
\texttt{reads} are tolerated for syntactic convenience but cannot affect
truth value.

\subsection{Node semantics: \TT{NodeSem}}

\paragraph{Lean.}
\TT{inductive NodeSem (P)(F)(L)} with constructors
\TT{assign}\label{def:assign}, \TT{sample}\label{def:sample}, \TT{commit}\label{def:commit}, \TT{reveal}.

\paragraph{Math.}
Each event node carries one of four semantic payloads, mirroring (but not
identical to) the four non-\Ret{} source constructors. Several projections
are derived: $\mathrm{actor}$ (which player owns this node, if any),
$\mathrm{reads}$ (the fields read), $\mathrm{writes}$ (the field-write list),
$\mathrm{writeTarget}$ (the unique field this node writes), $\mathrm{writeFields}$
(its singleton finset form), and the predicates \TT{WritesField}\label{def:WritesField} and
\TT{WritesWithMode}\label{def:WritesWithMode} that locate a write by field and storage mode. Only
\TT{commit} has an actor — \TT{assign}, \TT{sample}, and \TT{reveal} are
internal. The single-field-write convention is exposed as the simp lemma
\TT{writeFields_eq_singleton}\label{def:writeFields_eq_singleton}: every node writes exactly one field.
The Boolean recognisers \TT{NodeSem.isCommit}\label{def:NodeSem.isCommit} and
\TT{NodeSem.isReveal}\label{def:NodeSem.isReveal} are used by the program
event-graph prerequisites to state commit/reveal phase barriers independently
of the storage fields read by a reveal.

\paragraph{Field reindexing.}
\TT{NodeSem.mapFields}\label{def:NodeSem.mapFields} transports a node-semantic payload through a field map
that preserves field types. It is the workhorse used by the program event-graph
elaborator to reuse the abstract \TT{NodeSem}\label{def:NodeSem} carrier over its concrete program
fields rather than maintaining a parallel ad-hoc layer.

\subsection{The event graph itself: \TT{EventGraph}}
\label{def:EventGraph}

\paragraph{Lean.}
\begin{quote}
\TT{Node, Field : Type}\\
\TT{nodeDecEq : DecidableEq Node}, \TT{fieldDecEq : DecidableEq Field}\\
\TT{nodes : Finset Node}, \TT{fields : Finset Field}\\
\TT{fieldTy : Field} $\to$ \TT{L.Ty}\label{def:L.Ty}, \TT{fieldOwner : Field} $\to$ \TT{Option P}\\
\TT{initial : (f : Field)} $\to$ \TT{Option (L.Val (fieldTy f))}\\
\TT{sem : Node} $\to$ \TT{NodeSem}, \TT{prereqs : Node} $\to$ \TT{Finset Node}\\
\TT{rank : Node} $\to$ \TT{Nat}\label{def:Nat}\\
\TT{prereqs_subset_nodes}\label{def:prereqs_subset_nodes}, \TT{prereq_rank_lt}\label{def:prereq_rank_lt}\\
\TT{read_fields_mem}\label{def:read_fields_mem}, \TT{write_fields_mem}\label{def:write_fields_mem}, \TT{no_duplicate_writes}\label{def:no_duplicate_writes}\\
\TT{sliceLegal : Node} $\to$ \TT{WriteSlice}\label{def:WriteSlice} $\to$ \TT{Prop}\label{def:Prop}\\
\TT{actionLegal : ResultAssignment} $\to$ \TT{Node}\label{def:Node} $\to$ \TT{WriteSlice} $\to$ \TT{Prop}\\
\TT{internalKernel : Node} $\to$ \TT{ResultAssignment}\label{def:ResultAssignment} $\to$ \TT{PMF WriteSlice}
\end{quote}
The last three lines use the namespace abbreviations \TT{WriteSlice G} and
\TT{ResultAssignment G}; these are definitionally the raw dependent function
types appearing in the structure fields.

\paragraph{Math.}
A \emph{event graph} is a finite, ranked DAG of typed nodes over a finite
set of typed fields, equipped with three pieces of legality and execution
data:
\begin{itemize}\itemsep1pt
\item \emph{sliceLegal} (\emph{shape}): a predicate on a single node and a
  candidate write slice. It is purely structural: for every kind of node,
  it checks that the slice writes the expected fields with the expected
  visibility/typing. It does \emph{not} consult guards or other nodes'
  results.
\item \emph{actionLegal} (\emph{dynamic}): a predicate that may consult the
  result assignment of already-completed nodes. For player commit actions
  this is where the \emph{guard} is evaluated against the proposed action
  and the current public/owner-visible reads.
\item \emph{internalKernel}: a $\mathrm{PMF}$ over write slices for non-player
  (\texttt{assign}, \texttt{sample}, \texttt{reveal}) nodes — these are
  internal events of the machine, not player actions. For \texttt{assign}
  and \texttt{reveal} this is a Dirac kernel; for \texttt{sample} it is the
  source-level distribution evaluated at the current public state.
\end{itemize}

\paragraph{Why the rank.}
The \texttt{rank} field is a strict topological order witness; it is the
termination measure for inductions over event-graph executions. The DAG itself
encodes \emph{causal} dependency: a node may only execute when its
prerequisites have produced results.

\subsection{Configurations and the frontier}
\label{def:Configuration}

\paragraph{Lean.}
\begin{quote}
\TT{abbrev ResultAssignment G := G.Node} $\to$ \TT{Option (WriteSlice G)} \\
\TT{structure Configuration G where}\\
\quad\TT{result : ResultAssignment G}\\
\quad\TT{result_nodes :} completed nodes belong to \TT{G.nodes}\label{def:G.nodes}\\
\quad\TT{closed :} completed nodes are closed under \TT{G.prereqs}\label{def:G.prereqs}\\
\quad\TT{legal :} stored slices satisfy \TT{G.sliceLegal}\label{def:G.sliceLegal}\\
\TT{noncomputable def frontier (cfg : Configuration G) : Finset G.Node}
\end{quote}

\paragraph{Math.}
A \emph{configuration} of an event graph is an extensional record of which
nodes have produced results. The dependency-closure invariant says that if
$n$ has a result, all of its prerequisites have results too:
\[
  \mathrm{result}(n)\ \mathrm{is\ }\mathsf{some}
  \;\implies\; \forall q \in \mathrm{prereqs}(n).\;
  \mathrm{result}(q)\ \mathrm{is\ }\mathsf{some}.
\]
The \emph{frontier} is the set of unfinished nodes whose prerequisites are
all finished: nodes that are \emph{enabled} to execute. Crucially, the
frontier is \emph{computed} from the result assignment, not stored.

\paragraph{Why extensional state.}
Recording only the result assignment, not the order in which nodes were
executed, makes the diamond property \emph{definitional}: two
linearisations that finish the same set of nodes with the same per-node
slices end up at literally the same configuration. A scheduling-based
state would need a separate quotient.

\paragraph{Diagram.}
\begin{center}
\begin{tikzpicture}[
  every node/.style = {circle, draw, minimum size=6mm, font=\small},
  done/.style  = {fill=green!18},
  fr/.style    = {fill=yellow!28, very thick},
  todo/.style  = {fill=gray!8},
  arr/.style   = {-{Stealth[length=2mm]}},
  >=Stealth,
]
\node[done] (a) at (0,0) {$a$};
\node[done] (b) at (1.4,1) {$b$};
\node[done] (c) at (1.4,-1) {$c$};
\node[fr]   (d) at (3,1) {$d$};
\node[fr]   (e) at (3,-1) {$e$};
\node[todo] (f) at (4.6,0) {$f$};
\draw[arr] (a)--(b); \draw[arr] (a)--(c);
\draw[arr] (b)--(d); \draw[arr] (c)--(e);
\draw[arr] (d)--(f); \draw[arr] (e)--(f);
\node[draw=none, font=\scriptsize, below right=0mm and -2mm of c] {done};
\node[draw=none, font=\scriptsize, below=0mm of e] {frontier};
\node[draw=none, font=\scriptsize, below=0mm of f] {pending};
\end{tikzpicture}
\end{center}

In this fragment, $\{a,b,c\}$ are completed, $\{d,e\}$ are on the frontier,
and $f$ is still waiting on its prerequisites. The frontier diamond
property says $d$ and $e$ can be executed in either order with the same
result.

% =====================================================================
\section{Machines}

The event graph is a static description; \TT{Machine} is the abstract
asynchronous execution carrier into which it is interpreted.

\subsection{The machine carrier}
\label{def:Machine}

\paragraph{Lean.}
\TT{structure Machine (P : Type)} with fields
\TT{State, Action : P} $\to$ \TT{Type}, \TT{Internal, Public, Outcome},
\TT{Obs : P} $\to$ \TT{Type}, \TT{init}\label{def:init}, \TT{available}\label{def:available},
\TT{availableInternal}\label{def:availableInternal}, \TT{stepPlay}\label{def:stepPlay}, \TT{stepInternal}\label{def:stepInternal} (PMF-valued),
\TT{terminal}\label{def:terminal}, \TT{publicView}\label{def:publicView}, \TT{observe}\label{def:observe}, \TT{outcome}\label{def:outcome}, \TT{utility}\label{def:utility}.

\paragraph{Math.}
A \emph{machine} is a tuple
\[
\begin{aligned}
  M \;=\; \bigl(\;
    & S,\; \{A_p\}_p,\; I,\; \mathrm{Pub},\; \{O_p\}_p,\; \Outcome,\;
       s_0, \\
    & \mathrm{avail},\; \mathrm{availInt},\;
       \mathrm{stepPlay},\; \mathrm{stepInternal},\\
    & \mathrm{term},\; \mathrm{pubView},\; \mathrm{obs},\; \mathrm{out},\; u
  \;\bigr)
\end{aligned}
\]
of: a state space; per-player action sets; an internal-event set; a public
observation type and per-player observation types; an outcome type with
utility $u : \Outcome \to (\Players \to \mathbb{R})$; an initial state; two
availability predicates (player-action $\mathrm{avail}$ and internal
$\mathrm{availInt}$); two transition kernels:
\[
  \mathrm{stepPlay}_M : (p : \Players) \to A_p \to S \to \PMF S, \qquad
  \mathrm{stepInternal}_M : I \to S \to \PMF S;
\]
a terminal predicate; and the projections to public/per-player observations
and to the terminal outcome. A primitive \emph{event} is either
$(\mathsf{play}, p, a)$ for $a \in A_p$ or $(\mathsf{internal}, e)$ for
$e \in I$; the combined kernel $\mathrm{step}_M : \mathrm{Event}_M \times S \to \PMF S$
is total.

\paragraph{Total step kernels, separate availability.}
The transition kernels are total functions; \emph{legality} of an event is a
\emph{separate} predicate. The event-graph machine instance (next subsection)
implements a \emph{stuttering} convention — unavailable events leave the
state unchanged, returning $\mathrm{PMF}.\mathrm{pure}\,s$ (a Dirac mass
at the source state). The abstract \TT{Machine} carrier does not
pre-commit to that convention; the kernel is exposed unconstrained so
backends can choose their reaction.

\subsection{Run prefixes, schedulers, traces}

\paragraph{Lean.}
\TT{structure RunPrefix} pairs a list of events with a list of states with
$|\mathrm{states}| = |\mathrm{events}| + 1$. \TT{abbrev EventLaw} is a
function $S \to \PMF{\mathrm{Event}}$, with \TT{IsLegalEventLaw}\label{def:IsLegalEventLaw} the
predicate that the law respects availability outside terminal states.

\paragraph{Math.}
A \emph{scheduler} is a history-independent law $\lambda : S \to \PMF{\mathrm{Event}}$
choosing the next event from a state. Iterating it for $n$ steps from $s_0$
yields a distribution over $\mathrm{Tr}_n(M)$ — finite event-state traces of
length $n$. Marginalising on the final state and projecting via $\mathrm{out}$
gives the \emph{outcome kernel} $\mathrm{outcomeKernelFrom}_\lambda$ of the
machine.

\subsection{The event-graph machine interpretation}

\paragraph{Lean.}
\TT{Vegas/EventGraph/ToMachine.lean} interprets an event graph as a machine:
\TT{State := Configuration}; \TT{Action := PlayerAction G};
\TT{Internal := InternalEvent G}.  A \TT{PlayerAction G p} is a structure
with fields \TT{node : G.Node} and \TT{slice : G.WriteSlice}; availability
restricts it to frontier nodes owned by $p$ with a legal slice. Internal
events are \TT{InternalEvent.node n} for non-player frontier nodes plus an
unavailable \TT{idle}\label{def:idle} constructor. \TT{stepPlay} and \TT{stepInternal}
execute one frontier node at a time, total in the same way as the abstract
machine.

\paragraph{Math.}
Reading the event graph as a machine, an event is the execution of one frontier
node. \TT{stepPlay} stutters on unavailable actions (returns
$\mathrm{PMF}.\mathrm{pure}\,\mathrm{cfg}$) and otherwise commits the chosen slice
into the configuration. The event graph's diamond property and stability of
frontier reads (\TT{withResult_comm}\label{def:withResult_comm} and the lemmas in
\TT{Vegas/GameBridge/FOSG/FrontierStability.lean}) are exactly what is needed to lift this
asynchronous, single-event view into a \emph{round-based} FOSG view in
which a player's round-action covers the entire owned frontier.  That lift
lives in \TT{Vegas/GameBridge/FOSG/FromEventGraph.lean}.

\paragraph{Stable frontier rounds: \TT{HasStableFrontierRounds}.}
\label{def:stableFrontier}
The lift from machine to FOSG is not unconditional: it requires the graph
to satisfy the predicate \TT{HasStableFrontierRounds}\label{def:HasStableFrontierRounds}, which combines two
conditions. \emph{Existence}: every player-owned frontier node has at
least one available player action (\TT{HasAvailablePlayerActions}\label{def:HasAvailablePlayerActions}).
\emph{Stability}: if an action $a$ for player $p$ is available at $\mathrm{cfg}$
and $a$'s node is on the frontier with $a.\mathrm{node} \neq \mathrm{first}$,
then after executing any other frontier node \texttt{first} legally, $a$ is
still available. Without stability, two players' frontier choices could
\emph{race} — one execution could invalidate the other's still-pending
action — and round-level batching would be ill-defined. Vegas's syntax
graphs satisfy this predicate (\TT{GameBridge/FOSG/FrontierStability.lean}); a
backend graph need not.

% =====================================================================
\section{FOSG Views}

The shift from machine to game-theoretic structure happens at this layer.

\subsection{Bounded states}

\paragraph{Lean.}
\TT{structure BoundedState (M : Machine P) (horizon : Nat) where state, depth, depth_le}.

\paragraph{Math.}
A \emph{bounded state} pairs a machine state with a depth counter, used to
present a finite-horizon FOSG. The counter is presentation data — the
underlying machine state remains semantic.

\subsection{The FOSG view: \TT{Machine.FOSGView}}
\label{def:FOSGView}\label{def:Machine.FOSGView}

\paragraph{Lean.}
\TT{structure FOSGView (M : Machine P) where}
\begin{quote}
\TT{Act : P} $\to$ \TT{Type} \\
\TT{active : M.State} $\to$ \TT{Finset P} \\
\TT{availableActions : M.State} $\to$ \TT{(p : P)} $\to$ \TT{Set (Act p)} \\
\TT{transition : (state : M.State)} $\to$
  \TT{ \{a : JointAction Act //}
  \TT{ JointActionLegal Act active M.terminal availableActions state a\}} $\to$
  \TT{PMF M.State} \\
\TT{reward : (src : M.State)} $\to$
  \TT{ \{a : JointAction Act //}
  \TT{ JointActionLegal Act active M.terminal availableActions src a\}} $\to$
  \TT{M.State}\label{def:M.State} $\to$ \TT{P} $\to$ $\mathbb{R}$ \\
(\TT{terminal_active_eq_empty}\label{def:terminal_active_eq_empty}, \TT{nonterminal_exists_legal}\label{def:nonterminal_exists_legal})
\end{quote}

\paragraph{Math.}
A \emph{factored-observation stochastic game view} on top of a machine is
the \emph{round-level} view: at each state, a finite set of \emph{active}
players choose round-actions from \TT{Act}\label{def:Act}, the state transitions
stochastically, and a per-player reward is paid out at the transition.
Crucially, the round-action alphabet \TT{Act} is independent of the
machine action alphabet \TT{M.Action}\label{def:M.Action}: a single round may consist of many
machine events. For event-graph protocols, a round-action is a slice for every
owned frontier node.

\paragraph{Round menus and observation equivalence.}
For event-graph machines the player-facing menu at a frontier round is packaged
as \TT{EventGraph.roundMenu G cfg who}: a set of optional round-actions
where \texttt{none} means the player is not active in the current round and
\texttt{some action} means the player is active and \texttt{action} is legal
on every frontier node they own. The accompanying theorem
\TT{eventGraph_roundMenu_eq_of_observation_eq}\label{def:eventGraph_roundMenu_eq_of_observation_eq} shows that two configurations
that are indistinguishable to a player offer that player the same menu —
the structural form of ``a player's available choices depend only on what
they observe.''

\paragraph{Why FOSG and not extensive form.}
Standard extensive-form games index strategies by histories; FOSG indexes
strategies by \emph{information sets}, which are equivalence classes of
histories under per-player observation. FOSG is a modern state-based
formalism that cleanly separates public and private observation; the
GameTheory submodule of this project formalises FOSGs once and Vegas wires
its machine semantics into the existing carrier.

\subsection{The canonical EFG presentation}

\paragraph{Lean.}
\TT{Vegas/GameBridge/EFG/FromEventGraph.lean} applies GameTheory's
\TT{FOSG.AugmentedEFGBridge}\label{def:FOSG.AugmentedEFGBridge} to the bounded program event-graph FOSG:
\begin{quote}
\TT{eventGraphEFGAt g : EFG.EFGGame} \\
\TT{eventGraphEFGBehavioralProfileAt g beta} \\
\TT{eventGraphEFGAt_outcomeKernel_map_publicOutcome}\label{def:eventGraphEFGAt_outcomeKernel_map_publicOutcome}
\end{quote}
The bridge tree is not a hand-written textbook tree. It is the canonical
serialisation of one FOSG round: one EFG decision opportunity per player
(over optional local moves), followed by a chance node for the FOSG
transition. Translated legal FOSG strategies put all mass on \TT{none} for
inactive players, so these administrative decision opportunities do not add
strategic freedom.

\paragraph{Tree structure, not only distributions.}
The semantic adequacy theorem says that projecting EFG terminal histories
back to public Vegas outcomes gives the same outcome law as
\TT{pmfBehavioralKernelGameAt}\label{def:pmfBehavioralKernelGameAt}. For reviewers who want to see the actual
tree spine, \TT{EFG.GameTree.HasDecisionChainThenChance}\label{def:EFG.GameTree.HasDecisionChainThenChance} records the
structural predicate ``$n$ sequential decisions followed by chance''. The
example file \TT{Examples/EFGTranslation.lean} proves this predicate for the
first generated round of the checked examples. Thus the code now checks both
the law induced by the EFG and the expected root shape of the generated tree.

\subsection{Bounded FOSGs and outcome kernels}

\paragraph{Math.}
\TT{toBoundedFOSG horizon} wraps a FOSG view in \TT{BoundedState}\label{def:BoundedState} so that
games are presented with a strict horizon — a hard prerequisite for many of
the finite-game theorems imported from the GameTheory library. The bounded
view exposes outcome kernels
\[
  \mathrm{outcomeFromPure}_h\,\pi : \PMF{\Outcome\,M},
  \qquad
  \mathrm{outcomeFromBehavioral}_h\,\beta : \PMF{\Outcome\,M},
\]
from pure and behavioural profiles respectively. The horizon $h$ is fixed
by the program: \TT{syntaxSteps} (\TT{Vegas/Core/Config.lean}) counts every
source constructor except \Ret, including administrative \Let{}s. The depth
counter in \TT{BoundedState} is presentation data needed to prove the
\emph{FOSG tree's} finiteness; the underlying event graph is already finite, but
without bounding the FOSG presentation could in principle re-visit
configurations.

\paragraph{Round-batching diagram.}
\begin{center}
\begin{tikzpicture}[
  node distance=4mm,
  ev/.style = {draw, rounded corners, fill=blue!8, font=\small,
    minimum height=6mm, minimum width=20mm},
  arr/.style = {-{Stealth[length=2mm]}, thick},
  bra/.style = {decorate, decoration={brace, amplitude=4pt, mirror}},
]
\node[ev] (e1) {$\mathrm{play}\,p_1\,a_1$};
\node[ev, right=of e1] (e2) {$\mathrm{play}\,p_2\,a_2$};
\node[ev, right=of e2] (e3) {$\mathrm{internal}\,e_3$};
\node[ev, right=of e3] (e4) {$\mathrm{play}\,p_1\,a_4$};
\node[ev, right=of e4] (e5) {$\mathrm{play}\,p_3\,a_5$};
\node[ev, right=of e5] (edots) {$\cdots$};
\draw[bra] (e1.south west) -- (e3.south east)
  node[midway, below=8pt, font=\small] {round 1: $\{p_1{:}a_1, p_2{:}a_2\}$};
\draw[bra] (e4.south west) -- (e5.south east)
  node[midway, below=8pt, font=\small] {round 2: $\{p_1{:}a_4, p_3{:}a_5\}$};
\end{tikzpicture}
\end{center}

\noindent The machine is the bottom layer (every primitive event explicit);
the FOSG view collapses each \emph{frontier batch} into a single FOSG round
whose action alphabet is a slice for every owned frontier node. Internal
events are absorbed silently. The stability predicate of
\S\ref{def:stableFrontier} is what makes this collapse well-defined.

% =====================================================================
\section{The Strategic API}
\label{sec:strategic}

For a checked program $g$, the codebase exposes a small public API of
strategy spaces and kernel games.

\subsection{Strategies and profiles}
\label{def:strategies}

\paragraph{Lean.}
\begin{align*}
\mathtt{pureStrategyAt}\;g\;p &: \mathbf{Type} \\
\mathtt{behavioralStrategyPMFAt}\;g\;p &: \mathbf{Type} \\
\mathtt{pureProfileAt}\;g &: \mathbf{Type} \\
\mathtt{behavioralProfilePMFAt}\;g &: \mathbf{Type}
\end{align*}

\paragraph{Math.}
For each player $p$ the FOSG view (\S\ref{def:FOSGView}) induces a finite
set of \emph{information states} $\mathcal{I}_p$: equivalence classes of
machine histories under $p$'s observation. The strategic carrier uses a
\emph{reachability-pruned} version
\[
  \mathrm{ReachableInfoState}\,p \;=\;
    \bigl\{\, s \in \mathcal{I}_p \;\bigm|\;
      \exists\, h : \mathrm{History}.\; h.\mathrm{playerView}\,p = s\,\bigr\}
\]
— information states \emph{realised by some history of the FOSG}, not by
some \emph{positive-probability} profile run. (This is purely combinatorial
and avoids quantifying over profiles in the carrier definition.)

A pure strategy for $p$ is then
\[
  \mathrm{ReachablePureStrategy}\,p \;=\;
    \mathrm{ReachableInfoState}\,p \;\to\; \mathrm{Option}\,(\mathrm{Act}\,p),
\]
returning \texttt{none} at information states where $p$ is not active and
\texttt{some}\,$a$ at active ones. A \emph{behavioural strategy} is
\[
  \mathrm{ReachableBehavioralStrategy}\,p \;=\;
    \mathrm{ReachableInfoState}\,p \;\to\; \PMF{\mathrm{Option}\,(\mathrm{Act}\,p)}.
\]
The Vegas \emph{legal} carriers (\TT{LegalPureProfile}\label{def:LegalPureProfile},
\TT{LegalBehavioralProfile}\label{def:LegalBehavioralProfile}) further restrict the support: at any reachable
information state where $p$ is active, the chosen \texttt{some}\,$a$ must
satisfy $a \in \mathrm{availableMoves}\,p\,h$ for an underlying realising
history $h$. \emph{This is what closes the gap from program-level \Legal{}
to strategic legality}: illegal moves are simply not in the strategy carrier
for the equilibrium predicates to range over.

\subsection{The four kernel games}
\label{def:kernelgames}

\paragraph{Lean.}
\begin{quote}
\TT{pureKernelGameAt g} (and \TT{Vegas.pureKernelGame}\label{def:Vegas.pureKernelGame}) — pure strategies, public outcome $\Outcome\,\Players$.\\
\TT{pmfBehavioralKernelGameAt g} (and \TT{Vegas.pmfBehavioralKernelGame}\label{def:Vegas.pmfBehavioralKernelGame}) — PMF behavioural strategies, public outcome.\\
\TT{pureBlockedTraceKernelGameAt g} — pure strategies, blocked-trace outcome.\\
\TT{pmfBehavioralBlockedTraceKernelGameAt g} — PMF behavioural strategies, blocked-trace outcome.
\end{quote}

\paragraph{Math.}
Each is an instance of \TT{GameTheory.KernelGame}\label{def:GameTheory.KernelGame}: per-player strategy types,
an outcome type, a stochastic kernel from profiles to outcomes, and a
utility function. The four games span two orthogonal axes:
\begin{itemize}\itemsep1pt
\item \emph{Strategy carrier}: pure vs.\ PMF-behavioural.
\item \emph{Outcome shape}: the public payoff $\Outcome\,\Players$ versus
  the richer \texttt{syntaxBlockedTraceAt}\,$g$ which retains the per-round
  primitive-event blocks and their endpoint machine state.
\end{itemize}
The blocked-trace games refine the public games by an outcome projection,
and \TT{Strategic/KernelGame.lean} packages this refinement (next paragraph).

\paragraph{Unsuffixed APIs.}
The unsuffixed names \TT{Vegas.pureKernelGame} and
\TT{Vegas.pmfBehavioralKernelGame} are thin definitions equal to their
\texttt{*At} counterparts under \TT{[Fintype P]} and \TT{[FiniteDomains g]};
they are definitionally equal, so all theorems transport.

\subsection{The kernel-game carrier}
\label{def:KernelGame}

\paragraph{Lean.}
\begin{quote}
\TT{structure KernelGame (}$\Players$\TT{) where} \\
\quad\TT{Strategy : }$\Players$ $\to$ \TT{Type} \\
\quad\TT{Outcome : Type} \\
\quad\TT{utility : Outcome} $\to$ \TT{Payoff}\label{def:Payoff} $\Players$ \\
\quad\TT{outcomeKernel : Kernel (}$\Pi$~\TT{i, Strategy i)} \TT{Outcome}\label{def:Outcome}
\end{quote}

\paragraph{Math.}
A \emph{kernel game} is the minimal object on top of which all
expected-utility solution concepts can be defined:
\[
\KG \;=\; \bigl(\{S_p\}_p,\,O,\,K : \textstyle\prod_p S_p \to \PMF{O},\;
  u : O \to \mathrm{Payoff}\,\Players\bigr),
\]
where the payoff vector $\mathrm{Payoff}\,\Players \;=\; \Players \to \mathbb{R}$.
Expected utility is the obvious functional
\[
  \EU(\sigma, p) \;=\; \mathbb{E}_{\omega \sim K(\sigma)}\bigl[u(\omega)(p)\bigr].
\]
Splitting $K$ from $u$ separates \emph{protocol} (\TT{GameForm}\label{def:GameForm}) from
\emph{preference}, which is what enables preference-parameterised solution
concepts (next subsection).

\paragraph{Two outcome types in the codebase.}
Vegas itself defines a simpler protocol-level outcome,
$\Outcome\,\Players = \Players \to_0 \mathbb{Z}$ (\S\ref{sec:outcomes}). The
public kernel games \TT{pureKernelGameAt}\label{def:pureKernelGameAt} and \TT{pmfBehavioralKernelGameAt}
take $O = \Outcome\,\Players$ and use the coercion
$u(o)(p) = (o\,p : \mathbb{R})$. The blocked-trace games set $O$ to
\[
  \mathtt{syntaxBlockedTraceAt}\,g \;=\;
    [\![\mathrm{event\ blocks}]\!] \times \mathtt{State}_M
\]
— each FOSG round's primitive-event block, paired with the round's endpoint
machine state. Their utility factors through the projection to the public
outcome: $\mathrm{tr} \mapsto u_{\mathrm{public}}(\mathrm{out}(\mathrm{tr}.\mathrm{state}))$.

\paragraph{Refinement layer.}
\TT{Strategic/KernelGame.lean} packages the relationship between the blocked-trace
and public games as a \TT{KernelGame.Simulation}\label{def:KernelGame.Simulation} (definitionally a
\TT{Morphism}\label{def:Morphism}, which preserves utility distributions
$\mathrm{udist}$ — i.e.\ the joint distribution over payoff vectors — under
profile mapping). It also provides explicit expected-utility equality
lemmas (\TT{pureBlockedTraceKernelGameAt_eu_eq}\label{def:pureBlockedTraceKernelGameAt_eu_eq} and the behavioural
counterpart) and \TT{Bisimulation}\label{def:Bisimulation}/\TT{EUBisimulation}\label{def:EUBisimulation} objects packaging
the blocked-trace versus public direction symmetrically. So reasoning at
either game level is sound: outcome-projection preserves utility
distributions and EU is provably equal across the projection.

\subsection{Linear source reading}
\label{def:LinearRead}

\paragraph{Lean.}
\TT{Vegas/Core/LinearRead.lean} defines
\TT{syntaxLinearRead g := ProgramNode.enumerate g.prog} and proves:
\begin{quote}
\TT{syntaxLinearRead_sufficient :}\\
\quad if a program event-graph configuration is nonterminal, then some
unfinished source node of minimal source rank is on the executable frontier.
\end{quote}
The reusable event-graph-level theorem is
\begin{quote}
\TT{EventGraph.Configuration.exists_rank_minimal_enabled_of_not_terminal}\label{def:EventGraph.Configuration.exists_rank_minimal_enabled_of_not_terminal}.
\end{quote}

\paragraph{Math.}
The source-order rank on \TT{ProgramNode}\label{def:ProgramNode}s is a topological order for the
elaborated event graph: every prerequisite of a node has strictly smaller
rank. Therefore a reader who scans the source program in rank order and stops
at the first unfinished source event has found an event whose dependencies
are already complete. The event-graph semantics may execute independent frontier
nodes in other orders, but the straight-line source presentation never blocks
early.

\paragraph{Commit/reveal information barrier.}
\begin{quote}\scriptsize
\begin{tabular}{@{}l@{}}
\TT{eventGraph_priorCommit_not_frontier_with_reveal}\label{def:eventGraph_priorCommit_not_frontier_with_reveal}\\
\TT{reveal_not_frontier_with_prior_commit}\label{def:reveal_not_frontier_with_prior_commit}\\
\TT{prior_commit_not_available_when_reveal_frontier}\label{def:prior_commit_not_available_when_reveal_frontier}
\end{tabular}
\end{quote}
The graph theorem and project-facing wrapper
state that an earlier commit and a reveal cannot be frontier-simultaneous.
The strategy-surface corollary says that a player action for such an earlier
commit is not available while the reveal is on the frontier. This is the
theorem-level guard against observing a reveal before choosing an earlier
same-phase commitment.

\paragraph{Role.}
Source order is a readable topological schedule for the dependency graph:
following it never gets stuck before the event graph is terminal. Strategic
equivalence across stable frontier schedules is then proved by the
frontier-stability, blocked-trace, and morphism layers.

\subsection{Solution concepts (Vegas-facing definitions)}
\label{def:solution}\label{def:IsNash}\label{def:IsStrictNash}\label{def:WeaklyDominates}\label{def:StrictlyDominates}\label{def:IsCorrelatedEq}\label{def:IsCoarseCorrelatedEq}\label{def:ParetoDominates}\label{def:IsParetoEfficient}\label{def:socialWelfare}\label{def:IsExactPotential}\label{def:IsOrdinalPotential}\label{def:IsZeroSum}\label{def:IsConstantSum}\label{def:IsTeamGame}\label{def:optimalWelfare}\label{def:bestNashWelfare}\label{def:worstNashWelfare}\label{def:worstNashWelfare_le_bestNashWelfare}\label{def:bestNashWelfare_le_optimalWelfare}\label{def:ciSup}

The Vegas definitions (\TT{Vegas.IsNash}\label{def:Vegas.IsNash}, \TT{IsBestResponse}\label{def:IsBestResponse}, \TT{IsDominant}\label{def:IsDominant},
\ldots) are thin wrappers around GameTheory's preference-parameterised
predicates over the PMF behavioural kernel game
\TT{pmfBehavioralKernelGame g}. They are spread across
\TT{Vegas/Strategic/Equilibrium.lean} (Nash, dominance, correlated equilibria, social
welfare) and \TT{Vegas/Strategic/Properties.lean} (approximate Nash, potentials,
constant-sum, welfare and price-of-anarchy). We list the underlying
mathematical definitions; for each, the Lean wrapper has the form
$\TT{Vegas.}X = (\TT{pmfBehavioralKernelGame}\label{def:pmfBehavioralKernelGame}\,g).X$. We use $\EU(\sigma, p)$
throughout for expected utility.

\smallskip
\noindent\begin{tabular}{@{}lp{0.55\textwidth}@{}}
\toprule
\textbf{Predicate} & \textbf{Mathematical content (over $\KG$, $\sigma : \Pi_p S_p$)} \\
\midrule
\TT{IsNash}~$\sigma$ &
  $\forall p,\, \forall s'_p \in S_p.\; \EU(\sigma, p) \ge \EU(\sigma[p \mapsto s'_p], p)$. \\
\TT{IsStrictNash}~$\sigma$ &
  As above with $>$ for $s'_p \neq \sigma_p$. \\
\TT{IsBestResponse}~$p\,\sigma\,s$ &
  $\forall s'.\; \EU(\sigma[p\mapsto s], p) \ge \EU(\sigma[p\mapsto s'], p)$. \\
\TT{IsDominant}~$p\,s$ &
  $\forall \sigma\, s'.\; \EU(\sigma[p\mapsto s], p) \ge \EU(\sigma[p\mapsto s'], p)$. \\
\TT{WeaklyDominates}~$p\,s\,t$ &
  $\forall \sigma_{-p}.\, \EU((s, \sigma_{-p}), p) \ge \EU((t, \sigma_{-p}), p)$. \\
\TT{StrictlyDominates}~$p\,s\,t$ &
  As above with $>$. \\
\TT{IsCorrelatedEq}~$\mu$ &
  Profile distribution $\mu$: no recommendation-dependent unilateral
  deviation strictly improves any player's outcome distribution. \\
\TT{IsCoarseCorrelatedEq}~$\mu$ &
  As above, restricted to constant deviations. \\
\TT{ParetoDominates}~$\sigma\,\tau$ &
  $\forall p.\,\EU(\sigma, p) \ge \EU(\tau, p)$ and $\exists p.\,>$. \\
\TT{IsParetoEfficient}~$\sigma$ &
  $\nexists \tau.\; \mathrm{ParetoDominates}\,\tau\,\sigma$. \\
\TT{socialWelfare}~$\sigma$ &
  $\sum_p \EU(\sigma, p)$. \\
\TT{IsExactPotential}~$\Phi$ &
  $\Phi : \Pi_p S_p \to \mathbb{R}$ such that
  $\Phi(\sigma[p\!\mapsto\!s']) - \Phi(\sigma) = \EU(\sigma[p\!\mapsto\!s'], p) - \EU(\sigma, p)$. \\
\TT{IsOrdinalPotential}~$\Phi$ &
  Sign-only version of above. \\
\TT{IsZeroSum}, \TT{IsConstantSum}~$c$ &
  $\sum_p u(\omega)(p)$ is constant ($=0$ or $=c$) on all outcomes. \\
\TT{IsTeamGame} &
  $u(\omega)(p) = u(\omega)(q)$ for all $p, q, \omega$. \\
\texttt{Is}$\varepsilon$\texttt{Nash}~$\sigma$ &
  $\forall p\, s'_p.\, \EU(\sigma, p) + \varepsilon \ge \EU(\sigma[p\mapsto s'_p], p)$. \\
\TT{optimalWelfare} &
  $\sup_\sigma \mathrm{socialWelfare}(\sigma)$ (\TT{ciSup}; finite-profile
  variants are picked up by Mathlib's boundedness lemmas). Vegas also
  defines \TT{bestNashWelfare} and \TT{worstNashWelfare}, plus the
  comparison lemmas \TT{worstNashWelfare_le_bestNashWelfare} and
  \TT{bestNashWelfare_le_optimalWelfare}. \\
\bottomrule
\end{tabular}

\paragraph{Preference parameterisation.}
GameTheory exposes \TT{IsNashFor pref}, \TT{IsBestResponseFor pref}, etc.,
where \TT{pref}\label{def:pref} is an arbitrary preference relation over $\PMF\Outcome$.
The Vegas-named EU versions arise by instantiating
$\mathtt{pref} = \mathtt{euPref}$ (i.e., expected-value comparison).

\paragraph{Existence corollaries.}
\TT{Vegas/Corollaries/GameProperties.lean} exposes three finite-game
existence theorems through the Vegas-facing API:
\TT{mixedNash_exists}\label{def:mixedNash_exists}, \TT{correlatedEq_exists}\label{def:correlatedEq_exists}, and
\TT{coarseCorrelatedEq_exists}\label{def:coarseCorrelatedEq_exists}. These are direct wrappers around the
GameTheory fixed-point and correlated-equilibrium existence theorems for the
kernel game \TT{pmfBehavioralKernelGame g}; they are intentionally stated at
the PMF-over-strategies level and do not by themselves construct a named
source-level policy.

\subsection{Mixed extension and Kuhn realisation}
\label{def:Kuhn}

\paragraph{Lean.}
\begin{quote}
\TT{KernelGame.mixedExtension G : KernelGame}; \\
\TT{Vegas.kuhn_finite}\label{def:Vegas.kuhn_finite}: as in the main report.
\end{quote}

\paragraph{Math.}
The \emph{mixed extension} of a finite kernel game $\KG$ replaces each
strategy space $S_p$ by $\PMF S_p$. The outcome kernel of the extension
samples $\sigma \sim \prod_p \mu_p$ from the independent product
distribution and applies the underlying kernel:
\[
  K^{\mathrm{mix}}(\mu) \;=\; \bigl(\textstyle\prod_p \mu_p\bigr)
    \bind \lambda \sigma. K(\sigma).
\]

\emph{Kuhn realisation} for finite Vegas programs (\TT{kuhn_finite}\label{def:kuhn_finite}): every
independent mixed profile over the pure strategy carrier of
\TT{pureKernelGameAt}$\,g$ is matched by a behavioural profile of
\TT{pmfBehavioralKernelGameAt}$\,g$ with the \emph{same outcome
distribution}. Symbolically, with $\KG^{\mathrm{pure}}$ and
$\KG^{\mathrm{beh}}$ the two kernel games:
\[
  \forall \mu \in \textstyle\prod_p \PMF{S_p^{\mathrm{pure}}}.\;\;
  \exists \beta : \mathrm{Profile}(\KG^{\mathrm{beh}}).\;\;
    K^{\mathrm{beh}}(\beta) \;=\;
    \bigl(\textstyle\prod_p \mu_p\bigr) \bind K^{\mathrm{pure}}.
\]

\paragraph{Hypotheses behind the slogan.}
The Lean statement is a Kuhn-style realisation theorem, not merely the
textbook phrase ``perfect recall'' copied verbatim. It is routed through the
bounded FOSG view of the program event-graph machine: finite domains provide the
finite-horizon carriers, and the required information condition is
\TT{LegalObservable}\label{def:LegalObservable} for the bounded FOSG, proved for checked program event graphs
from the visibility and graph-stability invariants.

\paragraph{Why this is the headline theorem.}
Pure strategies are easy to read and define; behavioural strategies are
what equilibrium analyses on extensive-form games actually use. The Kuhn
realisation says that for every mixture over the simple objects we can
build a behavioural-profile witness with identical observable behaviour.
This is the formal contract between the surface API (\TT{pureKernelGame}\label{def:pureKernelGame})
and the analysis API (\TT{pmfBehavioralKernelGame}).

% =====================================================================
\section{Outcome and utility evaluation}
\label{sec:outcomes}

\subsection{Outcomes: \TT{Outcome}, \TT{mkOutcome}, \TT{evalPayoffs}}

\paragraph{Lean.}
\begin{quote}
\TT{abbrev Outcome (P : Type) [DecidableEq P] :=} $P \to_0 \mathbb{Z}$ \\
\TT{def mkOutcome (entries : List (P} $\times$ \TT{Int)) : Outcome P} \\
\TT{def evalPayoffs (payoffs : List (P} $\times$ \TT{L.Expr (erasePubVCtx }$\Gamma$\TT{) L.int)) : VEnv L }$\Gamma$ $\to$ \TT{Outcome P}
\end{quote}

\paragraph{Math.}
A protocol \emph{outcome} is a finitely-supported $\mathbb{Z}$-valued payoff
vector indexed by players: $\Outcome = \Players \to_0 \mathbb{Z}$.
\TT{mkOutcome}\label{def:mkOutcome} sums duplicate keys; absent players default to payoff $0$.
\TT{evalPayoffs}\label{def:evalPayoffs} takes the syntactic payoff list of a \Ret{} node, evaluates
each integer expression in the public-erased environment, and packages the
result.

\paragraph{Why \TT{Finsupp}, not \TT{P} $\to$ \TT{Int}.}
\TT{Finsupp}\label{def:Finsupp} is decidably equal and decidably has-zero, both of which are
needed for $\PMF{\Outcome}$ to type-check. The runtime semantics of ``only
named players are paid'' is the same in either form.

% =====================================================================
\section{The Backend Layer (refinement)}
\label{def:Backend}

\paragraph{Lean.}
\TT{Vegas/Backend/Refinement.lean} defines two machine-to-machine relations.
Both carry
\begin{quote}
\TT{projectState}\label{def:projectState}, \TT{projectEvent}\label{def:projectEvent}, \TT{projectPublic}\label{def:projectPublic},
\TT{projectObs}\label{def:projectObs}, \TT{projectOutcome}\label{def:projectOutcome}, and \TT{init_project}\label{def:init_project}.
\end{quote}
\TT{WeakStepRefinement Impl Spec} then adds
\begin{quote}
\TT{external_step_sound}\label{def:external_step_sound}, \TT{internal_step_projectState_eq}\label{def:internal_step_projectState_eq},
\TT{publicView_project}\label{def:publicView_project}, \TT{observe_project}\label{def:observe_project}, \TT{terminal_project}\label{def:terminal_project},
\TT{terminal_outcome_projected}\label{def:terminal_outcome_projected}.
\end{quote}
\TT{StochasticStepRefinement Impl Spec} instead adds
\begin{quote}
\TT{step_project}\label{def:step_project}, \TT{publicView_project}, \TT{observe_project},
\TT{terminal_project}, \TT{terminal_reflect}\label{def:terminal_reflect},
\TT{terminal_outcome_projected}.
\end{quote}

\paragraph{Math.}
\emph{Refinement} here is the verification-style relation ``the
implementation cannot be distinguished from the specification by any
externally observable behaviour''. Concretely, a refinement is a state
projection $\pi : S_{\mathrm{Impl}} \to S_{\mathrm{Spec}}$ satisfying
matching conditions on transitions. The codebase formalises two strengths:
\begin{itemize}
\item \emph{Weak step refinement.} Implementation external (player) events
  match a specification event under the state projection; internal events
  preserve the projected state.
\item \emph{Stochastic step refinement.} A strictly stronger version: the
  pushforward of the implementation step kernel through the projection equals
  the specification's kernel (with internal events stuttering on the spec
  side). This implies weak refinement
  (\TT{StochasticStepRefinement.toWeakStepRefinement}\label{def:StochasticStepRefinement.toWeakStepRefinement}).
\end{itemize}
Every machine also refines itself by \TT{StochasticStepRefinement.refl}\label{def:StochasticStepRefinement.refl}; this
identity refinement is useful both as a baseline backend and as a compact
smoke test for trace-projection theorems.

\paragraph{Why a separate layer.}
Implementations on real runtimes have additional machinery (transactions,
schedulers, gossip) that does not appear in the source language. Refinement
is the way to attest that the user's reasoning at the FOSG/kernel-game level
remains valid against an implementation, without growing a second
``implementation semantics'' inside the source language. Trace-level
projection lemmas (\TT{runEventsFrom_project_eq}\label{def:runEventsFrom_project_eq},
\TT{blockTraceDist_project_eq}\label{def:blockTraceDist_project_eq}) carry the refinement from primitive
events to whole runs.

\paragraph{Backend kernel-game transport.}
The file \TT{Vegas/Backend/KernelGame.lean} packages a
\TT{StochasticStepRefinement}\label{def:StochasticStepRefinement} as a kernel-game morphism. The data are:
\begin{itemize}\itemsep1pt
\item \TT{BackendPureBlockLawLift}\label{def:BackendPureBlockLawLift}~$g~R$: a per-pure-profile pair of
  block-laws (one for the spec graph machine, one for the implementation),
  proved \emph{compatible} via \TT{R.BlockLawCompatible}\label{def:R.BlockLawCompatible}, and \emph{sound}
  in the sense that the spec block-law reproduces the canonical
  \TT{pureBlockedTraceKernelGameAt}\label{def:pureBlockedTraceKernelGameAt} outcome kernel.
\item \TT{BackendBehavioralBlockLawLift}\label{def:BackendBehavioralBlockLawLift}~$g~R$: the analogous datum for PMF
  behavioural profiles.
\item \TT{backendBlockedTraceKernelGameAt}\label{def:backendBlockedTraceKernelGameAt}: a kernel game whose strategy
  spaces are the canonical Vegas pure strategies but whose outcomes are
  full implementation block-traces.
\item \TT{toBackendBlockedTraceMorphism}\label{def:toBackendBlockedTraceMorphism} and its behavioural counterpart
  expose, for every refinement plus lift, a \TT{KernelGame.EUMorphism}\label{def:KernelGame.EUMorphism}
  whose direction is
  \[
    \mathrm{backend\ game}\;\longrightarrow\;\mathrm{spec\ blocked{-}trace\ game}.
  \]
  This is the direction needed to \emph{pull Nash equilibria back from the
  spec to the backend}: a Nash on the spec kernel game corresponds to a
  Nash on the backend game by composition with the morphism.
\end{itemize}
A backend author supplies \TT{StochasticStepRefinement} plus a block-law
lift and obtains an EU-preserving morphism from the implementation into
the canonical Vegas blocked-trace game.

\paragraph{Transport proofs use the morphism API.}
The two public pullback theorems
\begin{quote}
\TT{backendBlockedTraceKernelGameAt_isNash_pullback}\label{def:backendBlockedTraceKernelGameAt_isNash_pullback}, \\
\TT{backendPMFBehavioralBlockedTraceKernelGameAt_isNash_pullback}\label{def:backendPMFBehavioralBlockedTraceKernelGameAt_isNash_pullback}
\end{quote}
are proved by applying \TT{KernelGame.EUMorphism.nash_of_nash}\label{def:KernelGame.EUMorphism.nash_of_nash}. The composed
``public game to backend'' theorems first use the Stage-1
\TT{EUGameIsomorphism.nash_iff}\label{def:EUGameIsomorphism.nash_iff} between public and blocked-trace games,
then the backend morphism. Thus the code exercises the same morphism
composition story stated in the report rather than repeating
expected-utility calculations locally.

% =====================================================================
\section{Worked Examples}

The codebase contains checked game examples, constructor-level smoke tests,
and backend/refinement examples. We summarise the textbook games first, then
the files whose purpose is to exercise API surfaces.

\subsection{Prisoners' Dilemma}

\paragraph{Game.}
Two players \texttt{row}, \texttt{col} simultaneously choose
$\mathsf{C}$~(\texttt{true}, cooperate) or $\mathsf{D}$~(\texttt{false},
defect). The bimatrix:
\[
\arraycolsep=8pt
\begin{array}{c|cc}
& \mathsf{C} & \mathsf{D} \\
\hline
\mathsf{C} & (2,2) & (0,3) \\
\mathsf{D} & (3,0) & (1,1)
\end{array}
\]
Unique Nash: $(\mathsf{D}, \mathsf{D})$. Defect is dominant for both.

\paragraph{Vegas encoding.}
Two \Commit{}s (one per player), each followed later by a \Reveal{}, then
\Ret{} with payoffs:
\begin{quote}\small\ttfamily
.commit rowSecret row (.constBool true)\\
\hspace*{1em}(.commit colSecret col (.constBool true)\\
\hspace*{2em}(.reveal rowPublic row rowSecret \_\\
\hspace*{3em}(.reveal colPublic col colSecret \_\\
\hspace*{4em}(.ret [(row, rowPayoff), (col, colPayoff)]))))
\end{quote}
Both guards are constant \texttt{true}, so any value is a legal
commitment; the strategic content is in the payoff expressions, which read
the two public bindings. In the generated event graph, both commits can share
the initial frontier, but neither reveal can be frontier-simultaneous with an
earlier unchosen commit.

\paragraph{Why the trivial guards.}
A \Commit{} guard restricts \emph{which actions are legal at the protocol
level}. Here every action is legal, and the strategic question is which one
the player should choose — a question answered by the equilibrium predicates
on the kernel game, not by the program syntax.

\subsection{Matching Pennies}

\paragraph{Game.}
Two players, $\mathsf{matcher}$ and $\mathsf{mismatcher}$, simultaneously
choose Heads/Tails. The matcher is paid $+1$ on a match, $-1$ otherwise; the
mismatcher's payoff is the matcher's negated. Textbook fact: no pure Nash
exists; the unique mixed Nash plays each action with probability $1/2$.
The Vegas example mechanises the per-payoff calculations (zero-sum
identities on the payoff expressions). The non-existence of a pure Nash
is proved separately in a hand-written deterministic kernel-game example,
not on the generated kernel game from the checked Vegas program; the
analytical mixed Nash is textbook context, not a Lean theorem.

\paragraph{Vegas encoding.}
Identical commit/reveal shape to Prisoners: the two commits are simultaneous
frontier events, and the reveal phase begins only after both commitments have
completed. Payoff expressions use \TT{simpleExpr}'s \TT{eq} on the two public
bindings to test for a match.

\paragraph{\TT{VegasLang} surface presentation.}
\TT{Examples/VegasLang.lean} contains two surface presentations. The
non-nullable presentation uses the core-shaped \TT{commit}/\TT{reveal}
constructors and proves
\[
  \TT{VegasLang.lower canonicalLangProgram = game.prog}
\]
by \TT{rfl}\label{def:rfl}. The nullable presentation uses a single simultaneous yield phase:
\begin{quote}\small\ttfamily
.simultaneous simultaneousYieldPhase simultaneousYieldPhase\_distinct\\
\hspace*{1em}(.ret [(matcher, matcherYieldPayoff),\\
\hspace*{3em}(mismatcher, mismatcherYieldPayoff)])
\end{quote}
Here each public choice has type \TT{option bool}; the payoff expressions use
\TT{getD} to decide how the null branch is handled. This nullable presentation
therefore lowers to a different core program, and the file proves
\[
  \TT{VegasLang.lower langProgram = coreProgram}
\]
by \TT{rfl}, where \TT{coreProgram}\label{def:coreProgram} is the explicit nullable-commit program
with both commits emitted before either reveal.

\paragraph{Pedagogical role.}
This example is the smallest where mixed strategies are necessary. The
kernel game produced by \TT{pmfBehavioralKernelGame} carries enough
information to express mixed-Nash claims. \TT{kuhn_finite} is a
\emph{realisation} theorem (every mixed pure profile is matched by a
behavioural profile with the same outcome distribution); it does not
itself construct or prove the existence of any equilibrium. Existence
proofs would compose Kuhn realisation with a separate fixed-point
existence theorem.

\subsection{Battle of the Sexes}

\paragraph{Game.}
Two players prefer to coordinate but disagree on which venue:
\[
\arraycolsep=8pt
\begin{array}{c|cc}
& \mathsf{Opera} & \mathsf{Football} \\
\hline
\mathsf{Opera} & (2,1) & (0,0) \\
\mathsf{Football} & (0,0) & (1,2)
\end{array}
\]
Two strict pure Nash equilibria, plus a mixed one.

\paragraph{Vegas encoding.}
Same shape as the previous two. The interest is purely game-theoretic
(multiple equilibria), not protocol-level.

\subsection{Monty Hall}

\paragraph{Game.}
Two players, $\mathsf{host}$ and $\mathsf{guest}$, with a bounded action
type \TT{Door := range 0 2}.
\begin{enumerate}[nosep]
\item Host commits the car location.
\item Guest commits a first guess.
\item Guest's first guess is revealed.
\item Host commits an opened door, with guard:
  $\mathsf{opened} \neq \mathsf{guest's\ first\ guess}$ and
  $\mathsf{opened} \neq \mathsf{car}$.
\item Opened door is revealed.
\item Guest commits a switch decision (boolean).
\item Switch decision is revealed.
\item Car location is revealed.
\item Payoffs.
\end{enumerate}
The classical analysis: switching wins iff the first guess was wrong, which
happens with probability $2/3$ under a uniform prior on the car location.

\paragraph{Vegas encoding.}
The longest example (eight non-\Ret{} steps). It is the only example with a
non-trivial commit guard — the host's \TT{commit} of the opened door has a
guard depending on two earlier bindings — and it uses the bounded
\TT{range 0 2} type to enumerate doors.

\paragraph{Mathematically non-obvious bits.}
\begin{enumerate}\itemsep1pt
\item The example does not introduce a \Sample{} for the car location: the
  host \Commit{}s it. To recover the textbook ``car uniformly placed''
  analysis, one would quantify in the equilibrium claim over (mixed) host
  strategies. The examples now include a finite enumeration proof of the
  familiar closed forms: switching has average value $2/3$ and staying has
  average value $1/3$ under the uniform door count.
\item The host's distribution over legal doors can itself be informative.
  For example, if the guest first picks door $0$ and the host opens door $1$,
  write $q$ for the probability that the host opens door $1$ in the branch
  where the car is actually behind door $0$; in the branch where the car is
  behind door $2$, opening door $1$ is forced. The checked calculation gives
  posterior switch value $1/(q+1)$ and stay value $q/(q+1)$. Thus, under the
  standard uniform-car/legal-host assumptions, switching remains weakly
  optimal for $0 \le q \le 1$, with a tie at the maximally biased endpoint
  $q=1$.
\item The host's open-door guard reads the host's hidden \texttt{carSecret}
  binding and the freshly-revealed public \texttt{firstPublic} alias of the
  guest's first guess. It is precisely \TT{GuardUsesOnly} for the host:
  both reads are visible to the host (\texttt{carSecret} as owner;
  \texttt{firstPublic} as a public).
\end{enumerate}

\subsection{SyntaxConstructors}

\TT{Examples/SyntaxConstructors.lean} is a deliberately tiny checked program
whose purpose is not strategic novelty but constructor coverage. It uses a
deterministic \Let{} to bind a public boolean flag, then a public \Sample{}
from a fair coin, then \Ret{} with payoff $1$ exactly when both are true. The
file proves the \TT{FWeight.totalWeight}\label{def:FWeight.totalWeight} normalisation obligation for the sample,
checks the payoff expression at both coin outcomes, packages the program as a
\TT{WFProgram}, and exposes the generated behavioural kernel game. This is the
minimal example a reviewer should inspect when asking whether \Let{} and
\Sample{} are actually exercised by checked programs.

\subsection{NullableCommit}

\TT{Examples/NullableCommit.lean} is the surface-language smoke test for
nullable yields. The example uses an impossible boolean guard
\TT{impossibleBoolGuard := .constBool false}. A normal core commit with this
guard would fail \Legal{}, because no boolean value satisfies it. The
\TT{VegasLang.yield}\label{def:VegasLang.yield} surface spelling lowers to a core commit of
type \TT{BaseTy.option .bool} with guard
\TT{Expr.nullableCommitGuard impossibleBoolGuard}; the file proves
\[
  \TT{VegasLang.lower langProgram = coreProgram}
\]
by \TT{rfl}, and proves \TT{lowered_program_legal}\label{def:lowered_program_legal} using
\TT{nullableCommitGuard_satisfiable}.

The same file connects nullable yields to a simple griefing criterion. It
defines \TT{NonCostlyGriefing}\label{def:NonCostlyGriefing} for two moves, ``abort'' and ``honest'': the
abort is non-costly for the actor and strictly worse for the victim. Under
\TT{badActorPayoff}\label{def:badActorPayoff}/\TT{badVictimPayoff}\label{def:badVictimPayoff}, the null branch leaves the actor at
utility $0$ while charging the victim $-10$, so
\TT{bad_null_is_noncostly_griefing}\label{def:bad_null_is_noncostly_griefing} proves that \TT{none} is a non-costly
griefing move. Under \TT{goodActorPayoff}\label{def:goodActorPayoff}/\TT{goodVictimPayoff}\label{def:goodVictimPayoff}, the null
branch charges the actor $-10$, and the theorems
\TT{good_null_is_punished}\label{def:good_null_is_punished} and \TT{good_null_not_noncostly_griefing}\label{def:good_null_not_noncostly_griefing} prove
that the same null move is no longer non-costly.

\paragraph{Strategic lesson.}
Nullable yield makes an infeasible guarded action legal by adding an explicit
abort value. It does not decide who pays for that abort. Because the revealed
value has option type, later payoff expressions must handle \TT{none}. If the
author handles it by harming only the counterparty, the program exposes a
griefing action; if the author punishes the nulling player, that particular
non-costly griefing opportunity is removed.

\subsection{BackendExamples}

\TT{Examples/BackendExamples.lean} uses \TT{StochasticStepRefinement.refl} as
an identity backend. It proves that identity projection preserves event
blocks, block traces, compatible block laws, and blocked-trace distributions,
then instantiates the result for the \TT{SyntaxConstructors.game}\label{def:SyntaxConstructors.game} program event graph
machine. This is intentionally boring: it is a small executable witness that
the backend-refinement API is usable before a real implementation supplies
extra state or internal events.

\subsection{EFGTranslation}

\TT{Examples/EFGTranslation.lean} checks the extensive-form presentation of
the examples, not merely their induced outcome distributions. For
\TT{Prisoners}\label{def:Prisoners}, \TT{MatchingPennies}\label{def:MatchingPennies}, \TT{BattleOfTheSexes}\label{def:BattleOfTheSexes}, and
\TT{MontyHall}\label{def:MontyHall}, the file proves that the first generated EFG round has two
decision nodes followed by a chance transition. For
\TT{SyntaxConstructors}\label{def:SyntaxConstructors}, whose player type has one inhabitant, it proves one
decision node followed by chance. The same file also retains the public
outcome-law theorem for each example:
\[
\begin{aligned}
  & \mathrm{map}(\mathrm{publicOutcome})\,
    \mathrm{EFG.outcomeKernel}(\mathrm{translated}\ \beta) \\
  & \qquad =\;
    \mathrm{pmfBehavioralKernelGameAt}(g).\mathrm{outcomeKernel}(\beta).
\end{aligned}
\]
This is the intended notion of ``translated to the expected EFG'': expected
up to the canonical FOSG-to-EFG serialisation, with explicit checked tree
shape at the root.

\subsection{StrategicSemantics}

The file \TT{Examples/StrategicSemantics.lean} packages the strategic side of
the worked examples. It first proves three generic theorems for any checked
program $g$: the public-game and blocked-trace-game expected utilities are
equal (\TT{pure_eu_eq_blockedTrace}\label{def:pure_eu_eq_blockedTrace} and \TT{behavioral_eu_eq_blockedTrace}\label{def:behavioral_eu_eq_blockedTrace})
and Nash and dominance predicates transport across the projection
(\TT{pure_nash_iff_blockedTrace}\label{def:pure_nash_iff_blockedTrace}, \TT{behavioral_nash_iff_blockedTrace}\label{def:behavioral_nash_iff_blockedTrace},
\TT{pure_dominant_iff_blockedTrace}\label{def:pure_dominant_iff_blockedTrace}), and exposes a generic legal pure
strategy \TT{defaultPureStrategy}\label{def:defaultPureStrategy} witnessing that the strategy carrier is
inhabited. It then specialises these theorems to each example
(Prisoners, Matching Pennies, Battle of the Sexes, Monty Hall,
SyntaxConstructors), and for Prisoners additionally spells out the explicit
event-graph nodes and field assignments — \TT{rowCommitNode}\label{def:rowCommitNode},
\TT{rowDefectSlice}\label{def:rowDefectSlice}, \TT{rowDefectAction}\label{def:rowDefectAction} — and exercises \TT{sliceLegal} and
\TT{actionLegal} directly. It is the recommended starting point for a reader
who wants to see \emph{how} the source elaborator maps individual constructors
to event-graph data, and \emph{how} the public-game and blocked-trace-game views
relate at the strategic API.

% =====================================================================
\section{Reading order}

For a reader landing on the codebase for the first time, we suggest:
\begin{enumerate}[nosep]
\item Read this supplement's \S\ref{sec:birdseye} and the main report's
  abstract.
\item Skim \TT{Vegas/Lang/Basic.lean} for the surface syntax and its lowering
  into \TT{VegasCore}; the current surface adds simultaneous yield phases and
  nullable yields, but no handlers.
\item Read \TT{Vegas/Base/Basic.lean} and \TT{Vegas/Core/Basic.lean}, focusing
  on \TT{IExpr}, the \TT{VCtx} family, and \TT{VegasCore}.
\item Skim \TT{Vegas/Core/WF.lean}, \TT{Vegas/Core/Scope.lean},
  \TT{Vegas/Core/Program.lean}
  for the well-formedness story.
\item Read \TT{Vegas/EventGraph/Basic.lean} top to bottom: the shape of
  \TT{EventGraph} and \TT{Configuration}, the diamond and stability
  theorems.
\item Skim \TT{Vegas/Machine/Basic.lean} and
  \TT{Vegas/EventGraph/ToMachine.lean} to see the abstract carrier and its
  event-graph instance.
\item Read \TT{Vegas/GameBridge/FOSG/Basic.lean} for the FOSG view, then
  \TT{Vegas/GameBridge/FOSG/FromEventGraph.lean},
  \TT{Vegas/GameBridge/EFG/FromEventGraph.lean},
  \TT{Vegas/Strategic/KernelGame.lean}, \TT{Vegas/Strategic/Kuhn.lean}, and one
  of the \TT{Vegas/Examples/} files.
\item Browse \TT{Vegas/Strategic/Equilibrium.lean}, \TT{Vegas/Strategic/Pure.lean},
  \TT{Vegas/Strategic/Properties.lean} for the public solution-concept names.
\end{enumerate}

\noindent The corollary files in \TT{Vegas/Corollaries/} transport
generic \TT{GameTheory}\label{def:GameTheory} results through the Vegas-facing wrappers and are
useful as catalogues, not as introductions.

\end{document}
