\section{Notation and glossary}
\label{sec:glossary}

This report explains a mechanized semantics for a finite probabilistic
game-description language.  It is meant to be readable by someone who
can read ordinary mathematical notation and programs, but who may not
already know Lean~\citep{deMouraUllrich2021}, extensive-form game
theory~\citep{vonNeumannMorgenstern1944,OsborneRubinstein1994,
ShohamLeytonBrown2009}, refinement proofs, or programming-language
semantics.  The recurring question is practical: given a protocol
program, what strategic game does it define, and what must an
implementation prove in order to preserve that game?

\paragraph{Math first, Lean on the side.}
The body of the report uses ordinary mathematical notation.  Lean
identifiers appear only when their type or naming choice is itself the
subject of a design discussion.  Each section ends with a brief
\emph{Lean correspondence} note pointing to the relevant module paths and
public theorem names.  A reader who wants the story can read the report
straight through; a reader auditing the code can follow those notes back
into the Lean sources.

\paragraph{Terminology.}
Several words are used with their technical meanings.  A
\emph{strategy} is a player's rule for what to do when that player
reaches an observation.  A \emph{profile} is one strategy for every
player.  A \emph{payoff law} is the probability distribution over final
payoff vectors induced by a profile.  An \emph{adequacy theorem} says
that two presentations produce the same payoff law, so analysis can move
between them without changing the game.

\paragraph{Glossary.}
The following terms are used pervasively.  Readers from one
discipline can usually skip the entries that match their background.

\begin{description}[itemsep=0pt,leftmargin=1.2em,style=nextline]
\item[Finite-support distribution / probability mass function (PMF).]
  A function $\mu : \alpha \to [0,1]$ with finite support and
  $\sum_x \mu(x) = 1$.  We write $\PMF{\alpha}$ for the set of such
  functions.  $\PMF{\cdot}$ is a monad: the bind operator
  $\mu \bind f = \sum_x \mu(x)\,f(x)$ chains a stochastic step with a
  continuation indexed by its result, and the unit $\delta_x$ is the
  Dirac point mass at $x$.
\item[Finite-weight function.]
  An auxiliary unnormalized version, $\FWeight{\alpha} = \alpha \to_0
  \NNQ$, used internally during construction; converting to a PMF
  requires a side proof of total mass $1$.
\item[Kuhn's theorem (finite realization and deviation form).]
  A way of saying that ``randomize once over a complete plan'' and
  ``randomize locally whenever a decision point is reached'' are
  equivalent when players do not forget what they previously knew
  or did~\citep{Kuhn1953}.  In VegasCore the theorem is stronger than
  equality of final payoffs: one-player deviations can be matched in the
  directions needed for equilibrium transport, including around the
  canonical mixed-to-behavioral realization (\Cref{sec:pres:kuhn}).
\item[Factored-observation stochastic game (FOSG).]
  A state-based presentation of finite games with hidden information
  \citep{Kovarik2022}.  In each round, active players choose actions
  from their local menus, the state changes, and each player receives an
  observation.  VegasCore exports FOSGs so its frontier games can
  be compared with a standard imperfect-information presentation.
\item[Extensive-form game (EFG).]
  The classical tree presentation, with chance and decision nodes,
  per-player information sets, and outcomes at the leaves.
\item[Frontier.]
  The current decision boundary of the event graph: graph work whose
  prerequisites are complete but whose protocol event has not yet been
  executed.  Frontier round views close forced internal work and expose
  the ready player commitments at this boundary.
\item[Checkpoint.]
  A game-facing observation point at which strategies are allowed to
  react.  A checkpoint is not a claim that a concrete runtime literally
  pauses there; it is the abstraction level at which frontier histories
  record observations and choices.
\item[Kernel game.]
  The smallest game object VegasCore needs for strategic analysis: for
  each player a strategy type, an outcome type, a probability law from
  strategy profiles to outcomes, and a utility function
  $u : \Outcome \to \Players \to \mathbb{R}$.  The strategies may be
  contingent rules over the frontier observations described in
  \Cref{sec:fosg-strategies}, not just one-shot actions.  The word
  \emph{kernel} here means a stochastic map from strategy profiles to
  PMFs over outcomes, not a null-space.
\item[Nash equilibrium.]
  A profile where no single player can improve their expected utility by
  changing only their own strategy while the others stay
  fixed~\citep{Nash1951,OsborneRubinstein1994}.  The refinement results
  in this report preserve selected Nash equilibria only for
  implementation strategy spaces that match, project to, or are
  explicitly related to the frontier strategy space.
\item[Cheap talk / babbling.]
  \emph{Cheap talk} is a payoff-free communication channel that players
  may use before or during strategic play.  It is payoff-free, but not
  automatically strategically irrelevant: players may coordinate through
  the messages.  A \emph{babbling} profile is the special case where
  players send fixed default messages and ignore all received messages.
  The standard theorem is one-way: a base Nash equilibrium embeds as a
  babbling equilibrium of the cheap-talk game, but cheap talk may create
  additional non-babbling equilibria.
\item[Static-single-assignment (SSA).]
  A standard programming-language convention~\citep{cytron1991efficiently}
  in which every variable is assigned at exactly one program point.  A
  program in SSA form has a unique definition site for each name; we
  exploit this to make every event node carry a unique write target.
\item[Event structure / event graph.]
  In concurrency theory~\citep{NielsenPlotkinWinskel1981,Winskel1986EventStructures}
  an event structure abstracts a process by its events and a causal
  dependency relation; independent events are unordered and may be
  permuted without affecting the resulting state.  Our event graphs
  are a typed, finite, asymmetric variant adapted to the protocol
  setting.
\item[Refinement / forward simulation.]
  A proof that a lower-level implementation behaves like a higher-level
  specification after hiding implementation-only detail.  Concretely,
  one supplies a projection from implementation states, events,
  observations, and outcomes to specification states, events,
  observations, and outcomes, and proves that implementation steps
  project to specification steps with the same probability laws
  \citep{LynchVaandrager1995,AbadiLamport1991}.  Bisimulation and
  probabilistic transition-system accounts provide neighboring
  formulations~\citep{Park1981Bisimulation,Milner1989Bisimulation,
  Segala1995Probabilistic}.  For a contract backend, the implementation
  might be a chain or contract-execution machine, and the specification
  would be the primitive event-graph machine.
\item[Visibility / view.]
  The information-set structure of a multi-party protocol, presented
  here as a binding-level discipline: every variable is either public
  or hidden to a single owner, and every program point exposes a
  per-player projected context.
\end{description}
