\section{Introduction}
\label{sec:intro}

Mechanisms with real stakes increasingly run on public, shared runtimes.
Auctions, prediction markets, votes, and escrows are deployed as smart
contracts, where the code is open, every execution step is visible, and the
other participants are strategic and often adversarial.  A mechanism deployed
this way is two things at once.  It is a \emph{program}, whose correctness one
would like to establish with the tools of programming-language semantics; and
it is a \emph{game}, whose outcome turns on what rational, partially-informed
players will do, which is the province of game theory.  Today the two readings
live apart: a contract is written and verified as code, a game model is
analysed separately for its incentives, and the correspondence between them is
taken on trust.  Nothing guarantees that an equilibrium proved of the model is
an equilibrium of the thing that runs---or that the runtime's own character,
such as the order in which transactions are applied and the visibility of
pending moves, has been folded into the analysis rather than ignored by it.

This paper asks whether a single artifact can carry both readings, and at what
cost.  We study a small language whose programs \emph{are} games of incomplete
information, with one organising aim: a game-theoretic fact proved of a program
should survive, by theorem, down to the runtime that executes it, while that
runtime's scheduling and information structure are treated as part of the game
rather than swept beneath it.  What we report is less a finished platform than
a map---of what such a correspondence demands, which parts of it can be
mechanically checked today, and where the remaining obligations on a concrete
runtime lie.

A VegasCore program is a straight-line description of such a game.  It draws
public randomness, asks players to commit sealed values under a visibility
constraint, reveals committed values to everyone, and reads payoffs off the
public state.  Each value is either public or sealed to a single player;
\(\Reveal\) discloses a sealed value to all.  The control flow is
value-independent, so only the drawn data varies from run to run, never the
shape of what happens.  These few declarations---who draws, who commits under
whose view, what is revealed, and how payoffs read the public state---fix an
information structure and a strategic structure at once.  The program is
therefore both the description of a game of incomplete information and the
artifact one carries forward to execution.

The difficulty is that a program is written and run in sequence, yet its
strategic and informational content must not depend on the incidental order
in which independent events happen to be written down.  If the text lists one
player's sealed commitment before another's, that order must \emph{not} make
the second player able to observe the first: the two commitments are meant to
be simultaneous, independent choices.  Our resolution is to let meaning
factor through the program's \emph{dependency} order rather than its textual
order.  The linear order remains a sound way to reason about what is
computed---data flows forward, as one expects---but it is provably inert for
who-knows-what and who-chooses, except through the dependencies the program
explicitly declares.  Independent events may be reordered with no strategic
effect; dependent ones may not.  The order is not meaningless; it is
meaningful precisely up to the dependency partial order.

It is worth stating what this does \emph{not} license.  The semantics does
not validate the naive sequential reading ``\(B\) is written after \(A\), so
\(B\) sees \(A\).''  At a simultaneous-move phase that reading is simply
false.  The value to the writer is exactly the converse: textual order cannot
create an unintended information edge.  Order is safe for computation because
it has been made inert for information.

Three pieces of machinery deliver this guarantee.  A visibility-aware event
graph records the program's dependency structure.  A commit/reveal barrier
enforces simultaneity: no reveal becomes available until every preceding
commitment is sealed.  And schedule confluence ensures that reordering
independent events yields the same outcome and the same per-player
observation.  This is why meaning is factored through a dependency graph
rather than emitted directly as an extensive-form game tree: the graph is
load-bearing but instrumental, the mechanism rather than the headline.

The contribution of this paper is the machine-checked surface that connects a
program to the strategic games and runtime models it determines.  We organize
it as a small family of \emph{faithfulness relations} of three shapes---%
equivalences, an adequacy, and a refinement order---set out in
\Cref{sec:intro:relations}.  The remainder of this section argues the
dependency view against ordinary game trees, works a small Matching-Pennies
example, sketches the pipeline from program to games and runtimes, lays out
these relations, places the runtime backends, and summarizes the theorem
surface; \Cref{fig:overview} gives the roadmap.  Unfamiliar terms ---
FOSG, EFG, frontier, kernel game, adequacy --- are collected in the
glossary (\Cref{sec:glossary}).

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  >={Stealth[length=2.4mm]},
  font=\small,
  box/.style={draw, rounded corners, align=center, inner sep=4pt,
              minimum height=9mm, text width=46mm},
  core/.style={box, fill=blue!8, very thick},
  side/.style={box, fill=green!8, text width=42mm},
  goal/.style={box, dashed, fill=black!4, text width=46mm},
  flow/.style={->, thick},
  lab/.style={font=\footnotesize, midway, right=1mm, align=left},
]
% ── forward semantic spine (top-down) ────────────────────────────────
\node[box]  (prog)  at (0,0)    {Protocol program\\{\footnotesize public samples \(\cdot\) sealed commits \(\cdot\) reveals \(\cdot\) payoffs}};
\node[core] (graph) at (0,-2.1) {Visibility-aware \textbf{event graph}\\{\footnotesize the program's dependency structure}};
\node[box]  (mach)  at (0,-4.2) {Operational machine\\{\footnotesize schedule-agnostic execution}};
\draw[flow] (prog)  -- node[lab]{compile\\{\scriptsize source syntax erased}} (graph);
\draw[flow] (graph) -- node[lab]{execute} (mach);
% ── strategic analysis surface (to the right) ────────────────────────
\node[side] (games) at (7.2,-1.6) {Strategic games\\{\footnotesize pure \(\cdot\) behavioral \(\cdot\) mixed-pure}};
\node[side] (pres)  at (7.2,-3.6) {FOSG / EFG presentations};
\draw[flow] (mach.east) .. controls +(2.2,0) and +(-1.6,0) .. (games.west)
  node[pos=0.55,above,font=\footnotesize,align=center]{frontier\\view};
\draw[flow] (games) -- (pres);
\node[draw, rounded corners, fill=green!4, inner sep=4pt,
      align=left, text width=44mm, font=\footnotesize] at (7.2,0.5)
  {\textbf{Prove \& analyze:} payoff adequacy, visibility, Kuhn equivalence,
   Nash / equilibria};
% ── runtime tower (bottom-up refinement) ─────────────────────────────
\node[box]  (back) at (0,-6.4) {Abstract backend layers\\{\footnotesize audited steps \(\cdot\) encoded storage \(\cdot\) message buffers}};
\node[goal] (evm)  at (0,-8.6) {\textbf{EVM / smart-contract runtime}\\{\footnotesize the setting that grounds the design}};
\draw[flow] (back) -- node[lab]{refines\\{\scriptsize utility-preserving}} (mach);
\draw[flow] (evm)  -- node[lab]{refines\\{\scriptsize layers compose into a tower}} (back);
% ── legend ───────────────────────────────────────────────────────────
\node[draw, rounded corners, align=left, font=\scriptsize, fill=white,
      inner sep=3pt, anchor=north west] at (4.6,-6.0)
  {\tikz\node[draw,rounded corners,minimum width=4mm,minimum height=3mm,fill=blue!8]{};\ mechanized in this work\\
   \tikz\node[draw,rounded corners,dashed,minimum width=4mm,minimum height=3mm,fill=black!4]{};\ motivating target};
\end{tikzpicture}
\caption{VegasCore at a glance.  Solid boxes are mechanized in this
development.  The downward arrows are the forward semantics that turn a
protocol into a game; the strategic surface on the right is where
game-theoretic properties are proved.  Concrete runtimes enter from below as a
\emph{tower} of utility-preserving refinements that compose down to the
primitive machine; the EVM runtime (dashed) is the concrete setting whose
characteristics shaped the design, not a backend mechanized here.}
\label{fig:overview}
\end{figure}

\subsection{Protocols vs. game trees}

Many finite games are naturally written as protocols.  Each player
commits to a private value, a public sample is drawn, sealed values are
revealed, and the terminal public state determines the payoffs.  Writing
the same object directly as an
extensive-form tree forces an arbitrary linearization of events that are
intended to be simultaneous or independent.  The tree can repair that
choice with information sets, but the independence structure is then
spread across syntax and equivalence classes.

A program is checked and compiled to a finite event graph.  The graph records which
fields each event reads, which field each event writes, and which earlier
events are prerequisites because they provide information needed at the
event.  The graph state records only completed nodes and the typed store.
It does not record the order in which independent nodes happened to
complete.

The result is a schedule-agnostic operational semantics.  Primitive
execution steps one available event at a time, but the scheduler is an
external event law.  The strategic semantics use a frontier round view
(a view of the currently ready graph boundary): players see checkpoint
observations and choose for ready commit nodes; internal graph work is
closed by the checkpoint policy.  This separates the schedule-agnostic
game events of the language from implementation-level linearization
detail.

\subsection{A running example}
\label{sec:intro:matching-pennies}

Consider a Matching-Pennies protocol.  The matcher and mismatcher each
commit a sealed boolean, then both booleans are revealed.  The matcher
gets \(+1\) on equality and \(-1\) otherwise; the mismatcher gets the
opposite payoff.  A source program necessarily lists the constructors in
some order, for example:
\begin{quote}\small
\begin{tabular}{@{}l@{}}
\(\Commit_{\mathsf{m}}\,x_{\mathsf{m}} \mid \mathsf{true};\)\\
\(\Commit_{\mathsf{m'}}\,x_{\mathsf{m'}} \mid \mathsf{true};\)\\
\(\Reveal\,y_{\mathsf{m}} := x_{\mathsf{m}};\)\\
\(\Reveal\,y_{\mathsf{m'}} := x_{\mathsf{m'}};\)\\
\(\Ret\,[\mathsf{m}\mapsto u(y_{\mathsf{m}},y_{\mathsf{m'}}),
          \mathsf{m'}\mapsto -u(y_{\mathsf{m}},y_{\mathsf{m'}})].\)
\end{tabular}
\end{quote}
The intended game is not sequential at the commitment phase.  The two
commits are independent choices because neither player can observe the
other player's sealed field before committing.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  every node/.style={font=\small},
  ev/.style={draw, rounded corners, inner sep=2pt, minimum width=1.8cm,
             minimum height=4mm, fill=blue!6},
  arr/.style={-{Stealth[length=2mm]}, thick},
  barr/.style={-{Stealth[length=2mm]}, thick, densely dashed, gray},
]
\node[ev] (cm) at (-2,1) {\(\Commit_{\mathsf{m}}\)};
\node[ev] (cmp) at (2,1) {\(\Commit_{\mathsf{m'}}\)};
\node[ev] (rm) at (-2,-1) {\(\Reveal_{\mathsf{m}}\)};
\node[ev] (rmp) at (2,-1) {\(\Reveal_{\mathsf{m'}}\)};
\node[ev] (ret) at (0,-2.5) {\(\Ret\)};
\draw[arr] (cm) -- (rm);
\draw[arr] (cmp) -- (rmp);
\draw[barr] (cm) -- (rmp);
\draw[barr] (cmp) -- (rm);
\draw[arr] (rm) -- (ret);
\draw[arr] (rmp) -- (ret);
\end{tikzpicture}

\caption{Matching Pennies as an event graph.  Solid edges --- from a
commit to its own reveal, and from reveals to \(\Ret\) --- are data
dependencies that carry a sealed or public value.  Dashed cross-edges
from one player's commit to the other player's reveal are barrier
dependencies: no reveal fires until \emph{both} commitments are sealed,
and they carry no payload.  The two commits themselves are unordered ---
that simultaneity is the point.}
\label{fig:matching-pennies-eg}
\end{figure}

In \Cref{fig:matching-pennies-eg}, the two commit nodes share the
initial frontier, the graph boundary of currently ready events.  Each
reveal depends on its own sealed source field, and the reveal phase also
waits on the \emph{other} commit.  This second dependency is the
commit/reveal barrier: no reveal becomes available until both
commitments are sealed, so that no player can condition a commitment on
another's revealed value.  The barrier is graph structure, not a
strategy-side convention.  Thus the source text may list one commitment
before the other, but the compiled game still presents them as
simultaneous sealed choices.

\subsection{The semantic spine}
\label{sec:intro:spine}

The construction has the following shape:
\[
\begin{aligned}
  \mathsf{VegasLang}
  &\xrightarrow{\;\mathsf{lower}\;}
  \mathsf{VegasCore}
  \xrightarrow{\;\mathsf{compile}\;}
  \mathsf{EventGraph} \\
  &\xrightarrow{\;\mathsf{primitiveMachine}\;}
  \mathsf{Machine}
  \xrightarrow{\;\mathsf{frontierRoundView}\;}
  \mathsf{KernelGame}.
\end{aligned}
\]
\(\mathsf{VegasLang}\) is a surface layer over the concrete expression
language.  It contains administrative \(\Let\)-bindings and nullable
yield sugar.  Lowering erases administrative lets by substitution and
turns yields into a sealed optional commit followed by a reveal.  The
maintained core, \(\mathsf{VegasCore}\), contains only protocol
constructors: \(\Sample\), \(\Commit\), \(\Reveal\), and \(\Ret\).

Compilation produces a well-formed event graph together with terminal
payoff projections.  It needs only basic static hygiene, not the stronger
checked-game obligations --- reveal completeness and legal commit guards
--- that the later progress and game-facing theorems consume.  The event
graph is source-independent: it records protocol structure rather than
syntax, so two programs differing only by \(\alpha\)-renaming compile to
equal graphs (names are erased to positional field ids).  Programs that
differ by reordering \emph{independent} effects compile to graphs that are
not equal but equivalent up to schedule confluence
(\Cref{sec:eg:diamond}); the graph data is order-sensitive, while the
denotation it carries is not.

The final arrow to \(\mathsf{KernelGame}\) --- the frontier game named in
the figures and in \Cref{sec:intro:relations} --- is where strategic
analysis starts.  A kernel game does not remember the syntax of the protocol; it
remembers the strategy spaces, the probability law over outcomes, and
the utility function.  The source-payoff adequacy theorems prove that
the outcomes used there are exactly the source payoffs evaluated at
terminal graph states.

\subsection{A map of the guarantees: five relations, three shapes}
\label{sec:intro:relations}

The semantic spine of \Cref{sec:intro:spine} lists the \emph{objects}; the
development is easier to hold in one's head through the \emph{relations} that
connect them.  There are five.  Separating them by \emph{algebraic shape}
--- an equivalence, an adequacy, or an order --- makes each one answer a
single question: \emph{what may I forget?}  Each is a separate theorem with
its own proof technique, so naming them apart, rather than lumping them as
``equivalence results,'' turns the end-to-end guarantee into a short
composite instead of a monolith.

\begin{description}[itemsep=2pt,leftmargin=1.3em,style=nextline]
\item[Schedule confluence \quad{\small(equivalence, inside one graph)}.]
  Two completions of the same graph in different orders reach the same
  terminal state, and batch-independent events commute
  (\Cref{thm:diamond,thm:barrier}).  \emph{Forget:} the order in which
  independent events were scheduled.
\item[Source-payoff adequacy \quad{\small(adequacy, across \texttt{compile})}.]
  The compiled graph \emph{is} the program's denotation: terminal
  outcomes are the source payoffs reconstructed from the terminal store
  (\Cref{sec:eg:adequacy}).  Adequacy here is the one-directional sense
  ``the graph means the program, by theorem'' --- not the
  operational/denotational coincidence of the same name --- and in
  particular not a symmetric equivalence between arbitrary programs and
  graphs.  \emph{Forget:} the source syntax.
\item[Presentation equivalence \quad{\small(equivalence, across a presentation)}.]
  The FOSG (factored-observation stochastic game) and EFG (extensive-form
  game) presentations carry the same utility distribution as the native
  frontier game (\Cref{thm:fosg,thm:efg}).  \emph{Forget:} which game-form
  a result was stated in.
\item[Strategic (Kuhn) equivalence \quad{\small(equivalence, inside the strategic layer)}.]
  Behavioral and product mixed-pure frontier play agree on outcome laws
  \emph{and} match one-player deviations, so they transport solution
  concepts, not merely payoffs
  (\Cref{thm:kuhn-beh-to-mixed,thm:kuhn-mixed-to-beh,thm:kuhn-strategic-equivalence}).%
  \statusnote{The behavioral\,$\leftrightarrow$\,mixed-pure pair is a full
  two-way equivalence with deviation matching; the correlated-pure
  direction is mechanized only as a one-way outcome-law realization.}
  \emph{Forget:} which strategy carrier a player reasons with.
\item[Runtime refinement \quad{\small(order, across an implementation)}.]
  An implementation machine refines the primitive machine by projection,
  and --- given an adequate trace law --- preserves the frontier game's
  utilities (\Cref{thm:runtime-trace}).  Unlike the others this one is an
  \emph{order}: a richer concrete strategy space can carry equilibria the
  specification does not, so the guarantee is directional in general.%
  \statusnote{At a fixed (frontier) strategy space the pull-back is in
  fact two-way under bounded utility: the implementation trace game
  indexed by the frontier strategy type has \emph{exactly} the frontier
  Nash set, so the backend introduces no spurious equilibria.  The order's
  asymmetry is genuine only across strategy-space \emph{enlargement} --- a
  backend whose concrete strategies have extra coordinates (e.g.\ an
  out-of-band channel) can add equilibria; that is the cheap-talk case of
  \Cref{sec:ref:cheap-talk}.}
  \emph{Forget:} the runtime's administrative bookkeeping.
\end{description}

\Cref{fig:relations} draws these as labelled edges on the object spine.
Reading the runtime question end-to-end --- ``why does a theorem about a
backend say anything about my program?'' --- is exactly a walk through the
composite: runtime refinement lands on the frontier game; strategic and
presentation equivalence say that game is carrier- and
presentation-independent; schedule confluence says the graph underneath is
schedule-independent; and source-payoff adequacy says the graph is the
program.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  >={Stealth[length=2.2mm]},
  font=\small,
  obj/.style={draw, rounded corners, align=center, inner sep=3pt,
              minimum height=8mm, text width=30mm, fill=blue!8},
  side/.style={obj, fill=green!8, text width=30mm},
  impl/.style={obj, fill=black!4},
  rel/.style={->, thick},
  rlab/.style={font=\scriptsize, align=center},
]
% spine
\node[obj]  (prog)  at (0,0)    {Protocol program};
\node[obj]  (graph) at (0,-1.9)  {Event graph};
\node[obj]  (mach)  at (0,-3.8)  {Primitive machine};
\node[side] (game)  at (5.6,-1.9){Frontier game};
\node[side] (pres)  at (5.6,-3.8){FOSG / EFG};
\node[impl] (impl)  at (0,-5.7)  {Impl.\ machine};
% R2 adequacy
\draw[rel] (prog) -- node[rlab,right=1mm]{$\adeq$\\{\scriptsize adequacy}} (graph);
% R1 confluence: self relation on the graph (an internal equivalence)
\draw[rel,<->] (graph.west) .. controls +(-2.1,0.8) and +(-2.1,-0.8) .. (graph.west)
  node[rlab,left=1mm,pos=0.5]{$\eqrel$\\{\scriptsize confluence}};
% execute
\draw[rel] (graph) -- node[rlab,right=1mm]{execute} (mach);
% frontier view
\draw[rel] (mach.east) .. controls +(2.4,0) and +(-1.6,0) .. (game.west)
  node[rlab,pos=0.5,above]{frontier\\view};
% R4 Kuhn: self relation on the frontier game (an internal equivalence,
% drawn like R1 so its shape on the page matches its structure)
\draw[rel,<->] (game.north west) .. controls +(-0.3,1.0) and +(0.3,1.0) .. (game.north east)
  node[rlab,above,pos=0.5]{$\eqrel$ Kuhn equivalence\\{\scriptsize beh.\,$\eqrel$\,mixed-pure}};
% R3 presentation
\draw[rel] (game) -- node[rlab,right=1mm]{$\eqrel$\\{\scriptsize presentation}} (pres);
% R5 refinement (order, upward)
\draw[rel] (impl) -- node[rlab,right=1mm]{$\ordrel$\\{\scriptsize refinement}} (mach);
% legend
\node[draw, rounded corners, align=left, font=\scriptsize, fill=white,
      inner sep=4pt, anchor=north west] at (3.0,-5.2)
  {$\eqrel$~equivalence \;(symmetric)\\
   $\adeq$~adequacy \;(one object \emph{is} the other's meaning)\\
   $\ordrel$~refinement \;(an order; directional facts)};
\end{tikzpicture}
\caption{The five faithfulness relations, by shape.  Equivalences
($\eqrel$) let a result move freely; adequacy ($\adeq$) is the
one-directional bridge from source syntax into the source-free graph; the
refinement order ($\ordrel$) carries directional guarantees up from a
concrete backend.  The end-to-end claim ``a backend preserves my
program's game'' is their composite, read bottom-up.}
\label{fig:relations}
\end{figure}

\subsection{Where runtime backends fit}
\label{sec:intro:runtime-backends}

Concrete runtimes enter below the primitive event-graph machine.  A
smart-contract runtime, a payment-channel state machine, or a local
simulator is modeled as a lower-level machine: its states contain
runtime storage and administrative data, its events model the relevant
calls or operational steps, and its observations say what each player can
infer from that runtime.  To connect such a machine to VegasCore, the
backend proves a stochastic refinement to the primitive event-graph
machine and supplies a trace law showing that the implementation realizes
the selected frontier game surface.  Ethereum/EVM is the concrete setting
this pattern is designed around: its public mempool, transaction ordering,
and execution costs are what motivate the visibility discipline, the
schedule-agnostic graph, and the refinement interface in the first place.  A
verified EVM backend is not among the results here, but the abstract
message-in-flight backend below already realizes the EVM's strategically
distinctive feature --- a public mempool whose pending transactions are
visible before they are applied.

This is intentionally stronger than saying that a contract ``looks like''
the protocol on examples: the refinement proof identifies which
implementation details are erased, which events correspond to graph
events, and why the final utilities are unchanged.  Extra communication
requires a separate strategy-space argument; \Cref{sec:ref:cheap-talk}
gives the three sound positions.

The backend layer is a tower-building interface: stochastic refinements
compose, event-batch law families compose with them, and abstract backend
examples demonstrate audited administrative steps, encoded storage, and
explicit message buffers.  The message-in-flight wrapper is deliberately
player-visible, so it can model settings in which strategies condition
on pending messages.  When those messages are irrelevant,
inert-extension theorems transport equilibria; when they are
strategically meaningful, the specification game includes the
corresponding protocol messages or cheap-talk channel.

\subsection{Contributions and theorem surface}
\label{sec:intro:contributions}

The main contribution is the checked semantic path from protocol syntax
to game presentations and runtime-refinement obligations.  The public
theorem surface is organized around the following claims.
\begin{enumerate}[leftmargin=1.4em,itemsep=2pt]
\item \textbf{Compilation.}  Checked programs compile to well-formed,
  source-free event graphs.  Payoff projections read only public fields
  that are available by graph termination.
\item \textbf{Progress.}  Source guard legality compiles to graph
  \(\mathsf{GuardLive}\), and graph guard liveness gives progress for
  the primitive downset checkpoint model.
\item \textbf{Outcome adequacy.}  Terminal primitive-machine outcomes
  agree with the source payoff expressions reconstructed from the
  terminal compiled store.
\item \textbf{Visibility.}  Commit reads are exactly the committing
  player's visible dependency footprint; public samples and payoff
  projections read declared public dependencies; reveals are the only
  sealed-to-public flow.
\item \textbf{Frontier semantics.}  Pure, behavioral, and product
  mixed-pure frontier games are defined from the canonical frontier
  round view.  Their outcome kernels are source-payoff projections of
  the same canonical run distribution.  A source-native behavioral
  strategy --- a function of the player's source-local history view ---
  and the source checkpoint game it induces are also mechanized: that
  game shares the frontier outcome law and is two-way Nash-deviation
  bisimilar to the frontier game on the checkpoint information surface,
  where source and frontier strategies coincide.  Carrying this to the
  raw source-local strategy surface is reduced to one bridge lemma,
  stated but not yet discharged.%
  \statusnote{So ``my program's equilibria are the frontier game's
  equilibria'' is, on the checkpoint surface, a theorem --- a two-way
  bisimulation --- not merely the definition of strategic meaning; only
  the raw source-local bridge remains assumed.}
\item \textbf{FOSG and EFG presentations.}  The FOSG and EFG layers are
  payoff-faithful presentations of the frontier semantics.
\item \textbf{Kuhn.}  Behavioral and product mixed-pure frontier profiles
  are related by a strategy/deviation equivalence at the canonical
  round-view information model: outcome laws agree, and one-player
  deviations can be matched across the two presentations.
\item \textbf{Runtime refinement.}  A stochastic refinement plus an
  adequate strategy-indexed trace law preserves the utility distribution
  of the selected frontier game surface.  Refinements and law-family
  lifts compose, so a tower of abstract backend layers can be collapsed
  to one proof obligation against the primitive machine.  With the
  strategy-space conditions of \Cref{sec:ref:cheap-talk} and bounded
  utility for the compared trace utilities, the implementation trace game
  indexed by the frontier strategy space has \emph{exactly} the frontier
  Nash set --- the backend adds no spurious equilibria.  The remaining
  asymmetry is across strategy-space enlargement: a concrete machine with
  additional strategic coordinates, such as an out-of-band channel, can
  carry equilibria the specification does not, unless those coordinates
  are ruled out, proved inert, or modeled explicitly.
\end{enumerate}

\leancorr{The end-to-end, program-facing claims of this list are collected,
with their fully qualified names, in the build-checked index
\TT{Vegas.Theorems.Claims}.  A few load-bearing supporting results --- the
visibility facts of item~4 and the schedule diamond and barrier
(\Cref{thm:diamond,thm:barrier}) --- are proved in their own theorem modules
(\TT{Vegas.Theorems.Visibility}, \TT{Vegas.Theorems.Execution}) rather than
re-exported through the index; the per-section \emph{Lean correspondence} notes
point to the module where each result is established.
The construction modules are \TT{Vegas.Language}, \TT{Vegas.Core},
\TT{Vegas.EventGraph}, \TT{Vegas.Frontier},
\TT{Vegas.Presentation}, \TT{Vegas.Export}, and \TT{Vegas.Machine}.}
