\section{Mathematical preliminaries}
\label{sec:preliminaries}

We collect the ambient probabilistic and language-level notation.  All
operational data that must be enumerated is finite: player sets, action
domains, stored value domains, programs, event graphs, reachable states,
and reachable strategy carriers.  Terminal payoff expressions return
integers, but those integers are not treated as an enumerable runtime
domain.  Distributions are finite-support PMFs; continuous types and
measure-theoretic kernels are out of scope (\Cref{sec:discussion}).

\subsection{Finite-support distributions}
\label{sec:prelim:pmf}

\begin{definition}[PMF]
\label{def:pmf}
For a type $\alpha$ with decidable equality, a \emph{probability mass
function} is a finite-support function
$\mu : \alpha \to [0,1]$ with $\sum_{x \in \mathrm{supp}(\mu)} \mu(x)
= 1$.  We write $\PMF{\alpha}$ for the set of such functions.  Because the
masses sum to one, the support is nonempty: every PMF offers at least one
value.  This is why a $\Sample$ step (\Cref{sec:source:sos}), which ranges
over the support of a PMF, is always enabled --- there is no stuck
nonterminal configuration arising from an empty draw.
\end{definition}

The PMF construction is a \emph{commutative monad}~\citep{Giry1982,Moggi1991}:
\begin{align*}
  \mathsf{pure}(x)    &= \delta_x,\\
  (\mu \bind f)(b)    &= {\textstyle\sum_{a \in \alpha}}\,\mu(a)\cdot f(a)(b),\\
  \mathrm{map}(g)\,\mu &= \mu \bind (\mathsf{pure}\circ g).
\end{align*}
Commutativity (i.e.\ that $\mu_1 \bind \lambda x.\mu_2 \bind \lambda y.f(x,y)
= \mu_2 \bind \lambda y.\mu_1 \bind \lambda x.f(x,y)$) is the
probabilistic version of a familiar idea: if two random steps are
independent, doing the left one first or the right one first gives the
same joint law.  This is one ingredient behind the frontier diamond for
independent graph events.  Finite probabilistic languages and their
denotational semantics provide the surrounding programming-language
background~\citep{Kozen1981,GordonProbabilistic,BorgstromEtAl2016Fun}.

\begin{remark}[Finite weights vs.\ PMFs]
\label{rem:fweight}
Inside the construction we work with an unnormalized auxiliary,
$\FWeight{\alpha} = \alpha \to_0 \NNQ$, the type of finite-support
$\mathbb{Q}_{\ge 0}$-valued weight functions.  Rational weights are
decidably equal, finite weights compose without per-step normalization
proofs, and the bridge $\mathrm{toPMF} : \FWeight{\alpha} \to
\PMF{\alpha}$ takes a single proof of total mass one.  The code uses
finite weights at construction time and PMFs for semantic distributions;
this paper uses PMFs throughout.  $\FWeight{\cdot}$ surfaces only when
the construction itself is being discussed.
\end{remark}

\begin{notation}[Dirac and product]
\label{not:dirac-product}
$\delta_x \in \PMF{\alpha}$ denotes the point mass at $x$, equal to
$\mathsf{pure}(x)$.  For a finite player set $\Players$ and a family
$\{\mu_p : \PMF{S_p}\}_{p \in \Players}$, the \emph{independent
product} $\bigotimes_{p}\mu_p \in \PMF{\prod_p S_p}$ is the joint PMF
$(\bigotimes_p \mu_p)(s) = \prod_p \mu_p(s_p)$ on tuples $s = (s_p)_p$.
We use the
shorthand $\Kbeh, \Kpure, \Kmix$ for the behavioral, pure, and
mixed-extension frontier kernel games introduced in
\Cref{sec:fosg-kernel-games}.
\end{notation}

\subsection{Typed contexts and environments}
\label{sec:prelim:contexts}

A \emph{typed context} associates each bound name with a type drawn
from a small fixed type universe (described in
\Cref{sec:source-language}).  Because the names in a checked program are
unique (they are in SSA form), we read a context \emph{extensionally},
as a finite map
\[
  \Gamma : x \mapsto \tau,
\]
and use map notation for it throughout: $\mathrm{dom}(\Gamma)$ for the
bound names, $\Gamma(x)$ for the binding at $x$, and
$\Gamma[x \mapsto \tau]$ for extension.  Lean represents the same data as
an ordered list of bindings and recovers the map view from uniqueness of
names; the order is a representational artifact and carries no source
meaning.  An \emph{environment} for $\Gamma$ is a dependent assignment
$\rho \in \Env(\Gamma)$ that maps each name to a value of its declared
type.  In Lean, variable lookup carries a proof that the named variable
is present with the requested type; because names are unique, two proofs
that refer to the same binding give the same value.  This is the formal
reason the compiler can talk about field names without depending on
arbitrary proof objects.

We write $\rho \mathrel{\equiv_S} \rho'$ for two environments that
agree on a variable set $S$.  The fundamental soundness law of the
expression language (\Cref{sec:source-language}) is then
\emph{evaluation respects the dependency set}:
$\rho \mathrel{\equiv_{\mathrm{deps}(e)}} \rho'$ implies
$\mathrm{eval}(e, \rho) = \mathrm{eval}(e, \rho')$, and similarly for
distribution expressions.

\subsection{Notation conventions}
\label{sec:prelim:notation}

Throughout the paper, $\Players$ denotes a finite set of players with
decidable equality, with generic player names $p, q$.  We also use:
\begin{itemize}\itemsep1pt
\item $\Outcome = \Players \to_0 \mathbb{Z}$, the type of public
  finitely-supported integer payoff vectors (the notation
  $\alpha \to_0 \beta$ denotes functions $\alpha \to \beta$ with
  finite support, i.e.\ taking the zero of $\beta$ outside a
  finite set).  A $\Ret$ supplies one payoff expression per player
  (\Cref{fig:vegas-syntax}), so over a finite player set every player is
  assigned a value and the zero default is only a representational
  convenience, not a way to leave a player unscored;
\item $u : \Outcome \to (\Players \to \mathbb{R})$ for the canonical
  utility (obtained by coercing each integer payoff to a real);
\item $\Profile$ for a strategy profile and $\Profile_{-p}$ for the
  partial profile of $p$'s opponents;
\item $\EU(\Profile, p)$ for the expected utility of profile
  $\Profile$ to player $p$;
\item $\KG = (\{S_p\}_p, \Outcome, K, u)$ for a kernel game with
  per-player strategy types, outcome type, outcome kernel, and
  utility (\Cref{sec:frontier-strategies});
\item $\bigotimes_p \mu_p$ for an independent product distribution.
\end{itemize}
A checked finite-domain program has finitely many reachable outcomes, and
each carries integer payoffs, so $\EU(\Profile,p)$ is a finite sum of finite
terms: it is always well defined and bounded, with no separate boundedness
hypothesis needed at the frontier level.  Boundedness re-enters only at the
runtime layer (\Cref{sec:refinement}), where implementation \emph{trace}
utilities may fold in unbounded economic costs such as gas or fees; we flag it
explicitly there.
We use sans-serif math symbols for type-level constants
($\Cfg$, $\State$, $\Action$), italic letters for nodes and mathematical
objects, and roman identifiers for predicates and projections.
The bracketed proposition $[\varphi]$ denotes the Iverson bracket: it is
\(1\) when \(\varphi\) holds and \(0\) otherwise.

\leancorr{Probability and finite-weight
definitions are in \TT{Vegas.Foundation.FWeight} and \TT{GameTheory.Math};
typed contexts, environments, and visibility-aware syntax support are in
\TT{Vegas.Foundation.Basic} and \TT{Vegas.Core.Basic}.}
