\section{The source language and visibility discipline}
\label{sec:source-language}

A VegasCore program reads like a probabilistic program in which some of
the sampling is delegated to strategic actors that see only part of the
state: where an ordinary probabilistic program draws a value from a
distribution, a VegasCore program may instead ask a player to choose it.
Read operationally, it is simply a multiparty protocol with executable
parts --- parties seal commitments and later reveal them, nature draws
public randomness, and guards and payoffs are computed --- ending in a
per-player payoff.

\subsection{Expression languages}

The core protocol syntax is parameterized by an expression language
\(L\).  An expression language supplies a type universe \(\Ty\), a
family of semantic values \(\Val(\tau)\), expression and distribution
syntax, evaluation functions, finite dependency sets, and soundness
laws saying that evaluation depends only on the declared dependencies.
The event-graph compiler treats these dependencies as declared read
footprints.  An instance may over-approximate, but it may not omit a
semantic dependency.

The phrase ``read footprint'' is literal.  If an expression says it
depends on fields \(a\) and \(b\), then graph execution, observations,
and proofs are allowed to treat \(a\) and \(b\) as the information that
must be available.  The soundness law prevents a bad expression-language
instance from secretly inspecting a field it did not declare.

The surface language specializes to the simple expression
language used by the examples.  That specialization is deliberate:
nullable yield sugar needs options, default values, and a commitment
payload predicate that are concrete to that expression language.  The
maintained core remains generic in \(L\).

\subsection{Visibility-aware contexts}

A visibility context is the finite metadata used by checking and by
configurations: a typed context (\Cref{sec:prelim:contexts}) in which
each name additionally carries a visibility.  A name is either
\emph{public}, visible to every player, or \emph{sealed} to an owner
\(p\), readable only by \(p\); in the map these are written \(\pub\) and
\(\sealed(p)\).  \(\Commit\) introduces a sealed name, \(\Reveal\)
reclassifies a sealed name as public, and \(\Sample\) introduces a public
one.  The surface syntax leaves visibility implicit
--- it is fixed by the effect --- writing a public variable \(x\) and a
sealed one \(x_p\); the absence of an owner subscript marks a public
variable.  The language is simply typed --- types come from the
expression language \(L\), with boolean guards and integer payoffs ---
but types are orthogonal to the protocol and visibility structure studied
here, so we suppress them throughout.  From a context \(\Gamma\) we use
three projections:
\[
  \erasepub{\Gamma}, \qquad
  \erasectx{\Gamma}, \qquad
  \view{p}.
\]
The first keeps only the public names; the second keeps all of them but
forgets visibility, yielding a plain typed context; and the third keeps
what player \(p\) may see --- the public names together with the names
sealed to \(p\).

\subsection{Core protocol syntax}

A program is a sequence of protocol effects terminated by a payoff
return.  \Cref{fig:vegas-syntax} gives the grammar.  Well-scopedness is
checked against a visibility context: \(\Commit\) and
\(\Sample\) each introduce a fresh name, while \(\Reveal\) discloses an
existing committed one.

\begin{figure}[t]
\centering
\begin{bnf}
$s$ : \textsf{effect} ::=
  | $\Commit~x_p~\Where~\varphi_p$ : sealed commitment, owned by \(p\)
  | $\Reveal~x_p$ : disclose a secret (\(x\) becomes public)
  | $\Sample~x \sim D$ : draw public randomness
;;
$t$ : \textsf{program} ::=
  | $s;~t$ : sequence an effect
  | $\Ret~\overline{e}$ : terminate with payoffs \(\overline{e}\)
\end{bnf}
\caption{Core protocol syntax. A program \(t\) is a sequence of effects
ending in \(\Ret\). The payoff expressions \(\overline{e}\) (one per
player), the distribution \(D\), and the boolean guard \(\varphi_p\) come
from the parameter expression language \(L\); \(x\) is a variable and
\(p\) a player. Types are supplied by \(L\) and suppressed throughout.}
\label{fig:vegas-syntax}
\end{figure}

\(\Commit\) binds a variable \(x_p\) sealed to \(p\) --- player \(p\)'s
sealed choice --- under the guard \(\varphi_p\), evaluated in \(p\)'s
view. \(\Reveal\) reclassifies the sealed \(x_p\) as public, disclosing
that same variable. \(\Sample\) binds a public variable \(x\) and draws
\(D\) over the public part of the context (nature has no private
knowledge). \(\Ret\) terminates with payoff expressions over the public
context.

The corresponding context action is partial.  Writing
\(\mathrm{dom}(\Gamma)\) for the names already bound, \(\Commit~x_p\) and
\(\Sample~x\) require \(x \notin \mathrm{dom}(\Gamma)\), and extend
\(\Gamma\) with \(x\) sealed to \(p\) or public, respectively.
\(\Reveal~x_p\) requires \(x\) already sealed to \(p\), and updates that
same map entry to public.

The guard \(\varphi_p\) is typed over exactly the variables \(p\) can see
at that point: the proposed value together with the entire \(p\)-view of
\(\Gamma\), and nothing outside it.  It is evaluated in an environment
providing exactly those variables --- the player view
\(\rho\!\downarrow_p\) extended with the proposed value \(x \mapsto v\);
which of them the guard actually inspects is up to the expression.  How a
player's view is realised in a concrete execution is a question we do not
settle here.

A guard is the protocol's local admissibility test for a commitment.  It
can depend on the proposed value and on what the committing player is
allowed to see at that point.  It cannot depend on another player's
sealed commitment unless that value has already been revealed.  This is
the source-language counterpart of information sets in classical game
trees, but it is enforced at the binding level instead of by drawing
equivalence classes on tree nodes.

There is no core \(\Let\).  Deterministic administrative bindings live
in the surface layer and are erased by substitution before graph
compilation.  This keeps event graphs about protocol activity rather
than source convenience.  If a deterministic value must be materialized
as public protocol state, it should be expressed through a point-mass
public sample or an explicit future protocol constructor, not through
administrative \(\Let\).

\subsection{Operational semantics}
\label{sec:source:sos}

A program's source-level operational reading is a small-step relation on
\emph{configurations}.  A configuration \(\sconf{\Gamma}{\rho}{t}\) pairs
a visibility map \(\Gamma\), an environment \(\rho\) assigning a value to
every name in \(\mathrm{dom}(\Gamma)\), and the program \(t\) that
remains to run.  The initial configuration of a closed program \(t_0\) is
\(\sconf{\emptyset}{\emptyset}{t_0}\).  Each step runs the program's head
effect; \Cref{fig:vegas-sos} gives the rules.

\begin{figure}[t]
\centering
\begin{mathpar}
\inferrule[Commit]
  {x \notin \mathrm{dom}(\Gamma) \\
   \sem{\varphi_p}_{\rho\downarrow_p,\ x \mapsto v} = \mathsf{true}}
  {\sconf{\Gamma}{\rho}{\Commit\,x_p~\Where~\varphi_p;\ t}
     \;\sstep\;
   \sconf{\Gamma[x \mapsto \sealed(p)]}{\rho[x \mapsto v]}{t}}
\and
\inferrule[Reveal]
  {\Gamma(x) = \sealed(p)}
  {\sconf{\Gamma}{\rho}{\Reveal\,x_p;\ t}
     \;\sstep\;
   \sconf{\Gamma[x \mapsto \pub]}{\rho}{t}}
\and
\inferrule[Sample]
  {x \notin \mathrm{dom}(\Gamma) \\
   v \in \supp\,\sem{D}_{\rho^{\pub}}}
  {\sconf{\Gamma}{\rho}{\Sample\,x \sim D;\ t}
     \;\sstep\;
   \sconf{\Gamma[x \mapsto \pub]}{\rho[x \mapsto v]}{t}}
\end{mathpar}
\caption{Small-step operational semantics.  Here \(\rho^{\pub}\) is the
public part of the environment, \(\rho\!\downarrow_p\) is the environment
restricted to \(p\)'s visible context (the public bindings together with
\(p\)'s own), and \(\sem{\cdot}\) is evaluation in the parameter language
\(L\).  \textsc{Commit} admits any value for which the player's guard
holds in the player's own view; \textsc{Reveal} reclassifies \(x\) from
sealed-to-\(p\) to public, leaving the value in \(\rho\) unchanged;
\textsc{Sample} steps to any value in the support of the distribution.
There is no rule for \(\Ret\): it is terminal.}
\label{fig:vegas-sos}
\end{figure}

Two features set these rules apart from ordinary
environment-threading reduction.  \textsc{Commit} and \textsc{Reveal}
carry the visibility discipline: a commit guard is evaluated in the
committing player's view \(\rho\!\downarrow_p\), and a sealed value
becomes observable only when \textsc{Reveal} reclassifies its map entry
from sealed to public.  \textsc{Sample} is probabilistic: the step
ranges over the support of \(\sem{D}\), recording which draws are
possible, while the quantitative weights are supplied by the denotation
of \(D\).

A terminal configuration \(\sconf{\Gamma}{\rho}{\Ret\,\overline{e}}\)
reports the outcome
\[
  \outcomeof\sconf{\Gamma}{\rho}{\Ret\,\overline{e}}
    \;=\; \sem{\overline{e}}_{\rho^{\pub}},
\]
the payoff vector obtained by evaluating each payoff expression in the
public environment.  Writing \(\sstep^{\!*}\) for the reflexive--%
transitive closure, a complete run of \(t_0\) is a sequence
\(\sconf{\emptyset}{\emptyset}{t_0} \;\sstep^{\!*}\;
  \sconf{\Gamma}{\rho}{\Ret\,\overline{e}}\),
and its result is that terminal outcome.

\begin{remark}[Visibility is static]
\label{rem:visibility-static}
Although the rules thread \(\Gamma\) through the configuration and update
it at each step, visibility is not runtime information.  The core is
straight-line --- a fixed sequence of effects with no value-dependent
control flow --- so the remaining program \(t\) is simply a program
counter into \(t_0\), and the context is a function of that counter
alone.  Writing \(\mathrm{vis}(t)\) for the context determined by the
effects of \(t_0\) preceding \(t\), every reachable configuration
satisfies
\[
  \sconf{\emptyset}{\emptyset}{t_0} \sstep^{\!*}
  \sconf{\Gamma}{\rho}{t}
  \quad\Longrightarrow\quad
  \Gamma = \mathrm{vis}(t),
\]
by induction on \(\sstep^{\!*}\): each effect's action on \(\Gamma\)
depends only on the effect, never on the sampled or committed value in
\(\rho\).  The freshness premise \(x \notin \mathrm{dom}(\Gamma)\) and the
reveal premise \(\Gamma(x) = \sealed(p)\) are then well-scopedness facts
about \(t_0\), not runtime guards: a well-scoped program satisfies them at
every step, so execution never gets stuck on \emph{scoping}.

This is
separate from the commit guard \(\varphi_p\), whose satisfaction is
genuinely value-dependent --- that a proposed value passes the guard is an
admissibility condition, not a scoping fact.  What rules out a stuck commit
is the orthogonal guarantee that \emph{some} admissible value always exists;
that is the legality obligation discharged by \(\mathsf{WFProgram}\)
(\Cref{sec:source:wf}) and carried to the graph as \(\mathsf{GuardLive}\),
not part of this static-visibility invariant.  The configuration carries
\(\Gamma\) only to keep each rule self-contained; the mechanization
instead makes the context a static index of the term, so this invariant
holds by construction and the reveal premise is discharged by an
ownership witness rather than rechecked.  This is the source-level
counterpart of the event graph deriving prerequisites and visibility
\emph{canonically} rather than storing them (\Cref{sec:eg:prereqs}).
\end{remark}

\begin{remark}[Labelled transitions]
Annotating each step with the data it introduces or discloses ---
\(\mathsf{commit}\,x_p\,v\), \(\mathsf{reveal}\,x_p\,v\), or
\(\mathsf{sample}\,x\,v\) --- yields a labelled transition system
\(\sconf{\Gamma}{\rho}{t} \lstep{\ell} \sconf{\Gamma'}{\rho'}{t'}\) whose
label-erasure is the relation of \Cref{fig:vegas-sos}.  A label names the
source action, including a committed value; later epistemic or player
views can be defined as projections of such labels.
\end{remark}

\leancorr{Configurations, the small-step relation, its reflexive--transitive
closure, the terminal outcome reader, and the labelled refinement are
\TT{Vegas.Core.SmallStep}.  The paper presents contexts extensionally as finite
maps in configurations.  Lean uses list-indexed terms and environments as an
intrinsic well-scoped representation: \TT{VegasCore P L \(\Gamma\)} stores the
current context in the type of each continuation.  Under freshness, the list
context represents a finite map up to representation order.  Lean also represents
\(\Reveal\) as a fresh public alias while retaining the sealed source binding;
the paper's in-place reclassification is the source-facing projection of that
representation, and the event-graph context is a separate compiler construction
state related by compilation invariants.}

\subsection{Surface syntax}
\label{sec:source:surface}

\(\mathsf{VegasLang}\) is the surface language over the simple
expression language.  It mirrors the protocol core and adds:
\begin{itemize}\itemsep1pt
\item administrative \(\Let\)-bindings, erased by the lowering
  environment;
\item nullable \(\mathsf{yield}\), lowered to a sealed optional commit
  followed immediately by a public reveal of the optional value.
\end{itemize}

The lowering environment maps public surface variables to public core
expressions and sealed variables to sealed core variables.  This is why
surface lets do not allocate graph fields: later public references are
rewritten to the defining expression, while sealed commitments remain
stored values that can later be revealed.

\subsection{Static obligations}
\label{sec:source:wf}

The code distinguishes the obligations needed for graph construction
from the obligations needed for checked game semantics.

\begin{description}[itemsep=2pt,leftmargin=1.2em,style=nextline]
\item[\(\mathsf{GraphProgram}\).]
  Contains an initial visibility context, a core program, an initial
  environment, a well-formed-context proof, fresh-binder evidence, and
  normalized sample distributions.  This is the minimal input to
  \(\mathsf{compile}\).
\item[\(\mathsf{WFProgram}\).]
  Wraps a \(\mathsf{GraphProgram}\) and adds reveal completeness plus
  commit-guard legality.  Reveal completeness is a syntactic
  acquire--release discipline for sealed commitments: each \(\Commit\)
  acquires a pending secret name, each \(\Reveal\) releases a currently
  pending secret, and \(\Ret\) is allowed only once no pending secrets
  remain.  Legality states that each commit guard admits at least one
  action in every source-visible context.  These obligations are consumed
  by the game-facing progress layer, not by raw graph allocation.
\item[\(\mathsf{FiniteDomains}\).]
  Evidence that all operational value domains introduced by the program
  are finite.  This is needed to enumerate finite strategy carriers and
  finite machine states.  Terminal payoff integers need not themselves
  form a finite type.
\end{description}

This split is important.  A graph can be allocated from syntax plus
basic static hygiene, even before we know it is a good game.  A checked
game needs more: it must not leave commitments forever sealed
when payoffs expect public data, and it must not present a player with a
commitment point where every possible action is illegal.  Those are
semantic obligations about play, so they live in \(\mathsf{WFProgram}\)
rather than in the graph row format.

In this terminology, a \emph{checked program} is a
\(\mathsf{WFProgram}\), the input to the game-facing theorems.
Compilation turns a \(\mathsf{GraphProgram}\) into a
\(\mathsf{CompiledProgram}\) whose event graph is well formed; the
additional checked-program obligations are then used by progress,
frontier-game, and refinement theorems.

\leancorr{Core syntax is \TT{Vegas.Core.Basic}; it is continuation-passing
--- each effect bundles its continuation, typed in the extended context, which
is what makes every term well-scoped by construction, and is what the grammar's
\(s;~t\) denotes.  Checked-program bundles are in \TT{Vegas.WellFormed.Program}.
Surface syntax and let-erasing lowering are in \TT{Vegas.Language.Basic}.  The
concrete well-formedness predicates are collected by \TT{Vegas.Core.Obligations}.}
