\section{Event graphs}
\label{sec:eventgraphs}

The event graph is the central object of the construction.  It turns
the syntax tree of a checked program into a finite directed acyclic
graph of typed event nodes with explicit dependency edges, and
replaces the implicit ``state along the spine'' of an interpreter
with an extensional partial assignment from nodes to their write
slices.  The strategically relevant content of a configuration is
\emph{which} events have produced which results, not the order in
which they were produced.

\subsection{Nodes, fields, and slices}
\label{sec:eg:nodes-fields}

An \emph{event graph} is a tuple
\[
  \EG = \bigl(
    \Nodes, \Fields, \mathrm{fieldTy}, \mathrm{fieldOwner},
    \mathrm{initial}, \mathrm{sem}, \prereq, \rank,
    \mathrm{sliceLegal}, \mathrm{actionLegal}, \mathrm{internalKernel}
  \bigr)
\]
of finite typed-DAG data and three pieces of legality and execution
data.  We unpack each in turn.

\subsubsection*{Nodes and fields}

$\Nodes$ is a finite set of \emph{event nodes} with decidable
equality.  $\Fields$ is a finite set of \emph{typed storage fields},
each with a base type given by $\mathrm{fieldTy} : \Fields \to \Ty$
and an optional owner $\mathrm{fieldOwner} : \Fields \to
\mathrm{Option}(\Players)$.  The function $\mathrm{initial}(f)$
optionally provides an initial value for each field; absent fields
are uninitialised until written.

\subsubsection*{Stored values and write slices}

A field's storage is a \emph{stored value}: a value of the field's
base type, tagged as either \textsf{clear} (currently public) or
\textsf{hidden} (sealed).  A \emph{write slice} for a node $n$ is a
partial map
\[
  s : \Fields \;\rightharpoonup\; \mathrm{StoredValue},
\]
indicating which fields $n$ writes when it executes.  Every node
writes exactly one field — this convention, the
\emph{single-target} discipline, is key to the symbolic accounting
of who writes what.

\subsubsection*{Node semantics}

Each node carries a \emph{node semantic} payload
$\mathrm{sem}(n)$ in one of four shapes:
\[
  \mathrm{NodeSem} \;::=\;
    \mathsf{assign}
    \;\mid\;
    \mathsf{sample}
    \;\mid\;
    \mathsf{commit}_p
    \;\mid\;
    \mathsf{reveal}.
\]
Only $\mathsf{commit}_p$ has an actor; $\mathsf{assign}$,
$\mathsf{sample}$, and $\mathsf{reveal}$ are internal events of the
machine.  Several derived projections are used pervasively:
$\actor(n)$, $\readsOf(n)$ (a finset of fields read), and
$\writeOf(n)$ (the unique target of the node's write).

\subsubsection*{Prerequisites and rank}

The function $\prereq : \Nodes \to \mathrm{Finset}(\Nodes)$ records
the dependency relation, with the structural invariant that
$\prereq(n) \subseteq \Nodes$ and (\emph{strict topological order})
\[
  \forall n.\; \forall q \in \prereq(n).\; \rank(q) < \rank(n).
\]
$\rank : \Nodes \to \mathbb{N}$ is a strict topological-order
witness that serves as the termination measure for inductions over
graph executions.  The $\prereq$ relation, not the rank, is what
encodes \emph{semantic} dependency.

\subsubsection*{Three pieces of legality and execution data}

\begin{description}[itemsep=2pt]
\item[$\mathrm{sliceLegal}(n, s)$ — \emph{shape}.]
  Predicate on a node and a candidate write slice.  Purely
  structural: it checks that $s$ writes the expected field, with the
  expected base type, with the expected visibility (clear/hidden)
  for the node's $\mathrm{NodeSem}$.  Does not consult guards or
  results of other nodes.
\item[$\mathrm{actionLegal}(C, n, s)$ — \emph{dynamic}.]
  Predicate that may consult the result assignment $C$ of
  already-completed nodes.  For a player commit node, this is where
  the guard is evaluated against the proposed action and the
  current visible reads of the actor.
\item[$\mathrm{internalKernel}(n, C) \in \PMF{\Slice(n)}$.]
  Stochastic kernel for non-player ($\mathsf{assign}$,
  $\mathsf{sample}$, $\mathsf{reveal}$) nodes.  $\mathsf{assign}$
  and $\mathsf{reveal}$ kernels are Dirac; $\mathsf{sample}$ is the
  source-level distribution, evaluated against the public state.
\end{description}

\designnote{The split between \emph{shape} (\textsf{sliceLegal}) and
\emph{dynamic} (\textsf{actionLegal}) is the structural analogue of
the split between typing and the \emph{guard-uses-only}
predicate of
\S\ref{sec:source-language}: both put the part that can be checked
by a syntactic over-approximation in one place, and the part that
must consult the actual run-time state in another.  Internal kernels
are stochastic but never depend on the schedule, only on the result
assignment, which is what allows commit-internal commutation lemmas
to be stated as kernel equalities rather than as schedule-quantified
ones.}

\subsection{Configurations and the frontier}
\label{sec:eg:configurations}

\subsubsection*{Configurations}

A \emph{configuration} of $\EG$ is an extensional record of which
nodes have produced which results:
\begin{equation}
\label{eq:configuration}
  C : \Nodes \;\rightharpoonup\; \Slice(\cdot),
\end{equation}
satisfying three invariants.  (i) \emph{Result-nodes}: every $n$
with $C(n) = \mathsf{some}\,s$ is in $\Nodes$.  (ii)
\emph{Closed under prerequisites}: if $C(n) = \mathsf{some}\,s$ then
$C(q) = \mathsf{some}\,s_q$ for every $q \in \prereq(n)$.  (iii)
\emph{Locally legal}: for each completed node $n$, the recorded
slice satisfies $\mathrm{sliceLegal}(n, C(n))$.

\subsubsection*{The frontier}

The frontier of $C$ is the set of unfinished nodes whose
prerequisites have all completed:
\begin{equation}
\label{eq:frontier}
  \frontier(C) \;=\;
    \bigl\{\, n \in \Nodes \;\big|\;
      C(n) = \mathsf{none}
      \text{ and } \forall q \in \prereq(n).\, C(q) \neq \mathsf{none}
    \,\bigr\}.
\end{equation}
The frontier is \emph{computed}, not stored.  This single design
decision is responsible for most of the diamond properties below: a
schedule is not part of the configuration, so two linearisations
that finish the same set of nodes with the same per-node slices
yield literally the same configuration.

\begin{example}[Configurations of the Matching-Pennies graph]
For the Matching-Pennies graph of \S\ref{sec:intro:matching-pennies},
the initial configuration $C_0$ has $C_0(n) = \mathsf{none}$
everywhere, $\frontier(C_0) = \{\Commit_{\textsf{m}},
\Commit_{\textsf{m'}}\}$.  After both commits have completed (with
slices $s_{\textsf{m}}, s_{\textsf{m'}}$), the configuration $C_2$
records $C_2(\Commit_{\textsf{m}}) = s_{\textsf{m}}$ and
$C_2(\Commit_{\textsf{m'}}) = s_{\textsf{m'}}$, with
$\frontier(C_2) = \{\Reveal_{\textsf{m}}, \Reveal_{\textsf{m'}}\}$.
The order in which the two commits were scheduled is not visible to
$C_2$.
\end{example}

\subsection{Frontier preservation and the diamond}
\label{sec:eg:diamond}

We can now state the two structural theorems that justify the
extensional, scheduler-free design.

\begin{theorem}[Frontier preservation]
\label{thm:front-preserv}
Let $C$ be a configuration, and let $a, b \in \frontier(C)$ with
$a \neq b$.  If $s_a$ is a slice for $a$ with $\mathrm{sliceLegal}
(a, s_a)$, then in the successor configuration $C[a \mapsto s_a]$,
the node $b$ is still on the frontier.
\end{theorem}

\begin{theorem}[Frontier diamond]
\label{thm:diamond}
Let $C$ be a configuration, $a, b \in \frontier(C)$ with $a \neq b$,
and $s_a, s_b$ legal slices for $a, b$ respectively.  Then
\[
  \bigl(C[a \mapsto s_a]\bigr)[b \mapsto s_b]
  \;=\;
  \bigl(C[b \mapsto s_b]\bigr)[a \mapsto s_a]
\]
as configurations.
\end{theorem}

\begin{proof}[Proof sketch]
\Cref{thm:front-preserv}: $b$ is unfinished in $C$, hence in
$C[a\mapsto s_a]$ since $a \neq b$.  Its prerequisites are completed
in $C$ and remain completed in any extension, so $b$'s frontier
membership survives.  \Cref{thm:diamond}: configurations are
extensional functions, and the two updates write disjoint
node-keys; the resulting partial maps agree pointwise.  The closure
and local-legality invariants hold for either order by symmetry.
\end{proof}

\Cref{thm:diamond} is the formal sense in which the order of
independent events is irrelevant: \emph{any} two linearisations of
the same set of frontier choices produce the same configuration.

\subsection{Adequacy: lowering preserves declared payoffs}
\label{sec:eg:adequacy}

The elaboration $\mathsf{VegasCore} \to \mathsf{EventGraph}$ assigns
each source constructor an event node, with the SSA discipline
making the storage field of the unique writer of each variable
unambiguous.  Adequacy is the obligation that this preserves
declared payoffs.

\begin{theorem}[Adequacy]
\label{thm:adequacy}
Let $g$ be a checked program with payoff list $\overline{p}$ at its
$\Ret$.  For any terminal configuration $C$ of the elaborated graph
$\mathsf{eg}(g)$, the payoff-vector outcome assigned by the event graph
agrees with $\mathrm{evalPayoffs}(\overline{p}, \rho_C)$, where
$\rho_C$ is the public-projected environment induced by $C$.
\end{theorem}

The proof is a structural induction over the source program: each
source constructor's elaborated node performs the exact source-level
side-effect (read, sample, write, expose), so the public environment
read at $\Ret$ matches the source-level public environment at the
same program point.

\subsection{Commit/reveal information barrier}
\label{sec:eg:barrier}

\Cref{thm:diamond} says \emph{independent} events commute.  We also
need a \emph{dependency} theorem: that every reveal depends on every
prior commit, including those whose payload it does not directly
read.  This is what protects the strategic story from a
``$\Reveal$ races ahead of an unmade commitment'' pathology.

The elaborator emits, for each $\Reveal$ node, prerequisites
including:
\begin{itemize}\itemsep0pt
\item the commit node that wrote the revealed payload, and
\item every $\mathsf{commit}$ node that lexically precedes the
  reveal in the source order.
\end{itemize}
The second item is the \emph{phase barrier}; it is what distinguishes
the event-graph elaboration from a naive ``track only causal reads''
elaboration.  The next theorem records its strategic consequence.

\begin{theorem}[Commit/reveal information barrier]
\label{thm:barrier}
For any configuration $C$ and any reveal node $r$:
\[
  r \in \frontier(C)
  \;\Longrightarrow\;
  \forall c \in \mathrm{priorCommits}(r).\;
  C(c) \neq \mathsf{none}.
\]
Equivalently: a reveal cannot be frontier-simultaneous with any
unfinished prior commit.
\end{theorem}

\begin{corollary}[Strategy-surface form]
\label{cor:barrier-strat}
At any configuration $C$ where reveal $r$ is on the frontier, no
player action whose underlying node is a prior commit of $r$ is
available to its actor.
\end{corollary}

\Cref{thm:barrier} is a direct unfolding of the
prerequisite-closure invariant of configurations: $r \in
\frontier(C)$ requires every $q \in \prereq(r)$ to satisfy $C(q) =
\mathsf{some}\,s_q$, and every prior commit of $r$ is in $\prereq(r)$
by elaboration.  \Cref{cor:barrier-strat} re-expresses the same fact
in the strategic vocabulary that becomes available in
\S\ref{sec:fosg-strategies}.

\designnote{The elaborator could in principle emit a sparser
prerequisite relation that records only causal reads.  The barrier
fails for such a relation: a reveal that does not read a sibling's
commit payload could otherwise become enabled before the sibling
commits, allowing the sibling's strategy to condition on the
revealed value.  The phase prerequisites are the price of preserving
the simultaneous-move semantics that the protocol writer wrote
down.  This is a place where the dependency graph carries strictly
more information than the per-node causal read-set alone.}

\paragraph{Lean correspondence.}
Supplement~\S5 (event graph): \TT{NodeSem}, \TT{EventGraph},
\TT{Configuration}, \TT{frontier}, the diamond and barrier theorems
(\TT{withResult\_comm}, \TT{withResult\_mem\_frontier\_of\_ne},
\TT{eventGraph\_priorCommit\_not\_frontier\_with\_reveal},
\TT{prior\_commit\_not\_available\_when\_reveal\_frontier}).
