\section{Machines: an asynchronous probabilistic carrier}
\label{sec:machines}

The event graph of \S\ref{sec:eventgraphs} is a static description.
Game-theoretic constructions need an operational object: states,
actions, transitions, observations, and outcomes.  We introduce a
generic \emph{machine} carrier and exhibit the canonical event-graph
machine as one of its instances.  The same machine carrier is the
target of the backend refinement layer (\S\ref{sec:refinement}), so
backend implementations and the canonical event-graph semantics live
in the same universe.

\subsection{The machine carrier}
\label{sec:mach:carrier}

A \emph{machine} on a player set $\Players$ is a tuple
\[
  \Mach = \bigl(\;
    \State,\; \{\Action_p\}_p,\; \Internal,\; \Pubobs,\; \{\Obs_p\}_p,\;
    \Outcome,\;
    s_0,\;
    \avail,\; \availInt,\;
    \stepPlay,\; \stepInt,\;
    \term,\; \pubView,\; \mathrm{obs},\; \out,\; u
  \;\bigr)
\]
of a state space $\State$; per-player action types $\{\Action_p\}$;
an internal-event type $\Internal$; a public observation type
$\Pubobs$ and per-player observation types $\{\Obs_p\}$; an outcome
type $\Outcome$ with utility $u : \Outcome \to (\Players \to
\mathbb{R})$; an initial state $s_0$; two availability predicates
\begin{align*}
  \avail    &: (p : \Players) \to \Action_p \to \State \to \Prop, \\
  \availInt &: \Internal \to \State \to \Prop,
\end{align*}
two transition kernels
\begin{align*}
  \stepPlay &: (p : \Players) \to \Action_p \to \State \to \PMF{\State}, \\
  \stepInt  &: \Internal \to \State \to \PMF{\State},
\end{align*}
a terminal predicate $\term : \State \to \Prop$, and projections
$\pubView : \State \to \Pubobs$, $\mathrm{obs}_p : \State \to \Obs_p$,
$\out : \State \to \Outcome$.  A primitive \emph{event} is either
$(\mathsf{play}, p, a)$ for $a \in \Action_p$ or $(\mathsf{internal},
e)$ for $e \in \Internal$.

\designnote{The transition kernels are \emph{total}.  Legality of an
event is a separate predicate ($\avail$ or $\availInt$).  Total
kernels make composition and trace semantics easy to write:
$\stepPlay$ never returns a partial result.  The standard
implementation choice for unavailable events is to \emph{stutter} —
return $\delta_s$ at the source state — and the event-graph instance
below makes that choice.  But the abstract carrier does not commit
to it; a backend may legitimately reduce to an absorbing failure
state instead, and the refinement theorems still go through, as long
as the projection commutes with both choices on the spec side.}

\subsection{Schedulers, traces, and outcome kernels}
\label{sec:mach:traces}

A \emph{scheduler} (or \emph{event law}) is a state-indexed law
$\lambda : \State \to \PMF{\Event}$ that selects the next event from
each state.  We say $\lambda$ is \emph{legal} if its support at any
non-terminal state is contained in the available events.  Iterating
$\lambda$ for $n$ steps from $s_0$ yields a distribution over
length-$n$ event-state traces, $\mathrm{Tr}_n(\Mach, \lambda) \in
\PMF{\mathrm{RunPrefix}_n}$.  Marginalizing on the final state and
projecting via $\out$ yields the \emph{outcome kernel}
\[
  \mathrm{outcomeKernel}_\lambda \in \PMF{\Outcome}.
\]

The outcome kernel is the bridge between operational traces and
game-theoretic objects: every kernel game built later
(\S\ref{sec:fosg-strategies}) factors through some
$\mathrm{outcomeKernel}_\lambda$ for an appropriate scheduler
constructed from the strategy profile.

\subsection{The event-graph machine}
\label{sec:mach:eg-instance}

The event-graph instance is the canonical interpretation of an
event graph $\EG$ as a machine $\Mach_{\EG}$:
\begin{itemize}\itemsep1pt
\item $\State = \Cfg(\EG)$, the configurations of $\EG$.
\item $\Action_p = \{(n, s) : n \in \Nodes,\, \actor(n) = p,\,
  s \in \Patch(n)\}$, a frontier node owned by $p$ paired with a
  candidate field patch.
\item $\Internal = \{(n) : n \in \Nodes,\, \actor(n) = \bot\} \cup
  \{\mathsf{idle}\}$, a frontier non-player node, with an
  always-unavailable $\mathsf{idle}$ added to keep the type
  inhabited at terminal states.
\item $\avail(p, (n, s), C)$ holds iff $n \in \frontier(C)$,
  $\actor(n) = p$, $\mathrm{patchLegal}(n, s)$, and
  $\mathrm{actionLegal}(C, n, s)$.
\item $\stepPlay(p, (n, s), C) = \delta_{C[n \mapsto s]}$ when the
  action is available, $\delta_C$ (stuttering) otherwise.
\item $\stepInt(n, C) = \mathrm{internalKernel}(n, C) \bind
  \lambda s.\,\delta_{C[n \mapsto s]}$ when $n$ is enabled, $\delta_C$
  otherwise.
\item $\term(C)$ holds iff $\frontier(C) = \emptyset$.
\item $\out(C)$ is the payoff-vector outcome of a terminal configuration
  (the public projection of the payoff patch at the $\Ret$ node).
\item $\pubView$ projects $C$ to the visible-to-all public state;
  $\mathrm{obs}_p(C)$ projects to $p$'s view.
\end{itemize}

\Cref{thm:diamond,thm:barrier} immediately become statements about
$\Mach_{\EG}$: independent player events commute, and a player
action whose underlying node is a prior commit of an enabled reveal
is unavailable at any configuration where the reveal is enabled.

\begin{remark}[Stuttering vs.\ blocking]
\label{rem:stutter}
The choice to stutter on unavailable events makes $\stepPlay$ and
$\stepInt$ total without introducing an extra ``error'' state.  An
externally observable consequence is that strategies that prescribe
an unavailable action effectively waste a turn.  This is benign:
strategy carriers later (\S\ref{sec:fosg-strategies}) are restricted
to legal actions by construction, so the stuttering branch is
never reached for a strategy in the carrier — but it makes the
base machine semantics robust to off-carrier inputs, which is
useful for refinement reasoning.
\end{remark}

\paragraph{Lean correspondence.}
Supplement~\S6.1 (machine carrier), \S6.2 (run prefixes,
schedulers, traces), \S6.3 (the event-graph machine
interpretation).
