\section{Native round views, strategies, and kernel games}
\label{sec:fosg-strategies}

The shift from operational machine to game-theoretic structure
happens at this layer.  We define a native round view on top of a
machine, exhibit the event-graph round view, build pure and
PMF-behavioral strategy carriers from that view, and present the
game forms and kernel games over which the standard solution-concept
inventory is defined.  The FOSG view mirrors the same round
structure as a presentation layer; it is used by the FOSG, EFG, and
Kuhn bridges, not by the native strategy definitions.

\subsection{Why native round views, and where FOSG enters}

The factored-observation stochastic game formalism
\citep{Kovarik2022} indexes strategies by sequences of per-player
observations and packages chance and decision in the same
transition.  Two of its features fit the construction directly.
First, a player's epistemic state is given by their per-step
observation sequence — the same data that the visibility discipline
of \S\ref{sec:source-language} produces from the source program.
Second, FOSG admits a round-batched view of an asynchronous
machine: a single FOSG round collapses many primitive events of
the underlying machine into one transition, which is exactly the
shape we want for the event-graph machine.

The native strategic layer uses this round-batched shape directly; the
definitions do not route through FOSG.  FOSG provides the external
game-theory presentation used to reuse the GameTheory library's
FOSG-to-EFG bridge and finite Kuhn theorem.  The relationship between
the native round view and the FOSG presentation is therefore a bridge
theorem, not a definition.

\subsection{The native round view and FOSG presentation}
\label{sec:fosg:view}

A \emph{round view} on top of a machine $\Mach$ is a tuple
\[
  \mathcal{V} = \bigl(
    \{\Act_p\}_p,\;
    \activeSet : \State \to \mathrm{Finset}(\Players),\;
    \mathrm{availActions},\;
    \mathrm{transition},\;
    \mathrm{eventBlock}
  \bigr)
\]
of: a per-player round-action alphabet $\Act_p$ (independent of
$\Mach$'s action type); an \emph{active set} $\activeSet(s)$ at each
state; a per-player legal-action predicate
$\mathrm{availActions}(s, p) \subseteq \Act_p$; a transition
kernel
\[
  \mathrm{transition} : (s : \State) \times
    \{\,\sigma : \mathrm{JointAction}\,\Act \mid
      \mathrm{legal}(s, \sigma)\,\}
    \to \PMF{\State},
\]
and an event-block projection recording which primitive machine
events are represented by the round.  Two side conditions hold:
terminal states have empty
active sets, and non-terminal states admit at least one legal joint
action.

\subsubsection*{The event-graph round view}

Concretely, one native round of the event-graph machine
$\Mach_{\EG}$ is a canonical whole-frontier transition.  At a
configuration $C$, the active set is the set of players who own at
least one current frontier node.  Each active player selects a
round-action that contains candidate write slices; availability
checks only the slices for that player's owned nodes in
$\frontier(C)$.  The transition then executes every node currently
in $\frontier(C)$ exactly once, in the canonical
$\frontier(C).\mathrm{toList}$ order.  Player-owned nodes use the
selected player slices; internal nodes use the event graph's
internal kernels (deterministic for $\mathsf{assign}$ and
$\mathsf{reveal}$, stochastic for $\mathsf{sample}$).  Nodes that
become enabled while this list is being executed wait for a later
round.

The round-action alphabet is static: independently of the
configuration, $\Act_p$ is the type
\[
  \Act_p \;=\; (n : \Nodes) \to \mathrm{Slice}(n),
\]
with the player menu using $\mathsf{Option}(\Act_p)$ so that
$\mathsf{none}$ denotes \emph{not active in this round}.  The total
function is intentionally wider than the current menu: irrelevant
coordinates are ignored.  The configuration enters at the level of
\emph{availability}, not the alphabet.  At configuration $C$, the
available $\mathsf{some}\,a$ actions for $p$ are those whose slices
are legal on every node in
$\frontier(C) \cap \{n : \actor(n) = p\}$.  Keeping the alphabet
static means strategy types do not depend on the run-time
configuration; the per-state menu is recovered by intersecting with
the frontier.

The list order inside the round is definitional presentation data.
The frontier diamond and its whole-frontier linearization theorem
state that, for the same current frontier and slices, independent
frontier nodes reach the same endpoint regardless of admissible
linearization.  The code therefore does not rely on the canonical
list order carrying strategic information.

\subsubsection*{The stability condition}

The collapse from machine to FOSG is sound only when frontier
choices do not interfere across players.  This is the
\emph{frontier-stability} condition: if action $a$ for player $p$ is
available at $C$, $a$'s node is on $\frontier(C)$ with $a.n \neq
n^\star$, and any other frontier node $n^\star$ executes legally,
then $a$ is still available at the resulting configuration.

\begin{theorem}[Frontier stability for elaborated event graphs]
\label{thm:stability}
For every elaborated graph $\mathsf{eg}(g)$ of a checked program
$g$, frontier stability holds.
\end{theorem}

The proof is a case analysis on the kinds of frontier nodes:
distinct commits do not write to each other's storage (by SSA),
distinct internal nodes likewise, and the
$\mathrm{actionLegal}$ predicate of a commit reads only owner-visible
or public fields, none of which are written by other players'
frontier slices (by the guard-uses-only predicate).

\Cref{thm:stability} is what makes the round-batched view
well-defined: without it, two simultaneously selected frontier
choices could interfere, and the FOSG transition would be
ambiguous.

The corresponding FOSG view uses the same action alphabet, active
sets, available-action predicates, and transition kernel.  It adds the
reward and observation fields required by the generic FOSG library.

\subsection{Reachable information states and strategies}
\label{sec:fosg:strategies}

For the bounded native round view of a checked program, the set of
\emph{information states} of player $p$ is the set of equivalence
classes of bounded round histories under $p$'s observation.  We work with the
\emph{reachable} subset:
\[
  \mathrm{ReachableInfoState}(p) =
    \bigl\{\, I \in \InfoState_p \mid
      \exists h \in \mathrm{History}.\, h.\mathrm{view}_p = I \,\bigr\},
\]
i.e.\ those information states realized by \emph{some} history in
the native round view.  Reachability is a purely combinatorial property of
histories, not a property quantified over profiles; this keeps the
strategy carrier definable without circularly referring to the
profiles that range over it.

A \emph{pure strategy} for $p$ is a function
\[
  \sigma^{\mathrm{pure}}_p : \mathrm{ReachableInfoState}(p)
    \to \mathrm{Option}(\Act_p),
\]
returning $\mathsf{none}$ at information states where $p$ is not
active and $\mathsf{some}\,a$ at active ones.  A
\emph{behavioral strategy} is the analogous function valued in
$\PMF{\mathrm{Option}(\Act_p)}$.

The \emph{legal} carriers further restrict: at every reachable
information state where $p$ is active, the chosen $\mathsf{some}\,a$
satisfies $a \in \mathrm{availActions}(h, p)$ at every native history
that realizes that information state.  This is what closes the gap from program-level
\emph{Legal} (existence of a legal action at each commit site) to
\emph{strategic} legality (every strategy in the carrier prescribes
only legal actions): illegal moves are absent from the carrier, not
filtered after the fact.

\designnote{Reachability-pruning matters because illegal moves of
unreachable information states would otherwise pollute the
carrier without affecting any equilibrium.  Pruning is purely
combinatorial: it does not require quantifying over profiles.  The
carrier is still finite by the bounded horizon imposed in
\S\ref{sec:fosg:horizon}.}

\subsection{Bounded horizon and outcome kernels}
\label{sec:fosg:horizon}

A finite event graph admits a finite execution horizon: the source
program has a finite syntax-step count
$h(g) = |\,\mathrm{constructors of }\,g \text{ except }\Ret\,|$, and
no native round can advance the depth counter beyond $h(g)$.  Bounding
the round view by $h(g)$ produces finite strategy carriers and finite
outcome kernels.  The FOSG presentation has the same horizon for the
FOSG/EFG bridge theorems.

The bounded view exposes outcome kernels for both pure and
behavioral profiles:
\begin{align*}
  \mathrm{outcomeFromPure}_h\,\pi      &\in \PMF{\Outcome}, \\
  \mathrm{outcomeFromBehavioral}_h\,\beta &\in \PMF{\Outcome}.
\end{align*}

\subsection{Game forms and kernel games}
\label{sec:fosg:kernel-games}

We first separate \emph{game form} from \emph{preference}.  A game
form contains per-player strategy types, an outcome type, and a
profile-to-outcome kernel, but no utility.  Vegas exposes game
forms for payoff-vector outcomes, terminal public-state outcomes,
and blocked-trace outcomes.  The terminal-public-state game forms
are useful when the caller wants the final public observation itself
rather than only the payoff vector.

A \emph{kernel game} attaches utility to a game form and is the
minimal strategic-form carrier on top of which expected-utility
solution concepts make sense:
\begin{equation}
\label{eq:kernel-game}
  \KG \;=\; \bigl(\,
    \{S_p\}_p,\, O,\, K : {\textstyle\prod_p} S_p \to \PMF{O},\,
    u : O \to (\Players \to \mathbb{R})\,\bigr).
\end{equation}
Expected utility is then the obvious functional
\begin{equation}
\label{eq:eu}
  \EU(\Profile, p) \;=\;
  \mathbb{E}_{\omega \sim K(\Profile)}\bigl[u(\omega)(p)\bigr].
\end{equation}

For a checked program $g$, the payoff-vector and blocked-trace game
forms have canonical utilities, so four kernel games are exposed:
\begin{center}
\small
\begin{tabular}{@{}lp{0.55\textwidth}@{}}
\toprule
Name & Carrier and outcomes \\
\midrule
$\mathrm{pureKernelGame}(g)$ &
  pure strategies; outcomes in $\Outcome$ (public payoff). \\
$\mathrm{pmfBehavioralKernelGame}(g)$ &
  PMF behavioral strategies; outcomes in $\Outcome$. \\
$\mathrm{pureBlockedTraceKernelGame}(g)$ &
  pure strategies; outcomes in $\mathrm{BlockedTrace}(g)$. \\
$\mathrm{pmfBehavioralBlockedTraceKernelGame}(g)$ &
  PMF behavioral; outcomes in $\mathrm{BlockedTrace}(g)$. \\
\bottomrule
\end{tabular}
\end{center}

A \emph{blocked trace} is the sequence, indexed by native round, of
all primitive event-block executions that took place during that
round, paired with the configuration reached at the round's end.
It is strictly more informative than the payoff-vector outcome and
is the natural object on which to compare the canonical event-graph
machine to a backend implementation, since the latter may produce
additional implementation-specific events within the same round.
The two axes of the kernel-game table above are then \emph{strategy
carrier} (pure vs.\ PMF-behavioral) and \emph{outcome shape}
(payoff vector vs.\ blocked trace).  The blocked-trace games
\emph{refine} the payoff-vector games by an outcome projection.

\subsection{Trace-to-public agreement}
\label{sec:fosg:trace-public}

The blocked-trace games are useful for refinement reasoning
(\S\ref{sec:refinement}); the payoff-vector games are what users
usually reason about.  We need to know that they agree, both at
the kernel level and at the level of expected utility.

\begin{theorem}[Trace-to-public agreement]
\label{thm:trace-public}
For every checked program $g$, the payoff projection applied to
the blocked-trace kernel agrees with the payoff-vector outcome kernel,
\begin{align*}
  \mathrm{map}(\mathrm{publicOf})\,\circ\,
    K^{\mathrm{pure}}_{\mathrm{blocked}}
  &= K^{\mathrm{pure}}_{\mathrm{public}}, \\
  \mathrm{map}(\mathrm{publicOf})\,\circ\,
    K^{\mathrm{beh}}_{\mathrm{blocked}}
  &= K^{\mathrm{beh}}_{\mathrm{public}},
\end{align*}
and expected utilities are equal under the corresponding profile
inclusions.  Solution concepts (Nash, dominance) transport across
the projection: a strategy is best-response in the payoff-vector game
iff it is best-response in the blocked-trace game.
\end{theorem}

The proof packages the projection as a kernel-game
\emph{simulation} — a structural morphism that pushes outcome
distributions through a deterministic function while preserving
utility.  In this case the simulation is in fact an
\emph{EU-isomorphism}: the blocked-trace outcome retains
strategically inert per-event-block detail on top of the public
outcome, so the projection is information-losing but not
\emph{utility}-losing, and equilibrium predicates transport in both
directions.  This is the structural reason for the iff in the
trace-to-public statement, and is what distinguishes it from the
backend morphism of \S\ref{sec:refinement}, which is genuinely
one-way: the implementation may admit profiles that no spec
strategy realizes.

\subsection{Solution concepts}
\label{sec:fosg:solution}

The standard inventory is defined uniformly on
$\mathrm{pmfBehavioralKernelGame}(g)$.  We list the most important
predicates here and refer to the supplement for the complete table
and the corresponding existence corollaries.

\begin{itemize}\itemsep2pt
\item \textbf{Nash equilibrium.} $\mathrm{IsNash}(\Profile)$ iff
  $\forall p,\, \forall s'_p.\, \EU(\Profile, p) \ge
  \EU(\Profile[p \mapsto s'_p], p)$.
\item \textbf{Best response.} $\mathrm{IsBestResponse}(p, \Profile,
  s)$ iff $s$ maximizes $\EU(\Profile[p \mapsto \cdot], p)$.
\item \textbf{Dominance.} $\mathrm{WeaklyDominates}(p, s, t)$ iff
  $s$'s payoff weakly dominates $t$'s for $p$ uniformly in opponents'
  partial profiles.  Strict version with $>$.
\item \textbf{Correlated and coarse correlated equilibrium.}
  $\mathrm{IsCorrelatedEq}(\mu)$~\citep{Aumann1974} for a profile
  distribution $\mu$, and the coarse variant restricting to
  constant deviations.
\item \textbf{Pareto efficiency, social welfare.} $\Profile$ is
  Pareto efficient iff no $\Profile'$ weakly improves all players
  with strict improvement for some $p$.  Social welfare is
  $\sum_p \EU(\Profile, p)$.
\item \textbf{Potential games}~\citep{MonderShapley1996}. A potential
  $\Phi : \prod_p S_p \to \mathbb{R}$ tracking unilateral
  deviations exactly (exact potential) or in sign only (ordinal
  potential).
\item \textbf{Zero-sum and team games.}  Constant total payoff and
  identical per-player utility, respectively.
\end{itemize}

A mixed-extension profile is simply a profile of PMFs, one over each
player's strategy carrier.  Fixed-point existence theorems assume
finite nonempty strategy carriers and a finite outcome carrier.  The
library exposes finite-game existence corollaries
($\mathrm{mixedNash\_exists}$,
$\mathrm{correlatedEq\_exists}$, $\mathrm{coarseCorrelatedEq\_exists}$)
as wrappers around the GameTheory library's fixed-point and
correlated-equilibrium theorems.  For the payoff-vector game
$O = \Outcome\,\Players = \Players \to_0 \mathbb{Z}$,
finite outcomes are therefore a separate hypothesis; finite public-state
or bounded-history presentations are the natural carriers for
unconditional finite-game existence statements.

\paragraph{Lean correspondence.}
Supplement~\S6 (machines), \S7.2--\S7.3 (native round view, FOSG
presentation), \S8.1 (strategies and profiles, reachable information
states), \S8.2--\S8.3 (game forms, kernel-game carrier,
simulation/bisimulation), \S8.5 (solution concepts).
