\section{FOSG and EFG presentations; Kuhn equivalence}
\label{sec:presentations}

\subsection{FOSG as a presentation}

FOSG is the standard imperfect-information game-form presentation used
by the broader library.  VegasCore does not define its strategic
semantics by first translating to FOSG; native frontier games define the
Vegas-facing semantics.  Instead, the checked frontier round view is
exported as a bounded FOSG together with payoff-faithfulness theorems.

This is a presentation boundary.  A reader who already trusts FOSGs can
use the exported FOSG to compare VegasCore games with standard
imperfect-information models.  A reader auditing the compiler should
look one layer earlier, at the frontier round view over the event graph.

Although FOSG is a presentation rather than the semantic center, the
bridge is mathematically central: it is the standard presentation
through which the paper connects VegasCore to existing
imperfect-information game theory without making FOSG an implementation
IR.

\begin{theorem}[FOSG payoff adequacy]
\label{thm:fosg}
Let \(R\) be the canonical frontier round view of a checked program.  Two maps
connect native behavioral frontier profiles for \(R\) with profiles of the
exported FOSG --- a \emph{translation} from native to FOSG and a
\emph{restriction} from FOSG back to native --- and each preserves the joint
utility distribution.  Consequently the two games realize exactly the same set
of joint utility distributions: every native profile's distribution is attained
by its FOSG translation, and every FOSG profile's distribution is attained by
its restriction to a native profile.
\end{theorem}

\begin{proof}[Mechanized proof]
See \TT{Vegas.Theorems.FOSG}: the forward translation
(\TT{frontierFOSG\_sourcePayoff\_udist\_behavioral}) and the backward
restriction (\TT{frontierFOSG\_sourcePayoff\_udist\_restrictedBehavioral}) each
preserve the joint utility distribution.  The curated index
\TT{Vegas.Theorems.Claims} re-exports the forward direction
(\TT{claim\_fosg\_utility\_distribution\_adequacy}); both directions are proved
in \TT{Vegas.Theorems.FOSG}.
\end{proof}

\subsection{EFG presentation}
\label{sec:pres:efg}

The EFG layer serializes the bounded FOSG presentation through the
generic FOSG-to-Augmented-EFG bridge.  The resulting tree is a
presentation of the same payoff law.  It necessarily introduces a
sequential representation of simultaneous or checkpointed actions, so it
is not treated as the semantic owner.

This is exactly the point of deriving the EFG instead of writing it by
hand.  The tree is useful for export and comparison, but the theorem
attached to it records that its utility distribution is the same as the
frontier game it presents.

\begin{theorem}[EFG payoff adequacy]
\label{thm:efg}
Let \(R\) be the canonical frontier round view of a checked program.
For every native behavioral frontier profile for \(R\), the plain EFG
export of \(R\) preserves the joint utility distribution of the native
behavioral frontier game.
\end{theorem}

\begin{proof}[Mechanized proof]
See \TT{Vegas.Theorems.EFG}; the relevant statement is the
frontier plain-EFG payoff-distribution theorem for behavioral profiles.
\end{proof}

\subsection{Kuhn under the frontier information model}
\label{sec:pres:kuhn}

We use \emph{Kuhn realization} to mean the finite equivalence between
randomizing once over a complete contingent plan and randomizing locally
at each reached decision point.  The classical statement is for finite
trees~\citep{Kuhn1953}, and \citet{Heymann2022KuhnProduct} give a
measure-theoretic product-form generalization; ours is the finite frontier
setting, where we additionally mechanize the deviation-matching that
equilibrium transport needs.  In VegasCore this is
stated for the canonical frontier information model: players reason about
reachable checkpoint observations and available menus induced by the checked
program.  The main theorem has two directions.

Kuhn's equivalence requires \emph{perfect recall} --- a player never forgets
what they previously observed or chose.  The canonical frontier information
model has this by construction: a player's reachable information state is their
full checkpoint-observation history, and behavioral strategies are functions of
exactly that history, so the round view discards no information across rounds.
The round-view locality and canonical-information facts that record this are
mechanized (see the Lean-correspondence note below); the directions below and
the equilibrium transport they support both rest on it.

Informally, the theorem says that the two common ways to randomize a
strategy agree for these finite frontier games.  A player may either
sample a complete plan at the start, or randomize locally whenever their
view reaches a decision point.  VegasCore proves that these
presentations produce the same completed outcome laws, and that
one-player deviations can be translated across the boundary.

\begin{theorem}[Behavioral to product mixed pure]
\label{thm:kuhn-beh-to-mixed}
Every behavioral frontier profile has a product mixed-pure frontier
profile with the same payoff-vector outcome law.
\end{theorem}

\begin{theorem}[Product mixed pure to behavioral]
\label{thm:kuhn-mixed-to-beh}
Every product mixed-pure frontier profile has a behavioral frontier
profile with the same payoff-vector outcome law.  Moreover, around the
canonical mixed-to-behavioral realization, unilateral deviations on
either side have matching unilateral deviations on the other side with
the same outcome law on the deviating profile.
\end{theorem}

\begin{theorem}[Frontier Kuhn strategic equivalence]
\label{thm:kuhn-strategic-equivalence}
The canonical completed-frontier mixed-pure and behavioral games form a
strategy/deviation-level Kuhn equivalence:
\begin{enumerate}[leftmargin=1.6em,itemsep=1pt,topsep=2pt]
\item the two canonical profile translations preserve completed outcome
  kernels;
\item behavioral deviations from arbitrary behavioral profiles have matching
  mixed-pure deviations; and
\item mixed-pure or behavioral deviations from a canonical
  mixed-to-behavioral realization have matching deviations on the other side.
\end{enumerate}
\end{theorem}

\begin{proof}[Mechanized proofs]
See \TT{Vegas.Theorems.Kuhn}, especially the behavioral-to-mixed,
mixed-to-behavioral, and frontier-complete outcome Kuhn theorems.
\end{proof}

\Cref{thm:kuhn-beh-to-mixed,thm:kuhn-mixed-to-beh,thm:kuhn-strategic-equivalence}
are program-facing facts about the frontier games produced from
VegasCore, not EFG-export side effects.
The native completed-frontier RoundView/Kuhn development packages both
outcome-kernel equivalence and strategy/deviation content.  Nash-safe
transport is a corollary of these deviation-matching facts.  The FOSG
and EFG adequacy theorems, \Cref{thm:fosg,thm:efg}, present the same
frontier semantics to external game forms; they are not the semantic
proof route for this Kuhn statement.

\leancorr{Generic RoundView/FOSG and RoundView/EFG bridges are in
\TT{Vegas.Presentation.FOSG.FromRoundView} and
\TT{Vegas.Presentation.EFG.FromRoundView}.  Checked-program bridges are in
\TT{Vegas.Presentation.FOSG.FromCore} and
\TT{Vegas.Presentation.EFG.FromCore}; exports are in
\TT{Vegas.Export.FOSG} and \TT{Vegas.Export.EFG}.  Program-facing Kuhn
facts are in \TT{Vegas.Frontier.Kuhn} and
\TT{Vegas.Theorems.Kuhn}; the public strategy/deviation package is
\TT{frontierKuhnStrategicEquivalence\_schema}, with individual
deviation facts such as
\TT{mixedPureFrontier\_mixedDeviation\_to\_behavioral\_kernel} and
\TT{behavioralFrontier\_deviation\_to\_mixedPure\_kernel}.  Perfect recall of
the canonical information model is recorded by
\TT{claim\_behavioral\_strategies\_are\_roundView\_local} and
\TT{claim\_kuhn\_uses\_canonical\_roundView\_information} in
\TT{Vegas.Theorems.Claims}.}
