\section{Runtime refinement and trace adequacy}
\label{sec:refinement}

\subsection{Runtime-general refinement}

The refinement layer relates a concrete implementation machine to the
primitive event-graph machine.  It is deliberately runtime-general:
smart-contract runtimes, payment-channel state machines, and local
simulators are all natural backends, but the formal interface talks
about machines, states, observations, outcomes, and stochastic transition
kernels.  Smart-contract languages and blockchain mechanisms provide
important target examples, but they sit below this generic boundary
\citep{Sergey2019Scilla,Lamela2020Marlowe,Roughgarden2021EIP1559}.

A stochastic refinement supplies a projection from implementation states
to specification states and proves that implementation transition laws
project to the specification transition laws.  Internal implementation
detail may be erased by the projection.  What matters is preservation of
the observable machine behavior and of the probability laws used by the
strategic analysis.

The refinement relation is compositional.  If a low-level machine refines
an intermediate machine and the intermediate machine refines the
primitive event-graph machine, the two proofs compose to a single
low-level-to-primitive refinement.  The same is true of compatible
event-batch law families: a backend tower can add administrative audit
steps, encoded storage, message queues, or other runtime structure one
layer at a time, and then collapse those layers when applying the trace
transport theorem.

For this paper, the \emph{specification} is the primitive machine
obtained from the compiled event graph.  An \emph{implementation} is any
lower-level machine whose behavior should realize that specification.
For a contract backend, the implementation machine models
contract-level execution and runtime-visible observations; the projection
says which contract states and events count as which graph states and
events.  Events that are purely administrative may project to no
specification event, but then they must stutter at the specification
level.

\subsection{Trace game surfaces}
\label{sec:ref:trace-surfaces}

Runtime adequacy is stated uniformly for a selected frontier game
surface.  A \(\mathsf{TraceGameSurface}\) packages:
\begin{itemize}\itemsep1pt
\item a strategy-indexed kernel game to preserve;
\item a primitive event-batch trace kernel realizing that game on the
  compiled specification machine;
\item a theorem that mapping primitive traces to utilities gives the same
  joint utility distribution as the game.
\end{itemize}
The canonical surfaces are behavioral frontier play, pure frontier play,
and product mixed-pure frontier play.

The extra trace law is necessary because a refinement relates machines,
while game strategies live at the frontier round-view level.  The law is
the bridge saying: for this strategy profile, here is the distribution
over primitive event-batch traces that realizes the frontier game.  A
backend then proves that its own event-batch law is compatible with that
specification law.

A \(\mathsf{TraceSpecEventBatchLaw}\) then gives a profile-indexed
specification event-batch law whose finite trace distribution is exactly
the trace kernel of the selected surface.  A
\(\mathsf{RuntimeTraceAdequacy}\) adds an implementation law family and
compatibility with the stochastic refinement.

Legality of an event batch is support-wide: every intermediate state in
the support of a probabilistic prefix must make the remaining suffix
available.  This prevents a backend law from hiding an unavailable
intermediate branch that later converges to an otherwise acceptable final
state.  It is the batch-level counterpart of ordinary step availability.

\subsection{Transport theorem}
\label{sec:ref:transport}

The main transport statement is:
\begin{theorem}[Runtime trace adequacy]
\label{thm:runtime-trace}
Let \(S\) be a frontier trace game surface of a checked finite-domain
program, and suppose the data of a \(\mathsf{RuntimeTraceAdequacy}\) for
\(S\) is supplied: a stochastic refinement of the primitive compiled machine
by an implementation machine, a trace-adequate specification event-batch law
(\(\mathsf{TraceSpecEventBatchLaw}\)) whose finite trace distribution is the
trace kernel of \(S\), and an implementation event-batch law family
compatible with that refinement.  Then the implementation trace game has the
same joint utility distribution as \(S.\mathsf{game}\).
\end{theorem}

\begin{proof}[Mechanized proof]
See \TT{Vegas.Theorems.Refinement.implTraceGame\_udist\_surface}.
\end{proof}

A bounded-utility hypothesis upgrades this equality of utility distributions
to expected-utility equality.  Here ``bounded'' constrains the compared
\emph{trace} utilities, which at the runtime level may fold in economic costs;
the frontier payoffs themselves are bounded by construction
(\Cref{sec:prelim:notation}).  Under that hypothesis the implementation trace
game indexed by the same frontier strategy type has \emph{exactly} the Nash
equilibria of the selected frontier surface --- for the behavioral, pure, and
mixed-pure surfaces alike --- so the two equilibrium sets coincide and a
compatible backend introduces no spurious equilibria.

The order's asymmetry is therefore not between specification and
implementation at a fixed strategy space, but across strategy-space
\emph{enlargement}: implementation strategies with additional coordinates,
such as an extra message channel, must be ruled out by the observation model,
proved inert, or modeled on the specification side, and for them only the
one-way babbling/inert transport holds.  \Cref{sec:ref:cheap-talk} gives the
three sound positions.

\subsection{Extra messages and cheap talk}
\label{sec:ref:cheap-talk}

There is a subtle strategy-space mismatch here.  A concrete runtime may
permit players to exchange messages that are not part of the VegasCore
event graph: transaction metadata, mempool signals, off-chain messages,
or explicit no-op calls.  A refinement proof may erase those messages
operationally.  Strategic preservation additionally requires a relation
between the concrete strategy space and the frontier strategy space.  If
strategies in the concrete machine can condition on the extra messages,
then the concrete strategy space is strictly larger than the frontier
strategy space.

There are three sound ways to use such a runtime model.

\begin{description}[itemsep=2pt,leftmargin=1.2em,style=nextline]
\item[Closed-world assumption.]
  State that the analysis ignores out-of-band communication, or that the
  chosen implementation observation model does not expose it.  Then the
  refinement theorem applies to the selected strategy-indexed law family,
  with scope limited to behavior represented by that observation model.
\item[Inert extension.]
  Prove that the concrete game is an inert extension of the frontier
  game: every enriched strategy profile has exactly the same outcome law
  as its projection to the base frontier strategies.  This is the strong
  ``messages are genuinely irrelevant'' condition.  Under inertness,
  Nash equilibria are invariant at projected profiles, and base Nash
  equilibria embed into the enriched game.  Crucially, inertness covers
  unilateral enriched deviations, not merely the equilibrium profile:
  every such deviation has the same outcome law as its projection to a
  base-game deviation.
\item[Cheap-talk extension.]
  If the messages are observable and strategies may react to them, model
  the same message channel on the specification side and analyze the
  enlarged game.  The generic cheap-talk theorem gives the babbling
  result: every base Nash equilibrium embeds as an equilibrium where
  players send default messages and ignore the channel.  Cheap talk may
  also add coordination equilibria, so the theorem deliberately preserves
  the babbling equilibria rather than every equilibrium of the enlarged
  game.
\end{description}

The third option is often the right statement for protocols that run in
communicative environments because it makes the communication explicit.
The second option is stronger and requires a proof that messages cannot
affect the induced action law, even through contingent strategies.  A
runtime with arbitrary additional observable messages is
game-theoretically identical to the base frontier game only under an
inertness condition or an intentionally restricted babbling/no-talk
analysis.

\subsection{Abstract runtime examples}
\label{sec:ref:runtime-examples}

The identity and audited runtimes are minimal witnesses for the
abstraction.
The identity runtime is the baseline: the primitive compiled machine
refines itself, and any trace-adequate specification law yields runtime
adequacy by reflexivity.  The audited runtime executes an administrative
audit tick before specification event batches; the refinement erases
those ticks.  Both instantiate the generic backend boundary.

The message-in-flight runtime wraps a machine with public pending and
delivered message buffers.  Send and delivery events are implementation
events that refine away unless the specification deliberately models
them, but the buffers remain visible in runtime observations.  This makes
the model strong enough to express front-running-style behavior: a player
may condition an action on a pending message before it is delivered.  The
wrapper also prevents terminal traces from stranding pending messages:
terminal-producing source events are unavailable until the pending queue
has drained.

The protocol examples use these pieces in two different ways.  Inert
message coordinates are erased with no payoff effect, giving Nash
transport by inert-extension theorems.  Observable cheap talk is modeled
as an explicit enlarged protocol game; the compiled message runtime is
then proved faithful to that enlarged game.  Erasing the message channel
uses the inert or babbling hypotheses above.  The opaque commit/reveal
Matching-Pennies example
proves a scheduler invariant: every bounded message-runtime trace has no
extracted plaintext before both sealed moves are locked.  That invariant
is a protocol/runtime fact by construction; cryptographic hiding belongs
to a cryptographic backend assumption.

\leancorr{Generic machine refinement is in \TT{Vegas.Machine.Refinement}.  The
program-facing trace-adequacy surface is in
\TT{Vegas.Theorems.Refinement}: \TT{TraceGameSurface},
\TT{TraceSpecEventBatchLaw}, \TT{RuntimeTraceAdequacy},
\TT{behavioralFrontierTraceSurface}, \TT{pureFrontierTraceSurface},
\TT{mixedPureFrontierTraceSurface}, and the
\TT{implTraceGame\_udist\_surface} /
\TT{implTraceGame\_nash\_of\_surface\_nash} theorems.  The equilibrium-set
equalities are the \TT{claim\_runtime\_refinement\_*\_nash\_set\_eq} family
(behavioral, pure, and mixed-pure) and the matching \TT{\_nash\_iff} claims
in \TT{Vegas.Theorems.Claims}.  The generic
GameTheory support for this strategy-space issue is in
\TT{GameTheory.Core.InertExtension},
\TT{GameTheory.Concepts.Foundations.InertExtension},
\TT{GameTheory.Core.Babbling}, and
\TT{GameTheory.Concepts.Communication.Babbling}.  Runtime examples are in
\TT{Vegas.Machine.Audited}, \TT{Vegas.Machine.MessageInFlight},
\TT{Vegas.Machine.MessageSecurity}, and the message/refinement examples
under \TT{Vegas.Examples}.}
