\section{Discussion}
\label{sec:discussion}

\subsection{Architectural commitments}

Three architectural commitments shape the construction.

\paragraph{Dependency, not syntax, is the carrier of meaning.}
Meaning is factored through the event graph's dependency structure, not
the source syntax.  A consequence is that two source programs that
elaborate to the same event graph are strategically indistinguishable by
construction, not by an after-the-fact equivalence proof.  This frees the surface language to
add ergonomic constructors (simultaneous yields, nullable yields, future
handlers) without changing the strategic semantics; lowering is the only
translation that can change the event-graph shape.

This is also what makes backend work modular.  An implementation backend
does not need to justify every surface syntactic convenience separately;
it refines the compiled event graph.

\paragraph{Configurations are extensional.}
A configuration records which nodes have produced which results;
schedules are not part of the state.  This is what makes the
frontier diamond (\Cref{thm:diamond}) definitional rather than a
quotient construction.  The cost is that an interpreter that wants
to walk a specific schedule must supply that schedule explicitly as an
event law, batch witness, checkpoint policy, or runtime trace law.  The
graph itself does not privilege a source linearization.

\paragraph{Visibility is bound, not used.}
Every binding has a visibility tag; expression typing then
projects the context down to public or sealed as required.  The
discipline is owner-or-all: there is no group-visibility
constructor, and reveal-like operations introduce a fresh public
alias rather than mutating visibility.  Both decisions keep the
view algebra small (two erasures and a per-player view) and let
guard-uses-only-style predicates be about \emph{dependency
sets}, not about typing rules.

\paragraph{Faithfulness comes in three shapes.}
A single development can be held together by relations of genuinely
different algebraic character, and the guarantees map
(\Cref{fig:relations}) keeps them apart on purpose.  Schedule confluence
and Kuhn equivalence are \emph{equivalences} and move results freely in
either direction.  Compilation is an \emph{adequacy}: the graph is the
program's meaning, recovered by theorem, and there is no symmetric
``equal program'' relation hiding behind it.  Runtime refinement is an
\emph{order}, and its asymmetry is load-bearing rather than incidental ---
specification equilibria pull back to an implementation, but the converse
is a separate (and currently unproved) statement.  Reading the surface as
``a bunch of equivalence theorems'' loses exactly this structure; reading
it as one equivalence, one adequacy, and one order makes the end-to-end
guarantee a short composite.

\begin{remark}[A probabilistic-programming reading]
Structurally a VegasCore program is a probabilistic program whose latent
draws may be delegated to strategic players under the visibility discipline:
fixing strategies yields a distribution over outcomes, and ranging over them
yields a game.  This is the influence-diagram intuition---graphical models
of interacting decisions, of which multi-agent influence diagrams
\citep{koller2003multi} are the canonical game-theoretic form---but it is
only an intuition.  VegasCore and influence diagrams are \emph{incomparable}:
VegasCore adds an operational and refinement layer and a
sealed-then-revealed temporal discipline that influence diagrams do not
model, while the arbitrary information edges of an influence diagram are not
expressible under the owner-or-public visibility used here, where a
\(\Reveal\) discloses to everyone.  The reading is a lens on the denotation,
not a claim of containment.
\end{remark}

\subsection{Related and parallel artifacts}
\label{sec:disc:related-artifacts}

\paragraph{The \texttt{vegas} compiler.}
The Kotlin \texttt{vegas} compiler is parallel engineering work around
the same semantic boundary.  It explores concrete surface syntax,
typechecking, event-graph construction, analysis exports, and runtime
backends.  Where it uses the same vocabulary (event graphs, guards,
visibility, samples, and frontiers), the intended connection is a
certificate-producing boundary: compiled artifacts should expose the
VegasCore obligations they satisfy, such as freshness, reveal
completeness, distribution normalization, guard liveness, payoff
conservation, and backend refinement.

\paragraph{ConCertLean as a contract-runtime target.}
ConCertLean formalizes executable smart-contract chains in Lean:
account balances, serialized contract state, typed contracts, contract
calls, blocks, action queues, and induction principles for reachable
traces.  This makes it a plausible lower-level target for the
runtime-refinement API.  A backend package realizes a VegasCore event
graph as one or more ConCertLean contracts and proves projection
lemmas from ConCertLean chain states and calls to graph states and
events.  The same structure gives a useful proof template: entry-point
correctness, queue and block-step accounting, serialization soundness,
and trace induction.

\subsection{Limitations}
\label{sec:disc:limits}

\paragraph{Finiteness.}  Player sets, value domains, programs, event
graphs, and reachable strategy carriers are finite.  Distributions
are finite-support PMFs.  Continuous types and measure-theoretic
kernels~\citep{Giry1982,Kozen1981} are out of scope; the bridge
from $\FWeight{\cdot}$ to mathlib's $\PMF{\cdot}$~\citep{mathlib2020}
is one-way and takes a normalization proof, with no measure-theoretic
continuation.

\paragraph{No quit handlers.}  The surface language deliberately
omits user-level quit handlers.  A strategic quit changes the
action space and the control-flow shape, neither of which is a
value-level effect; supporting it requires adding a
prerequisite-cancellation rule to the event graph and revisiting
the frontier-read invariants.  Nullable yields are the language-level
substitute for ``opt out of revealing.''

\paragraph{Solution-concept inventory.}  The strategic API exposes
Nash, dominance, correlated and coarse correlated equilibria,
potential-game predicates, zero-sum/team predicates, social welfare,
and Pareto efficiency.  Refinement-based equilibrium concepts
(subgame-perfect and trembling-hand equilibria~\citep{Selten1975},
and sequential equilibria~\citep{KrepsWilson1982}) are not yet exposed
because the event-graph view does not maintain a sub-game decomposition;
this requires an additional presentation or subgame API rather than
only another predicate on the existing kernel games.

\paragraph{No source-native strategic semantics.}  Strategies, profiles,
and equilibria live on the compiled frontier games, not on
\(\mathsf{VegasCore}\) source.  The source layer supplies the
\emph{ingredients} of a strategy --- commit menus (the legal action set
at a choice point), per-player views, and choice-point classification ---
but no strategy object, source game, or source equilibrium is assembled
from them.  A consequence is that source-payoff adequacy (R2) is stated
for outcomes only; ``my program's equilibria are the frontier game's
equilibria'' holds by the \emph{definition} of strategic meaning (the
frontier construction), not as a theorem against an independent
source-native semantics.  Building source strategies and a source game,
then proving frontier adequacy for them, would be the strategic analogue
of R2 and is the most fundamental open item.

\paragraph{Presentation boundaries.}  The Vegas-facing pure and
PMF-behavioral strategy carriers are native round-view carriers over the
event-graph machine.  The bounded FOSG view is the standard
imperfect-information presentation, and the EFG layer is the serialized
tree presentation obtained from it.  The public Kuhn theorems are stated
and proved at the native completed-frontier level
(\Cref{thm:kuhn-beh-to-mixed,thm:kuhn-mixed-to-beh,thm:kuhn-strategic-equivalence}).
FOSG and EFG adequacy results
(\Cref{thm:fosg,thm:efg}) can be composed with those theorems when an
exported presentation is needed; they are bridges for exported
presentations, not the proof boundary for checked programs.

\paragraph{Concrete backend obligations.}  The refinement interface is
compositional and is exercised by abstract backend layers: identity,
audited bookkeeping, encoded storage examples, and message-in-flight
protocol runtimes.  EVM semantics, Solidity or Vyper compiler
correctness, a ConCertLean adapter, mempool and gas models, gas/fee
utility effects, and cryptographic commitment proofs are concrete
backend-package obligations below the generic refinement boundary.

\paragraph{Extra communication changes the game unless proved otherwise.}
Refinement can hide additional implementation messages from the projected
machine trace, but it does not automatically remove those messages from
the concrete players' strategy spaces.  A backend that allows arbitrary
observable messages must either make a closed-world/no-interesting-talk
assumption, prove that the extra strategic coordinates form an inert
extension, or expose a corresponding cheap-talk extension in the
specification and state only the babbling-equilibrium transport result.
This is not bookkeeping: cheap talk can create new non-babbling
equilibria even when the messages have no direct payoff.

\paragraph{Public samples only.}  The source $\Sample$ constructor
draws a public coin.  It is suitable for public randomness, but it
does not create private Harsanyi types or hidden chance moves.
Those require either a new hidden/private sample primitive or an
explicit external type parameter that a player commits to under the
visibility discipline.

\paragraph{Solver export.}  The finite pure kernel-game export object
contains player/strategy enumerations and kernel/utility oracles.  It is
deliberately narrower than a complete solver interchange format, and the
behavioral carrier is not exported as a finite strategy list.

\paragraph{Decidability.}  The graph-program obligations listed in
\Cref{sec:source:wf} are decidable: contextual well-formedness, fresh
bindings, and distribution normalization.  The checked-program
additions, reveal completeness and commit-guard legality, are also
decidable under the standing finite-domain assumption.  The Nash and
dominance predicates are quantifier-bounded over finite carriers, so
they are decidable in principle; whether they are computable in useful
time is a separate question \citep{DaskalakisGoldbergPapadimitriou2009}.

\subsection{Natural extensions}
\label{sec:disc:ext}

\paragraph{Mechanism design under epistemic types.}
Bayesian games with type
spaces~\citep{Harsanyi1967,MyersonGameTheory} are close to the
intended scope but need one more primitive: private type draws.
The $\Sample$ layer is public, so it cannot by itself
generate hidden per-player types.  A mechanism-design library is packaged
as a checked \TT{WFProgram} builder over a type-space
schema once the source language has either hidden samples or an
external typed-input interface.

\paragraph{Concurrent and partial-information protocols at scale.}
The event graph is naturally an event structure in the sense of
\citet{Winskel1986EventStructures,NielsenPlotkinWinskel1981}.
Concurrent-game-semantic and process-calculus
techniques~\citep{RideauWinskel2011Concurrent,
CastellanYoshidaSessionGame,
AbramskyJagadeesan1994,HylandOng2000PCF,Milner1999PiCalculus}
and session-typed multi-party
protocols~\citep{HondaYoshidaCarbone2008MPST} suggest two
extensions: typed channel structure beyond the
single-public-write SSA convention, and a richer notion of session
that tracks delegated authority.  Both are layerable on top of the
event-graph carrier.

\paragraph{Cryptographic backends.}
The commit/reveal pattern is operationally a cryptographic
commitment scheme~\citep{Blum1983CoinFlipping,
Naor1991BitCommitment,Pedersen1991Commitment}.  The
construction abstracts the commitment as a sealed-then-revealed
binding; a refinement layer that models hash-based or
homomorphic-commitment implementations is a natural backend
under stochastic refinement, with the cryptographic soundness
property appearing as a side condition on $\refines_s$.

\paragraph{Contract-runtime refinement packages.}
A backend package for an EVM-style runtime, a ConCertLean-style
executable chain, or a payment-channel state machine splits into several
layers.  The lower layers model enough runtime
execution to define states, observations, event batches, terminal
outcomes, and utility.  Intermediate layers add storage encodings,
queues, block/mempool scheduling, logs, or administrative accounting.  A
final proof composes the layer refinements into a
\(\mathsf{StochasticRefinement}\) from the concrete machine to the
compiled event-graph machine and instantiates
\(\mathsf{RuntimeTraceAdequacy}\) for the chosen frontier surface.  This
is the backend-facing use of the transport theorem
(\Cref{thm:runtime-trace}).  It keeps runtime-specific questions, such
as which observations are public, which administrative calls are
semantically invisible, which messages players may condition on, and
which economic costs affect utility, explicit in the backend proof
instead of hidden in the VegasCore language.

\paragraph{Compositional structure.}
Every checked program produces a kernel game, and morphisms between
kernel games are EU-preserving.  This suggests a categorical
structure on checked programs analogous to compositional game
theory~\citep{GhaniHedgesWinschel2018CGT,Hedges2016CGT};
making the composition explicit (parallel,
sequential, and conditional combinators on checked programs)
lets large protocols be assembled from small ones with
machine-checked strategic-structure preservation.

\paragraph{Concurrent-strategy realization.}
The frontier Kuhn bridge gives the native mixed-vs-behavioral
strategy/deviation correspondence for the completed frontier model.  In
a concurrent-strategy setting (where a player can commit to a function of \emph{joint}
information across simultaneous events), the corresponding
realization theorem is more delicate, and the formalism naturally
intersects logical accounts of partial information
\citep{HintikkaSandu1989,MannSanduSevenster2011IFL}.

\subsection{Outlook}

The construction addresses a small but central question: when the
strategic content of a finite multi-party protocol is mechanically
extracted from the straight-line program it is written as, and carried
toward a runtime, what holds the path together?  The answer it offers is
not a single object but a small family of preservation relations, pivoting
on a visibility-aware event graph with extensional configurations and a
checked frontier.  The strategic-form, FOSG, EFG, and runtime views are
then \emph{presentations} of this object that come with their own
adequacy and transport theorems
(\Cref{thm:fosg,thm:efg,thm:kuhn-beh-to-mixed,thm:kuhn-mixed-to-beh,thm:kuhn-strategic-equivalence,thm:runtime-trace}).

The deeper benefit is that extensions such as compositional combinators,
larger surface notation, richer cryptographic backends, and
mechanism-design packaging can all be added as new presentations, new
lowerings, or new backend refinement layers without touching the central
event-graph theorems.  Adequacy at each boundary is a checked theorem;
the inside of each boundary is free to use the representation that makes
the next proof tractable.
